---
title: "EcoFOCI Analysis Run 1-3"
author: "Zack Gold"
date: "2025-01-06"
output: html_document
---


```{r}
library(ggpmisc)

library(tidyverse)
library(here)
library(lubridate)
library(measurements)
library(sp)
library(readr)
library(sf)
library("rnaturalearth")
library("rnaturalearthdata")
library(cmocean)
library(phyloseq)

library(tidyverse)
library(phyloseq)
library(metagMisc)
library(iNEXT)
library(vegan)
library('phyloseq')
library(ranacapa)
library(here)
library(patchwork)

```

# Load Data
```{r}
long_COI_EcoFOCI <- readRDS(here("data","long_data","long_COI_EcoFOCI.RDS"))
long_MiFish_EcoFOCI<- readRDS(here("data","long_data","long_MiFish_EcoFOCI.RDS"))
long_Machida_EcoFOCI <- readRDS(here("data","long_data","long_Machida_EcoFOCI.RDS"))

```

# Site Mapping

```{r}
long_COI_EcoFOCI %>% ungroup() %>% 
  dplyr::select( Sample  ,         replicates   ,        group1   ,           
group2   ,            Sample_Name  ,        Biological_Replicate,
Technical_Replicate , Negative_control,     Cruise_ID_short  ,   
Cruise_ID_long ,      Cast_No.   ,          Rosette_position   , 
Station   ,           Sample_volume_ml ,    Time  ,              
Depth_m  ,            lat  ,                lon  ,               
Temp_C  ,             PSU_ppt  ,            OxySatPerc_percent  ,
OxyConc_umol.per.l ,  Chla_ugrams.per.l ,   PO4_umol.per.l  ,    
NO2_umol.per.l  ,     NO3_umol.per.l   ,    NH4_umol.per.l     , 
sites) %>% distinct() %>% 
  filter(., Cruise_ID_short !="RC_GOA_2022") %>% 
    mutate(., Depth_bin = if_else(Depth_m >29, "Below 30m","Above 30m" )) -> sample_metadata

```

```{r}
long_COI_EcoFOCI %>% 
  ungroup() %>% 
  dplyr::select(ASV_combo, Kingdom, Phylum, Class, Order, Family, Genus, Species, nReads, Sample) %>% distinct() -> quick_tax

quick_tax %>% 
  dplyr::select(ASV_combo, nReads, Sample) %>% 
  group_by(Sample) %>% 
  dplyr::summarise(tot_reads=sum(nReads)) -> sample_read_depths


sample_read_depths %>% 
  ggplot(aes(x=tot_reads))+geom_histogram()

sample_read_depths %>% 
  filter(., tot_reads < 30000) -> samples_to_drop
```


```{r}

sample_metadata %>% 
  ggplot(aes(x=Depth_m)) +   geom_histogram(aes(x=Depth_m,y=..density..), position="identity") + 
  geom_density(aes(x=Depth_m,y=..density..))
```

## Temp Map
```{r}

min_lat <- min(long_COI_EcoFOCI$lat, na.rm = T)
max_lat <- max(long_COI_EcoFOCI$lat, na.rm = T)

min_lon <- min(long_COI_EcoFOCI$lon, na.rm = T)
max_lon <- max(long_COI_EcoFOCI$lon, na.rm = T)

world <- ne_countries(scale = "medium", returnclass = "sf")

colnames(long_COI_EcoFOCI)

sample_metadata %>% 
  dplyr::select(Station, lat, lon) %>%  
  mutate(., lat=round(lat,1),
         lon=round(lon,1)) %>% 
  distinct() %>% 
  filter(., Station %in% c("M5","M8","DBO3.4")) -> sites_to_label

library(ggrepel)
ggplot(data = world) +
    geom_sf() +
    geom_point(data = sample_metadata, aes(x = lon, y = lat, colour=Temp_C), size=2) +scale_colour_cmocean(name="thermal") +
    coord_sf(xlim = c(min_lon-1, max_lon+1), ylim = c(min_lat-1, max_lat+1), expand = FALSE) +theme_bw() +xlab("Longitude") +ylab("Latitude") + facet_wrap(Depth_bin~Cruise_ID_short)+ geom_text_repel(data = sites_to_label, aes(x = lon, y = lat, label=Station),size =3,min.segment.length = unit(0, 'lines'), nudge_x = 4, nudge_y =  0.2)


```

## Salinity

```{r}
ggplot(data = world) +
    geom_sf() +
    geom_point(data = sample_metadata, aes(x = lon, y = lat, colour=PSU_ppt), size=2) +scale_colour_cmocean(name="haline") +
    coord_sf(xlim = c(min_lon-1, max_lon+1), ylim = c(min_lat-1, max_lat+1), expand = FALSE) +theme_bw() +xlab("Longitude") +ylab("Latitude") + facet_wrap(Depth_bin~Cruise_ID_short)+ geom_text_repel(data = sites_to_label, aes(x = lon, y = lat, label=Station),size =3,min.segment.length = unit(0, 'lines'), nudge_x = 4, nudge_y =  0.2)

```

## Chla
```{r}
ggplot(data = world) +
    geom_sf() +
    geom_point(data = sample_metadata, aes(x = lon, y = lat, colour=Chla_ugrams.per.l), size=2) +scale_colour_cmocean(name="algae", limits= c(0,15)) +
    coord_sf(xlim = c(min_lon-1, max_lon+1), ylim = c(min_lat-1, max_lat+1), expand = FALSE) +theme_bw() +xlab("Longitude") +ylab("Latitude") + facet_wrap(Depth_bin~Cruise_ID_short)+ geom_text_repel(data = sites_to_label, aes(x = lon, y = lat, label=Station),size =3,min.segment.length = unit(0, 'lines'), nudge_x = 4, nudge_y =  0.2)


sample_metadata 
```

## Oxy. Sat.

```{r}
ggplot(data = world) +
    geom_sf() +
    geom_point(data = sample_metadata, aes(x = lon, y = lat, colour=OxySatPerc_percent), size=2) +scale_colour_cmocean(name="oxy") +
    coord_sf(xlim = c(min_lon-1, max_lon+1), ylim = c(min_lat-1, max_lat+1), expand = FALSE) +theme_bw() +xlab("Longitude") +ylab("Latitude") + facet_wrap(Depth_bin~Cruise_ID_short)+ geom_text_repel(data = sites_to_label, aes(x = lon, y = lat, label=Station),size =3,min.segment.length = unit(0, 'lines'), nudge_x = 4, nudge_y =  0.2)

```


## Ammonium
```{r}
ggplot(data = world) +
    geom_sf() +
    geom_point(data = sample_metadata, aes(x = lon, y = lat, colour=NH4_umol.per.l), size=2) +scale_colour_cmocean(name="haline") +
    coord_sf(xlim = c(min_lon-1, max_lon+1), ylim = c(min_lat-1, max_lat+1), expand = FALSE) +theme_bw() +xlab("Longitude") +ylab("Latitude") + facet_wrap(Depth_bin~Cruise_ID_short)+ geom_text_repel(data = sites_to_label, aes(x = lon, y = lat, label=Station),size =3,min.segment.length = unit(0, 'lines'), nudge_x = 4, nudge_y =  0.2)
```



```{r}
long_COI_EcoFOCI %>% 
  filter(., !is.na(Species)) %>% 
  filter(., nReads >0) %>% 
  dplyr::group_by(ASV_combo, Kingdom, Phylum, Class, Order, Family, Genus, Species) %>% 
  dplyr::summarise(sum_reads=sum(nReads)) %>% 
  arrange(desc(sum_reads))

```

# CO1 Data Exploration

#### Build Phyloseq Object

##### ASV Reads

```{r}

varnames <- colnames(long_COI_EcoFOCI)
to_remove <- c("Run2:ASV","Run3:ASV","Run4:ASV","ASV_combo","Sequence","nReads","Prop.abund","eDNA.Index")


sample_metadata -> co1_sample_data

#Metadata
co1_sample_data %>%  as.data.frame() -> sampledata



rownames(sampledata) <- sampledata$Sample
sample_data(sampledata) -> sampledata

#ASV Reads

long_COI_EcoFOCI %>% 
  dplyr::select(ASV_combo,Sample, eDNA.Index) %>% 
  mutate(., eDNA.Index=as.numeric(eDNA.Index)) %>% 
  mutate(., eDNA.Index=if_else(eDNA.Index=="NaN", 0, eDNA.Index)) %>%
  pivot_wider(names_from = Sample, values_from = eDNA.Index, values_fill  =0) -> wide_PA

long_COI_EcoFOCI %>% 
  dplyr::select(ASV_combo, Kingdom,Phylum,Class, Order,Family, Genus, Species) %>% 
  distinct() %>% as.matrix() -> taxonomy_table

rownames(taxonomy_table) <- wide_PA$ASV_combo

TAX = tax_table(taxonomy_table)

wide_PA %>% 
  ungroup() %>% 
  dplyr::select(-ASV_combo) %>% as.matrix() -> otu_table
rownames(otu_table) <- wide_PA$ASV_combo

OTU = otu_table(otu_table, taxa_are_rows = TRUE)
physeq_obj._CO1 = phyloseq(OTU, TAX, sampledata)

surface_co1 <- subset_samples(physeq_obj._CO1, Depth_bin == "Above 30m")
subsurface_co1 <- subset_samples(physeq_obj._CO1, Depth_bin == "Below 30m")

physeq_obj._CO1_2020 <- subset_samples(physeq_obj._CO1,Cruise_ID_short != "SKQ2021")

subsurface_co1_2020 <- subset_samples(subsurface_co1, Cruise_ID_short != "SKQ2021")

surface_co1_2020 <- subset_samples(surface_co1, Cruise_ID_short != "SKQ2021")

```


# Alpha Diversity


```{r}


quick_tax %>% 
    group_by(Species) %>% 
  filter(., nReads >0) %>% 
  filter(., !is.na(Species)) %>% 
  dplyr::summarise(tot_ASVs = n_distinct(ASV_combo), tot_reads = sum(nReads), tot_occ= n_distinct(Sample)) %>% 
  arrange(desc(tot_occ))

quick_tax %>% 
    group_by(Genus) %>% 
  filter(., nReads >0) %>% 
  filter(., !is.na(Genus)) %>% 
  dplyr::summarise(tot_ASVs = n_distinct(ASV_combo), tot_reads = sum(nReads), tot_occ= n_distinct(Sample)) %>% 
  arrange(desc(tot_occ))
    

quick_tax %>% 
  filter(., Class=="Hexanauplia") %>% 
    group_by(Species) %>% 
  filter(., nReads >0) %>% 
  filter(., !is.na(Species)) %>% 
  dplyr::summarise(tot_ASVs = n_distinct(ASV_combo), tot_reads = sum(nReads), tot_occ= n_distinct(Sample)) %>% 
  arrange(desc(tot_occ))



quick_tax %>% 
  filter(., Species %in% c("Pseudocalanus mimus","Oithona similis","Acartia longiremis")) 

```


# Betdiveristy

## All
```{r}
physeq_obj._CO1_c = subset_samples(physeq_obj._CO1, !is.na(Temp_C) & !is.na(lat) & !is.na(Depth_m))

#Betadiversity
#Generate Vegan formatted data table
method.sampledf <- data.frame(sample_data(physeq_obj._CO1_c))

method.sampledf %>% 
  mutate(T2=as.Date(Time),
         month = month(Time)) %>% 
  mutate(., Season = case_when(month < 3 ~"Winter",
                               month < 6 ~"Spring",
                               month < 9~ "Summer",
                               month < 12 ~"Fall")) -> method.sampledf
method.rel_abun<- vegan_otu(physeq_obj._CO1_c)

#Jaccard dissimilarity matrix
method.d_carn <- vegdist(method.rel_abun, method="jaccard", binary=TRUE)

method.sampledf$Cruise_ID_short %>%  unique()
#PERMANOVA: Method+Site

method.adonis_results <- adonis2(method.rel_abun~Season+ Depth_m+lat+Temp_C+PSU_ppt+OxySatPerc_percent+OxyConc_umol.per.l+Chla_ugrams.per.l+PO4_umol.per.l+NO2_umol.per.l+NO3_umol.per.l+NH4_umol.per.l
, data=method.sampledf,  na.action =na.exclude, by = "terms", method="jaccard", binary=TRUE)

method.adonis_results

method.adonis_results_simple <- adonis2(method.rel_abun~Season +Depth_m+lat+Temp_C+PSU_ppt+OxyConc_umol.per.l+Chla_ugrams.per.l+lon, data=method.sampledf, by = "terms", method="jaccard", binary=TRUE)

method.adonis_results_simple

summary(method.adonis_results_simple)
```

```{r}


#Bray dissimilarity matrix
method.d_carn_b <- vegdist(method.rel_abun, method="bray")

#PERMANOVA: Method+Site

method.adonis_results_b <- adonis2(method.rel_abun~ Season + Depth_m+lat+Temp_C+PSU_ppt+OxySatPerc_percent+OxyConc_umol.per.l+Chla_ugrams.per.l+PO4_umol.per.l+NO2_umol.per.l+NO3_umol.per.l+NH4_umol.per.l
, data=method.sampledf,  na.action =na.exclude, by = "terms", method="bray")

method.adonis_results_b

method.adonis_results_simple_b <- adonis2(method.rel_abun~Season +Depth_m+lat+Temp_C+PSU_ppt+OxyConc_umol.per.l+Chla_ugrams.per.l+lon, data=method.sampledf, by = "terms",  method="bray")

method.adonis_results_simple_b

summary(method.adonis_results_simple)
```

## surface_co1 Jaccard
```{r}
surface_co1_c = subset_samples(surface_co1, !is.na(Temp_C) & !is.na(lat) & !is.na(Depth_m))

surface_co1_c = subset_samples(surface_co1_c, !Sample %in% samples_to_drop$Sample)


#Betadiversity
#Generate Vegan formatted data table
method.sampledf_s_co1_c <- data.frame(sample_data(surface_co1_c))


method.sampledf_s_co1_c %>% 
  mutate(T2=as.Date(Time),
         month = month(Time)) %>% 
  mutate(., Season = case_when(month < 3 ~"Winter",
                               month < 6 ~"Spring",
                               month < 9~ "Summer",
                               month < 12 ~"Fall")) -> method.sampledf_s_co1_c

method.rel_abun_s_co1_c<- vegan_otu(surface_co1_c)

#Jaccard dissimilarity matrix
method.d_carn_s_co1_c <- vegdist(method.rel_abun, method="jaccard", binary = TRUE)

#PERMANOVA: Method+Site

method.adonis_results_s_co1_c <- adonis2(method.rel_abun_s_co1_c~ Season + Depth_m+lat+Temp_C+PSU_ppt+OxySatPerc_percent+OxyConc_umol.per.l+Chla_ugrams.per.l+PO4_umol.per.l+NO2_umol.per.l+NO3_umol.per.l+NH4_umol.per.l
, data=method.sampledf_s_co1_c,  na.action =na.exclude, by = "terms")

method.adonis_results_s_co1_c

method.adonis_results_simple_s_co1_c <- adonis2(method.rel_abun_s_co1_c~Season + Depth_m+lat+Temp_C+PSU_ppt+OxyConc_umol.per.l+Chla_ugrams.per.l+lon, data=method.sampledf_s_co1_c, by = "terms")

method.adonis_results_simple_s_co1_c

summary(method.adonis_results_simple)
```


```{r}
ord <- ordinate(surface_co1_c, method = "NMDS", distance = ("jaccard"), binary = TRUE)
```

### NMDS Temp
```{r}
##Plot_Ordination
plot_ordination(surface_co1_c, ord, "samples", color = "Temp_C") +
  ggtitle(paste0("NMDS - Stress ",round(ord$stress,3))) + 
  geom_point(size = 4) + 
  theme_bw()+ 
  theme(plot.title = element_text(hjust = 0.5,size = 20, face = "bold"),
        plot.subtitle = element_text(hjust = 0.5,size = 18, face = "bold"),
        panel.grid.major = element_blank(), 
        panel.grid.minor = element_blank(),
        panel.background = element_blank(), 
        axis.text=element_text(size=16), 
        axis.title=element_text(size=16,face="bold"),
        legend.title = element_text( size=16, face="bold"),
        legend.text = element_text(size=14)
        ) + 
  labs(color = "Temp ˚C") +scale_colour_cmocean() 


```


### NMDS Cruise
```{r}
##Plot_Ordination
plot_ordination(surface_co1_c, ord, "samples", color = "Cruise_ID_short") +
  ggtitle(paste0("NMDS - Stress ",round(ord$stress,3))) + 
  geom_point(size = 4) + 
  theme_bw()+ 
  theme(plot.title = element_text(hjust = 0.5,size = 20, face = "bold"),
        plot.subtitle = element_text(hjust = 0.5,size = 18, face = "bold"),
        panel.grid.major = element_blank(), 
        panel.grid.minor = element_blank(),
        panel.background = element_blank(), 
        axis.text=element_text(size=16), 
        axis.title=element_text(size=16,face="bold"),
        legend.title = element_text( size=16, face="bold"),
        legend.text = element_text(size=14)
        ) + 
  labs(color = "Cruise")

```

## subsurface Jaccard
```{r}
subsurface_co1_c = subset_samples(subsurface_co1, !is.na(Temp_C) & !is.na(lat) & !is.na(Depth_m))

subsurface_co1_c = subset_samples(subsurface_co1_c, !Sample %in% samples_to_drop$Sample)

subsurface_co1_c = subset_samples(subsurface_co1_c, !Sample %in% c("E1720_DY2306", "E1723_DY2306", "E1948_SKQ23_12S" ,"E1949_SKQ23_12S"))




#Betadiversity
#Generate Vegan formatted data table
method.sampledf_sub_co1_c <- data.frame(sample_data(subsurface_co1_c))
method.rel_abun_sub_co1_c<- vegan_otu(subsurface_co1_c)


method.sampledf_sub_co1_c %>% 
  mutate(T2=as.Date(Time),
         month = month(Time)) %>% 
  mutate(., Season = case_when(month < 3 ~"Winter",
                               month < 6 ~"Spring",
                               month < 9~ "Summer",
                               month < 12 ~"Fall")) -> method.sampledf_sub_co1_c


#Jaccard dissimilarity matrix
method.d_carn_sub_co1_c <- vegdist(method.rel_abun_sub_co1_c, method="jaccard", binary=T)

method.adonis_results_sub_co1_c <- adonis2(method.rel_abun_sub_co1_c~ Season+ Depth_m+lat+Temp_C+PSU_ppt+OxySatPerc_percent+OxyConc_umol.per.l+Chla_ugrams.per.l+PO4_umol.per.l+NO2_umol.per.l+NO3_umol.per.l+NH4_umol.per.l
, data=method.sampledf_sub_co1_c,  na.action =na.exclude, by = "terms", method="jaccard", binary=T)

method.adonis_results_sub_co1_c

```

```{r}
ord <- ordinate(subsurface_co1_c, method = "NMDS", distance = ("jaccard"), binary="true")
```


```{r}
##Plot_Ordination
plot_ordination(subsurface_co1_c, ord, "samples", color = "Cruise_ID_short") +
  ggtitle(paste0("NMDS - Stress ",round(ord$stress,3))) + 
  geom_point(size = 2) + 
  theme_bw()+ 
  theme(plot.title = element_text(hjust = 0.5,size = 20, face = "bold"),
        plot.subtitle = element_text(hjust = 0.5,size = 18, face = "bold"),
        panel.grid.major = element_blank(), 
        panel.grid.minor = element_blank(),
        panel.background = element_blank(), 
        axis.text=element_text(size=16), 
        axis.title=element_text(size=16,face="bold"),
        legend.title = element_text( size=16, face="bold"),
        legend.text = element_text(size=14)
        ) + 
  labs(color = "Cruise")


```
```{r}
##Plot_Ordination
ord$points %>% 
  as.data.frame() %>% 
  filter(., MDS1 < -4)
# E1948_SKQ23_12S ,E1949_SKQ23_12S

ord$points %>% 
  as.data.frame() %>% 
  filter(., MDS2 > 1.5)

# E1720_DY2306, E1723_DY2306 

sample_data(subsurface_co1_c) <- method.sampledf_sub_co1_c

plot_ordination(subsurface_co1_c, ord, "samples", color = "Temp_C", shape="Season") +
  ggtitle(paste0("NMDS - Stress ",round(ord$stress,3))) + 
  geom_point(size = 2) + 
  scale_colour_cmocean() +
  theme_bw()+ 
  theme(plot.title = element_text(hjust = 0.5,size = 20, face = "bold"),
        plot.subtitle = element_text(hjust = 0.5,size = 18, face = "bold"),
        panel.grid.major = element_blank(), 
        panel.grid.minor = element_blank(),
        panel.background = element_blank(), 
        axis.text=element_text(size=16), 
        axis.title=element_text(size=16,face="bold"),
        legend.title = element_text( size=16, face="bold"),
        legend.text = element_text(size=14)
        ) + 
  labs(color = "Temp ˚C") + 
  stat_ellipse(type = "norm", linetype = 2) +
  stat_ellipse(type = "t") 


```
### CAP Analysis


## CAP Species

```{r, results="hide",warning=FALSE}
#Run CAP analysis to identify predictor species
vare.cap <- capscale(method.rel_abun_sub_co1_c~Season+Depth_m+Temp_C+lat+PSU_ppt+OxyConc_umol.per.l+lon, data=method.sampledf_sub_co1_c, dist="jaccard", binary=T)

#Retain species scores
sppscores(vare.cap) <- method.rel_abun_sub_co1_c

as.data.frame(vegan::scores(vare.cap, display="species")) %>% 
  rownames_to_column(var = "species") %>% 
  as.tibble %>%
 mutate(dist = sqrt((CAP1 - 0)^2 + (CAP2 - 0)^2)) -> vare.cap_species_distances


as.data.frame(vegan::scores(vare.cap, display="bp")) %>% 
  rownames_to_column(var = "env") %>% 
  as.tibble %>%
 mutate(dist = sqrt((CAP1 - 0)^2 + (CAP2 - 0)^2)) -> vare.cap_env_distances


```
##### Check For Break Point for Top Species
```{r}
species_distances <- as.data.frame(vare.cap_species_distances)
TAXcap = tax_table(subsurface_co1_c) %>% as.data.frame()

species_distances %>% 
  mutate(., ASV_combo=species) %>% 
  left_join(TAXcap) -> species_distances

rownames(species_distances) <- species_distances$species

# Now add the environmental variables as arrows
arrowmat <- species_distances

#Rename row names to clean up plot

species_distances %>% 
  filter(., !is.na(Species)) %>% 
  mutate(., name=str_c(ASV_combo,":",Species))-> namers

rownames(namers) <- namers$name


# Add labels, make a data.frame
arrowdf <- data.frame(labels = namers$Species, namers)

arrowdf %>% 
  filter(., !Phylum %in% c("Bacillariophyta","Chlorophyta","Dinophyceae__p","Dictyochophyceae__p","Pelagophyceae__p")) %>% 
  arrange(desc(dist)) %>% 
  dplyr::slice(1:100) 

arrowdf %>% 
arrange(desc(dist)) %>% 
    filter(., !Phylum %in% c("Bacillariophyta","Chlorophyta","Dinophyceae__p","Dictyochophyceae__p","Pelagophyceae__p")) %>% 
   dplyr::slice(1:10) %>% 
ggplot(., aes(x= reorder(labels, -dist), y=dist)) +
  geom_bar(stat="identity") + theme(axis.text.x = element_text(angle = 90, hjust = 1))

```

```{r}
arrowdf %>% 
arrange(desc(dist)) %>% 
    filter(., !Phylum %in% c("Bacillariophyta","Chlorophyta","Dinophyceae__p","Dictyochophyceae__p","Pelagophyceae__p")) %>% 
   dplyr::slice(1:6)-> arrowdf


summary(vare.cap)$cont %>% as.data.frame() %>% round(3)-> vare.cap_prop

as.tibble(vare.cap_prop,rownames = "rowname") %>% 
  pivot_longer(., cols=-rowname, values_to = "value", names_to = "name") %>% 
  filter(., rowname=="Proportion Explained") %>% 
  filter(., name %in% c("importance.CAP1","importance.CAP2"))-> vare.cap_prop_tib

# Add labels, make a data.frame
arrowdf_env <- data.frame(labels = vare.cap_env_distances$env, vare.cap_env_distances)

arrowdf_env %>% 
arrange(desc(dist)) %>% 
  filter(., env %in% c("SeasonSpring","lat", "Temp_C","Depth_m","PSU_ppt")) %>% 
  mutate(., labels = recode(labels,`lat`="Lat",
                            `Temp_C`="˚C",
                            `Depth_m` = "Depth",
                            `PSU_ppt`="PSU",
                            `SeasonSpring`="Season"))-> arrowdf_env

# Define the arrow aesthetic mapping
arrow_map <- aes(xend = 0.3 *CAP1, 
                 yend = 0.3 *CAP2, 
                 x = 0, 
                 y = 0, 
                 shape = NULL, 
                 color = NULL, 
                 label = labels)

arrow_map2 <- aes(xend = 2.3 *CAP1, 
                 yend = 2.3 *CAP2, 
                 x = 0, 
                 y = 0, 
                 shape = NULL, 
                 color = NULL, 
                 label = labels)

label_map <- aes(x = 2.3 * CAP1, 
                 y = 2.3 * CAP2, 
                 shape = NULL, 
                 color = NULL, 
                 label = labels,
                 alpha=0.7,
                 fontface="italic")
label_map2 <- aes(x = 0.27 * CAP1, 
                 y = 0.27 * CAP2, 
                 shape = NULL, 
                 color = NULL, 
                 alpha=0.7,
                 label = labels)

arrowhead = arrow(length = unit(0.02, "npc"))

method.sampledf$sample <- rownames(method.sampledf)
# Plot CAP
as.data.frame(vare.cap$CCA$wa) %>% 
  rownames_to_column(var = "sample") %>% 
  as.tibble() %>% 
  left_join(method.sampledf)-> data_cap


data_cap %>% 
  ggplot(aes(x=CAP1, y=CAP2)) + geom_point(aes(col=Temp_C, shape=Season), size =3) + geom_segment(
    mapping = arrow_map2, 
    size = .5, 
    data = arrowdf, 
    color = "dodgerblue3", 
    arrow = arrowhead
  ) +
  geom_label(
    mapping = label_map, 
    size = 4,  
        color = "dodgerblue3", 
    data = arrowdf, 
    show.legend = FALSE,
    position=position_jitter(width=0.03,height=0.03)
  ) + 
  geom_segment(
    mapping = arrow_map, 
    size = .5, 
    data = arrowdf_env, 
    color = "grey", 
    arrow = arrowhead
  ) +
  geom_label(
    mapping = label_map2, 
    size = 5, 
    data = arrowdf_env, 
    show.legend = FALSE,
    position=position_jitter(width=0.03,height=0.03)
  )  +
  ylab(paste0("CAP2 [",round(vare.cap_prop_tib$value[2]*100,1),"]%"))   + xlab(paste0("CAP1 [",round(vare.cap_prop_tib$value[1]*100,1),"]%")) +   
  scale_color_gradientn(colours = cmocean('thermal')(256), limits=c(-2,8)) + ggpubr::theme_pubr() +labs(colour="Temperature ˚C") + theme(axis.title = element_text(face="bold", size = 16))


ggsave(filename = here("gold_out","CAP_benthic_copepods.png"),
       width=12,
       height = 8,
       dpi = 400,
      units = c("in"))

```

## Copepod NMDS
```{r}
copepod_phy = subset_taxa(subsurface_co1_c, Class=="Hexanauplia")
copepod_phy = subset_samples(copepod_phy, !Sample %in% c("E1720_DY2306", "E1723_DY2306", "E1948_SKQ23_12S" ,"E1949_SKQ23_12S","E1890_SKQ23_12S","E1063_SKQ2021"))
copepod_phy = filter_taxa(copepod_phy, function(x) sum(x) > 0, TRUE)
copepod_phy <- prune_samples(sample_sums(copepod_phy) > 0, copepod_phy)
tax_table(copepod_phy)
```

## Copepod Bray
```{r}

#Betadiversity
#Generate Vegan formatted data table
method.sampledf_cop <- data.frame(sample_data(copepod_phy))
method.rel_abun_cop<- vegan_otu(copepod_phy)


method.sampledf_cop %>% 
  mutate(T2=as.Date(Time),
         month = month(Time)) %>% 
  mutate(., Season = case_when(month < 3 ~"Winter",
                               month < 6 ~"Spring",
                               month < 9~ "Summer",
                               month < 12 ~"Fall")) -> method.sampledf_cop


#Jaccard dissimilarity matrix
method.d_carn_cop <- vegdist(method.rel_abun_cop, method="jaccard", binary = T)

#PERMANOVA: Method+Site

method.adonis_results_cop <- adonis2(method.rel_abun_cop~Season+ Depth_m+lat+Temp_C+PSU_ppt+OxySatPerc_percent+OxyConc_umol.per.l+Chla_ugrams.per.l+PO4_umol.per.l+NO2_umol.per.l+NO3_umol.per.l+NH4_umol.per.l
, data=method.sampledf_cop,  na.action =na.exclude, by = "terms", method="jaccard")

method.adonis_results_cop



```

```{r}
ord <- ordinate(copepod_phy, method = "NMDS", distance = ("jaccard"), binary=T)
```

### NMDS Temp
```{r}

ord$points %>% 
  as.data.frame() %>% 
  filter(., MDS1 >0.2 )
sample_data(copepod_phy) <- method.sampledf_cop

##Plot_Ordination
plot_ordination(copepod_phy, ord, "samples", color = "Temp_C", shape="Season") +
  ggtitle(paste0("NMDS - Stress ",round(ord$stress,3))) + 
  geom_point(size = 4) + 
  theme_bw()+ 
  theme(plot.title = element_text(hjust = 0.5,size = 20, face = "bold"),
        plot.subtitle = element_text(hjust = 0.5,size = 18, face = "bold"),
        panel.grid.major = element_blank(), 
        panel.grid.minor = element_blank(),
        panel.background = element_blank(), 
        axis.text=element_text(size=16), 
        axis.title=element_text(size=16,face="bold"),
        legend.title = element_text( size=16, face="bold"),
        legend.text = element_text(size=14)
        ) + 
  labs(color = "Temp ˚C") +scale_colour_cmocean() 


```
### CAP Analysis

<br />
```{r}
#Constrained Analysis of Principle Components
cap_ord <- ordinate(
  physeq = copepod_phy, 
  method = "CAP",
  distance = method.d_carn_cop,
  formula = ~ method.sampledf_cop$Season + method.sampledf_cop$Depth_m+method.sampledf_cop$lat+method.sampledf_cop$Temp_C+method.sampledf_cop$PSU_ppt+method.sampledf_cop$OxyConc_umol.per.l+method.sampledf_cop$Chla_ugrams.per.l)
#Results of model
anova(cap_ord)
summary(cap_ord)
```


```{r}

cap_plot <- plot_ordination(
  physeq = copepod_phy, 
  ordination = cap_ord, 
    color = "Temp_C", 
    axes = c(1,2)
) + 
    geom_point(aes(colour = Temp_C), alpha = 0.8, size = 4) + 
  scale_color_gradientn(colours = cmocean('thermal')(256))+
    geom_point(colour = "grey90", size = 1.5) +
  scale_shape_manual(values = c(15,16,17,18))


# Now add the environmental variables as arrows
arrowmat <- vegan::scores(cap_ord, display = "bp")
arrowmat <- arrowmat[1:6,]
# Add labels, make a data.frame
rownames(arrowmat) -> namers

arrowdf <- data.frame(labels = stringr::str_remove(namers, "method.sampledf\\$"), arrowmat)

# Define the arrow aesthetic mapping
arrow_map <- aes(xend = CAP1, 
    yend = CAP2, 
    x = 0, 
    y = 0, 
    shape = NULL, 
    color = NULL, 
    label = labels)

label_map <- aes(x = 1.3 * CAP1, 
    y = 1.3 * CAP2, 
    shape = NULL, 
    color = NULL, 
    label = labels)

arrowhead = arrow(length = unit(0.02, "npc"))

# Make a new graphic
cap_plot + 
  geom_segment(
    mapping = arrow_map, 
    size = .5, 
    data = arrowdf, 
    color = "gray", 
    arrow = arrowhead
  ) + 
  geom_text(
    mapping = label_map, 
    size = 4,  
    data = arrowdf, 
    show.legend = FALSE
  ) +theme_bw()


```

```{r}
ord <- ordinate(subsurface_co1_2020, method = "NMDS", distance = ("bray"))
```


```{r}
##Plot_Ordination
plot_ordination(subsurface_co1_2020, ord, "samples", color = "Temp_C") +
  ggtitle(paste0("NMDS - Stress ",round(ord$stress,3))) + 
  geom_point(size = 4) + 
  theme_bw()+ 
  theme(plot.title = element_text(hjust = 0.5,size = 20, face = "bold"),
        plot.subtitle = element_text(hjust = 0.5,size = 18, face = "bold"),
        panel.grid.major = element_blank(), 
        panel.grid.minor = element_blank(),
        panel.background = element_blank(), 
        axis.text=element_text(size=16), 
        axis.title=element_text(size=16,face="bold"),
        legend.title = element_text( size=16, face="bold"),
        legend.text = element_text(size=14)
        ) + 
  labs(color = "Temp ˚C") +scale_colour_cmocean() 


```

# To Print



```{r}

varnames <- colnames(long_COI_EcoFOCI)
to_remove <- c("Run2:ASV","Run3:ASV","Run4:ASV","ASV_combo","Sequence","nReads","Prop.abund","eDNA.Index")
c(setdiff(varnames,to_remove),"Tot") -> varnames_to_group

varnames_to_group
long_COI_EcoFOCI %>% 
  dplyr::select(-eDNA.Index, -Prop.abund) %>% 
  ungroup() %>% 
 group_by(Sample) %>%
    mutate(Tot = sum(nReads)) %>% 
    group_by(!!!syms(varnames_to_group)) %>% 
    dplyr::summarise(nReads=sum(nReads)) %>% 
    mutate (Prop.abund = nReads / Tot) %>% 
     unite(sum.taxonomy, c("Kingdom",              "Phylum",               "Class"               
,"Order",                "Family",               "Genus",               "Species"), sep=";") %>% 
    group_by (sum.taxonomy) %>%
    filter(., nReads >0) %>% 
  group_by(sum.taxonomy) %>% 
  dplyr::summarise(Total_reads = sum(nReads), Maxprop= max(Prop.abund), N_detections = n_distinct(Sample_Name))  %>% 
  arrange(sum.taxonomy) -> co1_to_print


write_excel_csv(co1_to_print, file = here("co1_2_print"))
```

```{r}

varnames <- colnames(long_MiFish_EcoFOCI)
to_remove <- c("Run2:ASV","Run3:ASV","Run4:ASV","ASV_combo","Sequence","nReads","Prop.abund","eDNA.Index")
c(setdiff(varnames,to_remove),"Tot") -> varnames_to_group

varnames_to_group
long_MiFish_EcoFOCI %>% 
  dplyr::select(-eDNA.Index, -Prop.abund) %>% 
  ungroup() %>% 
 group_by(Sample) %>%
    mutate(Tot = sum(nReads)) %>% 
    group_by(!!!syms(varnames_to_group)) %>% 
    dplyr::summarise(nReads=sum(nReads)) %>% 
    mutate (Prop.abund = nReads / Tot) %>% 
     unite(sum.taxonomy, c("Kingdom",              "Phylum",               "Class"               
,"Order",                "Family",               "Genus",               "Species"), sep=";") %>% 
    group_by (sum.taxonomy) %>%
    filter(., nReads >0) %>% 
  group_by(sum.taxonomy) %>% 
  dplyr::summarise(Total_reads = sum(nReads), Maxprop= max(Prop.abund), N_detections = n_distinct(Sample_Name))  %>% 
  arrange(sum.taxonomy) -> mifish_to_print


write_excel_csv(mifish_to_print, file = here("mifish_2_print"))
```


## Group by Species

#### Build Phyloseq Object

```{r}

varnames <- colnames(long_COI_EcoFOCI)
to_remove <- c("Run2:ASV","Run3:ASV","Run4:ASV","ASV_combo","Sequence","nReads","Prop.abund","eDNA.Index")
c(setdiff(varnames,to_remove),"Tot") -> varnames_to_group


sample_metadata -> co1_sample_data

#Metadata
co1_sample_data %>%  as.data.frame() -> sampledata
rownames(sampledata) <- sampledata$Sample
sample_data(sampledata) -> sampledata

#ASV Reads


long_COI_EcoFOCI %>% 
  dplyr::select(-eDNA.Index, -Prop.abund) %>% 
  ungroup() %>% 
 group_by(Sample) %>%
    mutate(Tot = sum(nReads)) %>% 
    group_by(!!!syms(varnames_to_group)) %>% 
    dplyr::summarise(nReads=sum(nReads)) %>% 
    mutate (Prop.abund = nReads / Tot) %>% 
      ungroup() %>% 
    unite(sum.taxonomy, c("Kingdom",              "Phylum",               "Class"               
,"Order",                "Family",               "Genus",               "Species"), sep=";") %>% 
    group_by (sum.taxonomy) %>%
    mutate (Colmax = max(Prop.abund),
            eDNA.Index = Prop.abund / Colmax) %>% 
  dplyr::select(sum.taxonomy,Sample, eDNA.Index) %>% 
  mutate(., eDNA.Index=as.numeric(eDNA.Index)) %>% 
  mutate(., eDNA.Index=if_else(eDNA.Index=="NaN", 0, eDNA.Index)) %>%
  pivot_wider(names_from = Sample, values_from = eDNA.Index, values_fill  =0) -> wide_PA

wide_PA %>% 
  dplyr::select(sum.taxonomy) %>% 
   separate(sum.taxonomy, into=c("Kingdom",              "Phylum",               "Class"               
,"Order",                "Family",               "Genus",               "Species"), sep=";", remove=F) %>%  
  dplyr::select(sum.taxonomy, Kingdom,Phylum,Class, Order,Family, Genus, Species) %>% 
  distinct() %>% as.matrix() -> taxonomy_table
dim(taxonomy_table)
dim(wide_PA)
rownames(taxonomy_table) <- wide_PA$sum.taxonomy

TAX = tax_table(taxonomy_table)

wide_PA %>% 
  ungroup() %>% 
  dplyr::select(-sum.taxonomy) %>% as.matrix() -> otu_table
rownames(otu_table) <- wide_PA$sum.taxonomy

OTU = otu_table(otu_table, taxa_are_rows = TRUE)
physeq_obj._CO1_sum.tax = phyloseq(OTU, TAX, sampledata)

subsurface_co1_sum.tax <- subset_samples(physeq_obj._CO1_sum.tax, Depth_bin == "Below 30m")

surface_co1_sum.tax <- subset_samples(physeq_obj._CO1_sum.tax, Depth_bin == "Above 30m")


```


##  All Surface Bray

```{r}


surface_co1_sum.tax <- prune_samples(sample_sums(surface_co1_sum.tax) > 0, surface_co1_sum.tax)
surface_co1_sum.tax_sp = subset_samples(surface_co1_sum.tax, !is.na(Temp_C) & !is.na(lat) & !is.na(Depth_m))
surface_co1_sum.tax_sp = subset_taxa(surface_co1_sum.tax_sp, Kingdom!="Bacteria")
surface_co1_sum.tax_sp = subset_taxa(surface_co1_sum.tax_sp, Kingdom!="Environmental Unknown")
surface_co1_sum.tax_sp2 = filter_taxa(surface_co1_sum.tax_sp, function(x) sum(x > 0) > (0.1*length(x)), TRUE)


tax_table(surface_co1_sum.tax_sp) %>%  as.data.frame()
```


```{r}

#Betadiversity
#Generate Vegan formatted data table
method.sampledf <- data.frame(sample_data(surface_co1_2020_sum.tax_sp2))
method.rel_abun<- vegan_otu(surface_co1_2020_sum.tax_sp2)

#Jaccard dissimilarity matrix
method.d_carn <- vegdist(method.rel_abun, method="bray")

#PERMANOVA: Method+Site

method.adonis_results_simple <- adonis2(method.rel_abun~Depth_m+Temp_C+lat+PSU_ppt+OxyConc_umol.per.l+lon, data=method.sampledf, by = "terms")

method.adonis_results_simple

```

```{r}
ord <- ordinate(surface_co1_2020_sum.tax_sp2, method = "NMDS", distance = ("bray"))
```

### NMDS Temp
```{r}
##Plot_Ordination
plot_ordination(surface_co1_2020_sum.tax_sp2, ord, "samples", shape="Cruise_ID_short",color = "Temp_C") +
  ggtitle(paste0("NMDS - Stress ",round(ord$stress,3))) + 
  geom_point(size = 4) + 
  theme_bw()+ 
  theme(plot.title = element_text(hjust = 0.5,size = 20, face = "bold"),
        plot.subtitle = element_text(hjust = 0.5,size = 18, face = "bold"),
        panel.grid.major = element_blank(), 
        panel.grid.minor = element_blank(),
        panel.background = element_blank(), 
        axis.text=element_text(size=16), 
        axis.title=element_text(size=16,face="bold"),
        legend.title = element_text( size=16, face="bold"),
        legend.text = element_text(size=14)
        ) + 
  labs(color = "Temp ˚C") +scale_colour_cmocean() 


```
### CAP Analysis


## CAP Species

```{r, results="hide",warning=FALSE}
#Run CAP analysis to identify predictor species
vare.cap <- capscale(method.rel_abun~Depth_m+Temp_C+lat+PSU_ppt+OxyConc_umol.per.l+lon, data=method.sampledf, dist="bray")

#Retain species scores
sppscores(vare.cap) <- method.rel_abun

as.data.frame(vegan::scores(vare.cap, display="species")) %>% 
  rownames_to_column(var = "species") %>% 
  as.tibble %>%
 mutate(dist = sqrt((CAP1 - 0)^2 + (CAP2 - 0)^2)) -> vare.cap_species_distances


as.data.frame(vegan::scores(vare.cap, display="bp")) %>% 
  rownames_to_column(var = "env") %>% 
  as.tibble %>%
 mutate(dist = sqrt((CAP1 - 0)^2 + (CAP2 - 0)^2)) -> vare.cap_env_distances


```
##### Check For Break Point for Top Species
```{r}
species_distances <- as.data.frame(vare.cap_species_distances)
rownames(species_distances) <- species_distances$species

# Now add the environmental variables as arrows
arrowmat <- species_distances

#Rename row names to clean up plot

rownames(species_distances) %>% as.data.frame() -> namers
colnames(namers) <- c("path")
namers %>% separate(path, c("D1","P1","C1","O1","F1","G1","Species"), sep=";") %>% replace(is.na(.), "") %>% 
mutate(., name = ifelse(Species == "NA", G1, Species)) %>% 
  mutate(., name = ifelse(name == "NA", F1, name)) %>% 
  mutate(., name = ifelse(name == "NA", O1, name)) %>% 
  mutate(., name = ifelse(name == "NA", C1, name)) %>% 
  mutate(., name = ifelse(name == "NA", P1, name)) %>% 
  mutate(id = row_number(),
         name=str_c(name,"_",id))-> namers

rownames(arrowmat) <- namers$name


# Add labels, make a data.frame
arrowdf <- data.frame(labels = rownames(arrowmat), arrowmat)

arrowdf %>% 
arrange(desc(dist)) %>% 
ggplot(., aes(x= reorder(labels, -dist), y=dist)) +
  geom_bar(stat="identity") + theme(axis.text.x = element_text(angle = 90, hjust = 1))

```

```{r}

arrowdf %>% 
arrange(desc(dist)) %>% 
  slice(1:5) -> arrowdf


summary(vare.cap)$cont %>% as.data.frame() %>% round(3)-> vare.cap_prop

as.tibble(vare.cap_prop,rownames = "rowname") %>% 
  pivot_longer(., cols=-rowname, values_to = "value", names_to = "name") %>% 
  filter(., rowname=="Proportion Explained") %>% 
  filter(., name %in% c("importance.CAP1","importance.CAP2"))-> vare.cap_prop_tib

# Add labels, make a data.frame
arrowdf_env <- data.frame(labels = vare.cap_env_distances$env, vare.cap_env_distances)

arrowdf_env %>% 
arrange(desc(dist)) %>% 
  slice(1:5) -> arrowdf_env

# Define the arrow aesthetic mapping
arrow_map <- aes(xend = 0.5 *CAP1, 
                 yend = 0.5 *CAP2, 
                 x = 0, 
                 y = 0, 
                 shape = NULL, 
                 color = NULL, 
                 label = labels)

label_map <- aes(x = 0.4 * CAP1, 
                 y = 0.4 * CAP2, 
                 shape = NULL, 
                 color = NULL, 
                 label = labels)

arrowhead = arrow(length = unit(0.02, "npc"))

method.sampledf$sample <- rownames(method.sampledf)
# Plot CAP
as.data.frame(vare.cap$CCA$wa) %>% 
  rownames_to_column(var = "sample") %>% 
  as.tibble() %>% 
  left_join(method.sampledf)-> data_cap

data_cap %>% 
  ggplot(aes(x=CAP1, y=CAP2)) + geom_point(aes(col=Temp_C), size =3) + geom_segment(
    mapping = arrow_map, 
    size = .5, 
    data = arrowdf, 
    color = "orange", 
    arrow = arrowhead
  ) +
  geom_text(
    mapping = label_map, 
    size = 5,  
        color = "orange", 
    data = arrowdf, 
    show.legend = FALSE,
    position=position_jitter(width=0.08,height=0.08)
  ) + 
  geom_segment(
    mapping = arrow_map, 
    size = .5, 
    data = arrowdf_env, 
    color = "grey", 
    arrow = arrowhead
  ) +
  geom_text(
    mapping = label_map, 
    size = 5, 
    data = arrowdf_env, 
    show.legend = FALSE,
    position=position_jitter(width=0.08,height=0.08)
  )  +
  ylab(paste0("CAP2 ",round(vare.cap_prop_tib$value[2]*100,1),"%"))   + xlab(paste0("CAP1 ",round(vare.cap_prop_tib$value[1]*100,1),"%")) +   
  scale_color_gradientn(colours = cmocean('thermal')(256)) + ggpubr::theme_pubr()


```

## Benthic Species Copepod NMDS
```{r}

copepod_phy_sp = subset_taxa(physeq_obj._CO1_sum.tax, Class=="Hexanauplia")
copepod_phy_sp = subset_samples(copepod_phy_sp, !is.na(Temp_C) & !is.na(lat) & !is.na(Depth_m))
copepod_phy_sp = subset_samples(copepod_phy_sp, !Sample %in% samples_to_drop$Sample)
copepod_phy_sp = subset_samples(copepod_phy_sp, !Sample %in% c("E1720_DY2306", "E1723_DY2306", "E1948_SKQ23_12S" ,"E1949_SKQ23_12S","E1890_SKQ23_12S","E1063_SKQ2021"))
copepod_phy_sp = filter_taxa(copepod_phy_sp, function(x) sum(x > 0) > (0.01*length(x)), TRUE)
copepod_phy_sp = filter_taxa(copepod_phy_sp, function(x) sum(x) > 0, TRUE)
copepod_phy_sp <- prune_samples(sample_sums(copepod_phy_sp) > 0, copepod_phy_sp)

copepod_phy_sp <- subset_samples(copepod_phy_sp, Depth_bin == "Below 30m")
tax_table(copepod_phy_sp)
```

## Species Copepod Bray
```{r}

#Betadiversity
#Generate Vegan formatted data table
method.sampledf <- data.frame(sample_data(copepod_phy_sp))
method.rel_abun<- vegan_otu(copepod_phy_sp)


method.sampledf %>% 
  mutate(T2=as.Date(Time),
         month = month(Time)) %>% 
  mutate(., Season = case_when(month < 3 ~"Winter",
                               month < 6 ~"Spring",
                               month < 9~ "Summer",
                               month < 12 ~"Fall")) -> method.sampledf

#Jaccard dissimilarity matrix
method.d_carn <- vegdist(method.rel_abun, method="jaccard", binary=T)

#PERMANOVA: Method+Site

method.adonis_results_simple <- adonis2(method.rel_abun~Season+Depth_m+Temp_C+lat+PSU_ppt+OxyConc_umol.per.l+lon, data=method.sampledf, by = "terms",method="jaccard", binary=T)

method.adonis_results_simple

```

```{r}
ord <- ordinate(copepod_phy_sp, method = "NMDS", distance = ("jaccard"), binary=T)
```

### NMDS Temp
```{r}

sample_data(copepod_phy_sp) <- method.sampledf

##Plot_Ordination
plot_ordination(copepod_phy_sp, ord, "samples", color = "Temp_C", shape="Season") +
  ggtitle(paste0("NMDS - Stress ",round(ord$stress,3))) + 
  geom_point(size = 4) + 
  theme_bw()+ 
  theme(plot.title = element_text(hjust = 0.5,size = 20, face = "bold"),
        plot.subtitle = element_text(hjust = 0.5,size = 18, face = "bold"),
        panel.grid.major = element_blank(), 
        panel.grid.minor = element_blank(),
        panel.background = element_blank(), 
        axis.text=element_text(size=16), 
        axis.title=element_text(size=16,face="bold"),
        legend.title = element_text( size=16, face="bold"),
        legend.text = element_text(size=14)
        ) + 
  labs(color = "Temp ˚C") +scale_colour_cmocean() 


```
### CAP Analysis


## CAP Species

```{r, results="hide",warning=FALSE}
#Run CAP analysis to identify predictor species
vare.cap <- capscale(method.rel_abun~Depth_m+Temp_C+lat+PSU_ppt+OxyConc_umol.per.l+lon, data=method.sampledf, dist="bray")

#Retain species scores
sppscores(vare.cap) <- method.rel_abun

as.data.frame(vegan::scores(vare.cap, display="species")) %>% 
  rownames_to_column(var = "species") %>% 
  as.tibble %>%
 mutate(dist = sqrt((CAP1 - 0)^2 + (CAP2 - 0)^2)) -> vare.cap_species_distances


as.data.frame(vegan::scores(vare.cap, display="bp")) %>% 
  rownames_to_column(var = "env") %>% 
  as.tibble %>%
 mutate(dist = sqrt((CAP1 - 0)^2 + (CAP2 - 0)^2)) -> vare.cap_env_distances


```
##### Check For Break Point for Top Species
```{r}
species_distances <- as.data.frame(vare.cap_species_distances)
rownames(species_distances) <- species_distances$species

# Now add the environmental variables as arrows
arrowmat <- species_distances

#Rename row names to clean up plot

rownames(species_distances) %>% as.data.frame() -> namers
colnames(namers) <- c("path")
namers %>% separate(path, c("D1","P1","C1","O1","F1","G1","Species"), sep=";") %>% replace(is.na(.), "") %>% 
mutate(., name = ifelse(Species == "NA", G1, Species)) %>% 
  mutate(., name = ifelse(name == "NA", F1, name)) %>% 
  mutate(., name = ifelse(name == "NA", O1, name)) %>% 
  mutate(., name = ifelse(name == "NA", C1, name)) %>% 
  mutate(., name = ifelse(name == "NA", P1, name))-> namers

rownames(arrowmat) <- namers$name


# Add labels, make a data.frame
arrowdf <- data.frame(labels = rownames(arrowmat), arrowmat)
arrowdf %>% 
  dplyr::slice(1:10) 
arrowdf %>% 
arrange(desc(dist)) %>% 
  filter(., CAP1>0)
arrowdf %>% 
arrange(desc(dist)) %>% 
ggplot(., aes(x= reorder(labels, -dist), y=dist)) +
  geom_bar(stat="identity") + theme(axis.text.x = element_text(angle = 90, hjust = 1))

```

```{r}
arrowdf %>% 
arrange(desc(dist)) %>% 
  filter(., labels %in% c("Lucicutia flavicornis","Metridia pacifica","Pseudocalanus moultoni","Oithona similis","Pseudocalanus mimus","Calanus glacialis","Calanus pacificus","Pseudocalanus minutus"))-> arrowdf


summary(vare.cap)$cont %>% as.data.frame() %>% round(3)-> vare.cap_prop

as.tibble(vare.cap_prop,rownames = "rowname") %>% 
  pivot_longer(., cols=-rowname, values_to = "value", names_to = "name") %>% 
  filter(., rowname=="Proportion Explained") %>% 
  filter(., name %in% c("importance.CAP1","importance.CAP2"))-> vare.cap_prop_tib

# Add labels, make a data.frame
arrowdf_env <- data.frame(labels = vare.cap_env_distances$env, vare.cap_env_distances)

arrowdf_env %>% 
arrange(desc(dist)) %>% 
  filter(., env %in% c("lat", "Temp_C","Depth_m","PSU_ppt")) %>% 
  mutate(., labels = recode(labels,`lat`="Lat",
                            `Temp_C`="˚C",
                            `Depth_m` = "Depth",
                            `PSU_ppt`="PSU"))-> arrowdf_env

# Define the arrow aesthetic mapping
arrow_map <- aes(xend = 0.4 *CAP1, 
                 yend = 0.4 *CAP2, 
                 x = 0, 
                 y = 0, 
                 shape = NULL, 
                 color = NULL, 
                 label = labels)

label_map <- aes(x = 0.32 * CAP1, 
                 y = 0.32 * CAP2, 
                 shape = NULL, 
                 color = NULL, 
                 label = labels,
                 alpha=0.7,
                 fontface="italic")
label_map2 <- aes(x = 0.32 * CAP1, 
                 y = 0.32 * CAP2, 
                 shape = NULL, 
                 color = NULL, 
                 alpha=0.7,
                 label = labels)

arrowhead = arrow(length = unit(0.02, "npc"))

method.sampledf$sample <- rownames(method.sampledf)
# Plot CAP
as.data.frame(vare.cap$CCA$wa) %>% 
  rownames_to_column(var = "sample") %>% 
  as.tibble() %>% 
  left_join(method.sampledf)-> data_cap


data_cap %>% 
  ggplot(aes(x=CAP1, y=CAP2)) + geom_point(aes(col=Temp_C, shape=Season), size =3) + geom_segment(
    mapping = arrow_map, 
    size = .5, 
    data = arrowdf, 
    color = "dodgerblue3", 
    arrow = arrowhead
  ) +
  geom_label(
    mapping = label_map, 
    size = 5,  
        color = "dodgerblue3", 
    data = arrowdf, 
    show.legend = FALSE,
    position=position_jitter(width=0.03,height=0.03)
  ) + 
  geom_segment(
    mapping = arrow_map, 
    size = .5, 
    data = arrowdf_env, 
    color = "grey", 
    arrow = arrowhead
  ) +
  geom_label(
    mapping = label_map2, 
    size = 5, 
    data = arrowdf_env, 
    show.legend = FALSE,
    position=position_jitter(width=0.03,height=0.03)
  )  +
  ylab(paste0("CAP2 [",round(vare.cap_prop_tib$value[2]*100,1),"]%"))   + xlab(paste0("CAP1 [",round(vare.cap_prop_tib$value[1]*100,1),"]%")) +   
  scale_color_gradientn(colours = cmocean('thermal')(256), limits=c(-2,8)) + ggpubr::theme_pubr() +labs(colour="Temperature ˚C") + theme(axis.title = element_text(face="bold", size = 16))


ggsave(filename = here("gold_out","CAP_benthic_copepods.png"),
       width=12,
       height = 8,
       dpi = 400,
      units = c("in"))

```


```{r}
ord <- ordinate(subsurface_co1_2020, method = "NMDS", distance = ("bray"))
```


```{r}
##Plot_Ordination
plot_ordination(subsurface_co1_2020, ord, "samples", color = "Temp_C") +
  ggtitle(paste0("NMDS - Stress ",round(ord$stress,3))) + 
  geom_point(size = 4) + 
  theme_bw()+ 
  theme(plot.title = element_text(hjust = 0.5,size = 20, face = "bold"),
        plot.subtitle = element_text(hjust = 0.5,size = 18, face = "bold"),
        panel.grid.major = element_blank(), 
        panel.grid.minor = element_blank(),
        panel.background = element_blank(), 
        axis.text=element_text(size=16), 
        axis.title=element_text(size=16,face="bold"),
        legend.title = element_text( size=16, face="bold"),
        legend.text = element_text(size=14)
        ) + 
  labs(color = "Temp ˚C") +scale_colour_cmocean() 


```

# Taxa Plot
## Taxa Exploration


```{r}
#Chionoecetes opilio, Euphausiacea, Limacina helicina, Pteropoda

long_COI_EcoFOCI %>% 
    filter(, !is.na(Depth_m)) %>% 
  ungroup() %>% 
   unite(sum.taxonomy, c("Kingdom",              "Phylum",               "Class"               
,"Order",                "Family",               "Genus",               "Species"), sep=";", remove=F) %>%  
  dplyr::select(sum.taxonomy, Kingdom,Phylum,Class, Order,Family, Genus, Species) %>% 
  distinct() -> tax_df


tax_df %>% 
  filter(., str_detect(sum.taxonomy,"Echinodermata"))

tax_df %>% 
  filter(., str_detect(sum.taxonomy,"Mollusca"))

tax_df %>% 
  filter(., str_detect(sum.taxonomy,"Decapoda"))

tax_df %>% 
    filter(., Kingdom== "Eukaryota") %>% 
dplyr::select(Species) %>% distinct()
```

### CO1
```{r}

tax_df %>% 
  filter(., Kingdom== "Eukaryota") %>% 
  filter(., Species !="NA") %>% 
  dplyr::select(Phylum) %>%  distinct()

tax_df %>% 
  filter(., Kingdom== "Eukaryota") %>% 
  filter(., Species !="NA") %>% 
  group_by(Phylum) %>% 
  dplyr::summarise(Count = n_distinct(Species)) %>% 
  arrange(desc(Count)) %>% 
  dplyr::slice(1:10) %>% 
  ggplot(aes(x=reorder(Phylum,-Count), y=Count)) + geom_col(fill="dodgerblue4") + theme_pubr() + theme(axis.title = element_text(face="bold", size=16),
      axis.text.x = element_text(angle = 30, vjust=1, hjust = 1)) +xlab("Phylum") +ggtitle(label = "Leray CO1") -> co1_phyla_plot
co1_phyla_plot
```

```{r}
#Chionoecetes opilio, Euphausiacea, Limacina helicina, Pteropoda

long_Machida_EcoFOCI %>% 
    filter(, !is.na(Depth_m)) %>% 
  ungroup() %>% 
   unite(sum.taxonomy, c("Kingdom",              "Phylum",               "Class"               
,"Order",                "Family",               "Genus",               "Species"), sep=";", remove=F) %>%  
  dplyr::select(sum.taxonomy, Kingdom,Phylum,Class, Order,Family, Genus, Species) %>% 
  distinct() -> tax_df

tax_df %>% 
  filter(., str_detect(sum.taxonomy,"Echinodermata"))

tax_df %>% 
  filter(., str_detect(sum.taxonomy,"Mollusca"))


tax_df %>% 
  filter(., str_detect(sum.taxonomy,"Arthropoda"))


```

### Machida
```{r}

tax_df %>% 
  filter(., Kingdom== "Eukaryota") %>% 
  filter(., Species !="NA") %>% 
  dplyr::select(Species) %>%  distinct()


tax_df %>% 
  filter(., Kingdom== "Eukaryota") %>% 
  filter(., Species !="NA") %>% 
  group_by(Phylum) %>% 
  dplyr::summarise(Count = n_distinct(Species)) %>% 
  arrange(desc(Count)) %>% 
  dplyr::slice(1:10) %>% 
  ggplot(aes(x=reorder(Phylum,-Count), y=Count)) + geom_col(fill="purple4") + theme_pubr() + theme(axis.title = element_text(face="bold", size=16),
      axis.text.x = element_text(angle = 30, vjust=1, hjust = 1)) +xlab("Phylum") +ggtitle(label = "Machida 18Sv8-9") -> machida_phyla_plot
machida_phyla_plot
```

### MiFish
```{r}


long_MiFish_EcoFOCI %>% 
    filter(, !is.na(Depth_m)) %>% 
ungroup() %>% 
   unite(sum.taxonomy, c("Kingdom",              "Phylum",               "Class"               
,"Order",                "Family",               "Genus",               "Species"), sep=";", remove=F) %>%  
  dplyr::select(sum.taxonomy, Kingdom,Phylum,Class, Order,Family, Genus, Species) %>% 
  distinct() -> tax_df_mifish


tax_df_mifish %>% 
    filter(., Phylum== "Chordata") %>% 
  dplyr::select(Species) %>%  distinct()
  
tax_df_mifish %>% 
    filter(., Phylum== "Chordata") %>% 
  filter(., str_detect(sum.taxonomy,"Chondrichthyes"))


tax_df_mifish %>% 
  filter(., str_detect(sum.taxonomy,"Mammalia"))

```

```{r}
tax_df_mifish %>% 
  filter(., Phylum== "Chordata") %>% 
    filter(., Species !="NA") %>% 
  group_by(Class) %>% 
  dplyr::summarise(Count = n_distinct(Species)) %>% 
  arrange(desc(Count)) %>% 
  dplyr::slice(1:10) %>% 
  ggplot(aes(x=reorder(Class,-Count), y=Count)) + geom_col(fill="aquamarine4") + theme_pubr() + theme(axis.title = element_text(face="bold", size=16),
     axis.text.x = element_text(angle = 30, vjust=1, hjust = 1)) +xlab("Class") +ggtitle(label = "MiFish") -> mifish_class_plot
```

```{r}
mifish_class_plot + co1_phyla_plot  + plot_annotation(tag_levels = 'A')

ggsave(filename = here("gold_out","Total_species.png"),
       width=16,
       height = 6,
       dpi = 400,
      units = c("in"))

```


## Calanus
```{r}

long_COI_EcoFOCI %>% 
  dplyr::select(ASV_combo, Kingdom, Phylum, Class, Order, Family, Genus, Species) %>% 
  distinct() %>% 
  filter(., Genus=="Calanus")
```


```{r}
long_COI_EcoFOCI %>% 
  filter(., Genus=="Calanus") %>% 
  dplyr::group_by(ASV_combo, Kingdom, Phylum, Class, Order, Family, Genus, Species) %>% 
  dplyr::summarise(sum_reads=sum(nReads)) %>% 
  arrange(desc(sum_reads))


```

# Machida

#### Build Phyloseq Object

##### ASV Reads

```{r}

varnames <- colnames(long_Machida_EcoFOCI)
to_remove <- c("Run2:ASV","Run3:ASV","ASV_combo","Sequence","nReads","Prop.abund","eDNA.Index")


sample_metadata -> Machida_sample_data

#Metadata
Machida_sample_data %>%  as.data.frame() -> sampledata
rownames(sampledata) <- sampledata$Sample
sample_data(sampledata) -> sampledata

#ASV Reads

long_Machida_EcoFOCI %>% 
  dplyr::select(ASV_combo,Sample, eDNA.Index) %>% 
  mutate(., eDNA.Index=as.numeric(eDNA.Index)) %>% 
  mutate(., eDNA.Index=if_else(eDNA.Index=="NaN", 0, eDNA.Index)) %>%
  pivot_wider(names_from = Sample, values_from = eDNA.Index, values_fill  =0) -> wide_PA

dim(wide_PA)

long_Machida_EcoFOCI %>% 
  dplyr::select(ASV_combo, Kingdom,Phylum,Class, Order,Family, Genus, Species) %>% 
  distinct() %>% as.matrix() -> taxonomy_table

rownames(taxonomy_table) <- wide_PA$ASV_combo

TAX = tax_table(taxonomy_table)

wide_PA %>% 
  ungroup() %>% 
  dplyr::select(-ASV_combo) %>% as.matrix() -> otu_table
rownames(otu_table) <- wide_PA$ASV_combo

OTU = otu_table(otu_table, taxa_are_rows = TRUE)
physeq_obj._Machida = phyloseq(OTU, TAX, sampledata)

physeq_obj._Machida_2020 <- subset_samples(physeq_obj._Machida,Cruise_ID_short != "SKQ2021")

subsurface_Machida <- subset_samples(physeq_obj._Machida, Depth_bin == "Below 30m")
subsurface_Machida_2020 <- subset_samples(subsurface_Machida, Cruise_ID_short != "SKQ2021")

surface_Machida <- subset_samples(physeq_obj._Machida, Depth_bin == "Above 30m")
surface_Machida_2020 <- subset_samples(surface_Machida, Cruise_ID_short != "SKQ2021")

```

#### Build Phyloseq Object

```{r}

varnames <- colnames(long_Machida_EcoFOCI)
to_remove <- c("Run2:ASV","Run3:ASV","ASV_combo","Sequence","nReads","Prop.abund","eDNA.Index")


sample_metadata -> Machida_sample_data

#Metadata
Machida_sample_data %>%  as.data.frame() -> sampledata
rownames(sampledata) <- sampledata$Sample
sample_data(sampledata) -> sampledata

#ASV Reads

c(setdiff(varnames,to_remove),"Tot") -> varnames_to_group

long_Machida_EcoFOCI %>% 
  dplyr::select(-eDNA.Index, -Prop.abund) %>% 
  ungroup() %>% 
 group_by(Sample) %>%
    mutate(Tot = sum(nReads)) %>% 
    group_by(!!!syms(varnames_to_group)) %>% 
    dplyr::summarise(nReads=sum(nReads)) %>% 
    mutate (Prop.abund = nReads / Tot) %>% 
      ungroup() %>% 
    unite(sum.taxonomy, c("Kingdom",              "Phylum",               "Class"               
,"Order",                "Family",               "Genus",               "Species"), sep=";") %>% 
    group_by (sum.taxonomy) %>%
    mutate (Colmax = max(Prop.abund),
            eDNA.Index = Prop.abund / Colmax) %>% 
  dplyr::select(sum.taxonomy,Sample, eDNA.Index) %>% 
  mutate(., eDNA.Index=as.numeric(eDNA.Index)) %>% 
  mutate(., eDNA.Index=if_else(eDNA.Index=="NaN", 0, eDNA.Index)) %>%
  pivot_wider(names_from = Sample, values_from = eDNA.Index, values_fill  =0) -> wide_PA

wide_PA %>% 
  dplyr::select(sum.taxonomy) %>% 
   separate(sum.taxonomy, into=c("Kingdom",              "Phylum",               "Class"               
,"Order",                "Family",               "Genus",               "Species"), sep=";", remove=F) %>%  
  dplyr::select(sum.taxonomy, Kingdom,Phylum,Class, Order,Family, Genus, Species) %>% 
  distinct() %>% as.matrix() -> taxonomy_table
dim(taxonomy_table)
dim(wide_PA)
rownames(taxonomy_table) <- wide_PA$sum.taxonomy

TAX = tax_table(taxonomy_table)

wide_PA %>% 
  ungroup() %>% 
  dplyr::select(-sum.taxonomy) %>% as.matrix() -> otu_table
rownames(otu_table) <- wide_PA$sum.taxonomy

OTU = otu_table(otu_table, taxa_are_rows = TRUE)
physeq_obj._Machida_sum.tax = phyloseq(OTU, TAX, sampledata)

physeq_obj._Machida_2020_sum.tax <- subset_samples(physeq_obj._Machida_sum.tax,Cruise_ID_short != "SKQ2021")

subsurface_Machida_sum.tax <- subset_samples(physeq_obj._Machida_sum.tax, Depth_bin == "Below 30m")
subsurface_Machida_2020_sum.tax <- subset_samples(subsurface_Machida_sum.tax, Cruise_ID_short != "SKQ2021")

surface_Machida_sum.tax <- subset_samples(physeq_obj._Machida_sum.tax, Depth_bin == "Above 30m")
surface_Machida_2020_sum.tax <- subset_samples(surface_Machida_sum.tax, Cruise_ID_short != "SKQ2021")

```


## Benthic Species Copepod NMDS
```{r}
copepod_phy_sp_m = subset_taxa(physeq_obj._Machida_sum.tax, Class=="Hexanauplia")
copepod_phy_sp_m <- prune_samples(sample_sums(copepod_phy_sp_m) > 0, copepod_phy_sp_m)
copepod_phy_sp_m = filter_taxa(copepod_phy_sp_m, function(x) sum(x > 0) > (0.1*length(x)), TRUE)
copepod_phy_sp_m = subset_samples(copepod_phy_sp_m, !is.na(Temp_C) & !is.na(lat) & !is.na(Depth_m))
copepod_phy_sp_m <- subset_samples(copepod_phy_sp_m, Cruise_ID_short != "SKQ2021")
copepod_phy_sp_m <- subset_samples(copepod_phy_sp_m, Depth_bin == "Below 30m")

tax_table(copepod_phy_sp_m) %>%  as.data.frame()
```

## Species Copepod Bray
```{r}

#Betadiversity
#Generate Vegan formatted data table
method.sampledf <- data.frame(sample_data(copepod_phy_sp_m))
method.rel_abun<- vegan_otu(copepod_phy_sp_m)

#Jaccard dissimilarity matrix
method.d_carn <- vegdist(method.rel_abun, method="bray")

#PERMANOVA: Method+Site

method.adonis_results_simple <- adonis2(method.rel_abun~Depth_m+Temp_C+lat+PSU_ppt+OxyConc_umol.per.l+lon, data=method.sampledf, by = "terms")

method.adonis_results_simple

```

```{r}
ord <- ordinate(copepod_phy_sp_m, method = "NMDS", distance = ("bray"))
```

### NMDS Temp
```{r}
##Plot_Ordination
plot_ordination(copepod_phy_sp_m, ord, "samples", color = "Temp_C", shape="Depth_bin") +
  ggtitle(paste0("NMDS - Stress ",round(ord$stress,3))) + 
  geom_point(size = 4) + 
  theme_bw()+ 
  theme(plot.title = element_text(hjust = 0.5,size = 20, face = "bold"),
        plot.subtitle = element_text(hjust = 0.5,size = 18, face = "bold"),
        panel.grid.major = element_blank(), 
        panel.grid.minor = element_blank(),
        panel.background = element_blank(), 
        axis.text=element_text(size=16), 
        axis.title=element_text(size=16,face="bold"),
        legend.title = element_text( size=16, face="bold"),
        legend.text = element_text(size=14)
        ) + 
  labs(color = "Temp ˚C") +scale_colour_cmocean() 


```
### CAP Analysis


## CAP Species

```{r, results="hide",warning=FALSE}
#Run CAP analysis to identify predictor species
vare.cap <- capscale(method.rel_abun~Depth_m+Temp_C+lat+PSU_ppt+OxyConc_umol.per.l+lon, data=method.sampledf, dist="bray")

#Retain species scores
sppscores(vare.cap) <- method.rel_abun

as.data.frame(vegan::scores(vare.cap, display="species")) %>% 
  rownames_to_column(var = "species") %>% 
  as.tibble %>%
 mutate(dist = sqrt((CAP1 - 0)^2 + (CAP2 - 0)^2)) -> vare.cap_species_distances


as.data.frame(vegan::scores(vare.cap, display="bp")) %>% 
  rownames_to_column(var = "env") %>% 
  as.tibble %>%
 mutate(dist = sqrt((CAP1 - 0)^2 + (CAP2 - 0)^2)) -> vare.cap_env_distances


```
##### Check For Break Point for Top Species
```{r}
species_distances <- as.data.frame(vare.cap_species_distances)
rownames(species_distances) <- species_distances$species

# Now add the environmental variables as arrows
arrowmat <- species_distances

#Rename row names to clean up plot

rownames(species_distances) %>% as.data.frame() -> namers
colnames(namers) <- c("path")
namers %>% separate(path, c("D1","P1","C1","O1","F1","G1","Species"), sep=";") %>% replace(is.na(.), "") %>% 
mutate(., name = ifelse(Species == "NA", G1, Species)) %>% 
  mutate(., name = ifelse(name == "NA", F1, name)) %>% 
  mutate(., name = ifelse(name == "NA", O1, name)) %>% 
  mutate(., name = ifelse(name == "NA", C1, name)) %>% 
  mutate(., name = ifelse(name == "NA", P1, name))-> namers

rownames(arrowmat) <- namers$name


# Add labels, make a data.frame
arrowdf <- data.frame(labels = rownames(arrowmat), arrowmat)

arrowdf %>% 
arrange(desc(dist)) %>% 
ggplot(., aes(x= reorder(labels, -dist), y=dist)) +
  geom_bar(stat="identity") + theme(axis.text.x = element_text(angle = 90, hjust = 1))

```

```{r}

arrowdf %>% 
arrange(desc(dist)) %>% 
  dplyr::slice(1:5) -> arrowdf


summary(vare.cap)$cont %>% as.data.frame() %>% round(3)-> vare.cap_prop

as.tibble(vare.cap_prop,rownames = "rowname") %>% 
  pivot_longer(., cols=-rowname, values_to = "value", names_to = "name") %>% 
  filter(., rowname=="Proportion Explained") %>% 
  filter(., name %in% c("importance.CAP1","importance.CAP2"))-> vare.cap_prop_tib

# Add labels, make a data.frame
arrowdf_env <- data.frame(labels = vare.cap_env_distances$env, vare.cap_env_distances)

arrowdf_env %>% 
arrange(desc(dist)) %>% 
  filter(., env %in% c("lat", "Temp_C","Depth_m","PSU_ppt")) %>% 
  mutate(., labels = recode(labels,`lat`="Lat",
                            `Temp_C`="˚C",
                            `Depth_m` = "Depth",
                            `PSU_ppt`="PSU"))-> arrowdf_env

# Define the arrow aesthetic mapping
arrow_map <- aes(xend = 0.4 *CAP1, 
                 yend = 0.4 *CAP2, 
                 x = 0, 
                 y = 0, 
                 shape = NULL, 
                 color = NULL, 
                 label = labels)

label_map <- aes(x = 0.32 * CAP1, 
                 y = 0.32 * CAP2, 
                 shape = NULL, 
                 color = NULL, 
                 label = labels,
                 fontface="italic")
label_map2 <- aes(x = 0.32 * CAP1, 
                 y = 0.32 * CAP2, 
                 shape = NULL, 
                 color = NULL, 
                 label = labels)

arrowhead = arrow(length = unit(0.02, "npc"))

method.sampledf$sample <- rownames(method.sampledf)
# Plot CAP
as.data.frame(vare.cap$CCA$wa) %>% 
  rownames_to_column(var = "sample") %>% 
  as.tibble() %>% 
  left_join(method.sampledf)-> data_cap


data_cap %>% 
  ggplot(aes(x=CAP1, y=CAP2)) + geom_point(aes(col=Temp_C), size =3) + geom_segment(
    mapping = arrow_map, 
    size = .5, 
    data = arrowdf, 
    color = "dodgerblue3", 
    arrow = arrowhead
  ) +
  geom_label(
    mapping = label_map, 
    size = 5,  
        color = "dodgerblue3", 
    data = arrowdf, 
    show.legend = FALSE,
    position=position_jitter(width=0.03,height=0.03)
  ) + 
  geom_segment(
    mapping = arrow_map, 
    size = .5, 
    data = arrowdf_env, 
    color = "grey", 
    arrow = arrowhead
  ) +
  geom_label(
    mapping = label_map2, 
    size = 5, 
    data = arrowdf_env, 
    show.legend = FALSE,
    position=position_jitter(width=0.03,height=0.03)
  )  +
  ylab(paste0("CAP2 [",round(vare.cap_prop_tib$value[2]*100,1),"]%"))   + xlab(paste0("CAP1 [",round(vare.cap_prop_tib$value[1]*100,1),"]%")) +   
  scale_color_gradientn(colours = cmocean('thermal')(256), limits=c(-2,8)) + ggpubr::theme_pubr() +labs(colour="Temperature ˚C") + theme(axis.title = element_text(face="bold", size = 16))


ggsave(filename = here("gold_out","CAP_benthic_copepods.png"),
       width=12,
       height = 8,
       dpi = 400,
      units = c("in"))

```

# MiFish
## Group by Species

#### Build Phyloseq Object

```{r}

varnames <- colnames(long_MiFish_EcoFOCI)
to_remove <- c("Run2:ASV","Run3:ASV","Run4:ASV","ASV_combo","Sequence","nReads","Prop.abund","eDNA.Index")


sample_metadata -> co1_sample_data

#Metadata
co1_sample_data %>%  as.data.frame() -> sampledata
rownames(sampledata) <- sampledata$Sample
sample_data(sampledata) -> sampledata

#ASV Reads

c(setdiff(varnames,to_remove),"Tot") -> varnames_to_group

long_MiFish_EcoFOCI %>% 
  dplyr::select(-eDNA.Index, -Prop.abund) %>% 
  ungroup() %>% 
 group_by(Sample) %>%
    mutate(Tot = sum(nReads)) %>% 
    group_by(!!!syms(varnames_to_group)) %>% 
    dplyr::summarise(nReads=sum(nReads)) %>% 
    mutate (Prop.abund = nReads / Tot) %>% 
      ungroup() %>% 
    unite(sum.taxonomy, c("Kingdom",              "Phylum",               "Class"               
,"Order",                "Family",               "Genus",               "Species"), sep=";") %>% 
    group_by (sum.taxonomy) %>%
    mutate (Colmax = max(Prop.abund),
            eDNA.Index = Prop.abund / Colmax) %>% 
  dplyr::select(sum.taxonomy,Sample, eDNA.Index) %>% 
  mutate(., eDNA.Index=as.numeric(eDNA.Index)) %>% 
  mutate(., eDNA.Index=if_else(eDNA.Index=="NaN", 0, eDNA.Index)) %>%
  pivot_wider(names_from = Sample, values_from = eDNA.Index, values_fill  =0) -> wide_PA

wide_PA %>% 
  dplyr::select(sum.taxonomy) %>% 
   separate(sum.taxonomy, into=c("Kingdom",              "Phylum",               "Class"               
,"Order",                "Family",               "Genus",               "Species"), sep=";", remove=F) %>%  
  dplyr::select(sum.taxonomy, Kingdom,Phylum,Class, Order,Family, Genus, Species) %>% 
  distinct() %>% as.matrix() -> taxonomy_table
dim(taxonomy_table)
dim(wide_PA)
rownames(taxonomy_table) <- wide_PA$sum.taxonomy

TAX = tax_table(taxonomy_table)

wide_PA %>% 
  ungroup() %>% 
  dplyr::select(-sum.taxonomy) %>% as.matrix() -> otu_table
rownames(otu_table) <- wide_PA$sum.taxonomy

OTU = otu_table(otu_table, taxa_are_rows = TRUE)
physeq_obj_MiFish_sum.tax = phyloseq(OTU, TAX, sampledata)

subsurface_MiFish_sum.tax <- subset_samples(physeq_obj_MiFish_sum.tax, Depth_bin == "Below 30m")

surface_MiFish_sum.tax <- subset_samples(physeq_obj_MiFish_sum.tax, Depth_bin == "Above 30m")


surface_MiFish_sum.tax <- prune_samples(sample_sums(surface_MiFish_sum.tax) > 0, surface_MiFish_sum.tax)
surface_MiFish_sum.tax_sp = subset_samples(surface_MiFish_sum.tax, !is.na(Temp_C) & !is.na(lat) & !is.na(Depth_m))
surface_MiFish_sum.tax_sp = subset_taxa(surface_MiFish_sum.tax_sp, Kingdom!="Bacteria")
surface_MiFish_sum.tax_sp = subset_taxa(surface_MiFish_sum.tax_sp, Kingdom!="Environmental Unknown")


subsurface_MiFish_sum.tax <- prune_samples(sample_sums(subsurface_MiFish_sum.tax) > 0, subsurface_MiFish_sum.tax)
subsurface_MiFish_sum.tax_sp = subset_samples(subsurface_MiFish_sum.tax, !is.na(Temp_C) & !is.na(lat) & !is.na(Depth_m))
subsurface_MiFish_sum.tax_sp = subset_taxa(subsurface_MiFish_sum.tax_sp, Kingdom!="Bacteria")
subsurface_MiFish_sum.tax_sp = subset_taxa(subsurface_MiFish_sum.tax_sp, Kingdom!="Environmental Unknown")
```


# Power Analysis

## CO1 
### iNEXT
```{r}

subsurface_co1_sum.tax <- subset_samples(subsurface_co1_sum.tax, Cruise_ID_short != "SKQ2021")

subsurface_co1_sum.tax <- prune_samples(sample_sums(subsurface_co1_sum.tax) > 0, subsurface_co1_sum.tax)

subsurface_co1_sum.tax_sp = subset_samples(subsurface_co1_sum.tax, !is.na(Temp_C) & !is.na(lat) & !is.na(Depth_m))
subsurface_co1_sum.tax_sp = subset_taxa(subsurface_co1_sum.tax_sp, Kingdom!="Bacteria")
subsurface_co1_sum.tax_sp = subset_taxa(subsurface_co1_sum.tax_sp, Kingdom!="Environmental Unknown")


#isolate groups

#Convert groups into vegan outputs
veganComm_surface_CO1 <- vegan_otu(surface_co1_sum.tax_sp)
veganComm_benthic_CO1 <- vegan_otu(subsurface_co1_sum.tax_sp)

###### iNEXT Species Accumulation Curves

#Create List for iNEXT Species Incidence Frequencies
mor_inc <-list ("Above 30m" =t(veganComm_surface_CO1),
                "Below 30m"=t(veganComm_benthic_CO1))

species_incidence <-lapply(mor_inc, as.incfreq)

#Convert to iNEXT format
t <- seq(1, 300, by=1)
out.inc2 <- iNEXT(species_incidence, q=0, datatype="incidence_freq", size=t)

#out.inc2$DataInfo
```

```{r}
#Sample‐size‐based R/E curves
plot_co1 <- ggiNEXT(out.inc2, type=1, color.var="Both") + 
  theme_bw(base_size = 18) + theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
                                                                                                    panel.background = element_blank()) + xlab("Stations") + ylab("Taxa") + scale_colour_manual(values = c("dodgerblue4","#Ed820e"), name = "Group", labels = c("Above 30m", "Below 30m")) + scale_shape_discrete(name = "Group", labels = c("Above 30m", "Below 30m")) + scale_fill_manual(values = c("dodgerblue4","#Ed820e"),name = "Group", labels = c("Above 30m", "Below 30m")) +ggtitle(label = "Leray CO1")+ggtitle(label = "", subtitle="Leray CO1") +theme(axis.line = element_line(colour = "black"),panel.grid.major = element_blank(), panel.grid.minor = element_blank(),panel.background = element_blank(), axis.text=element_text(size=16),
        axis.title=element_text(size=20,face="bold"),legend.title = element_text( size=12, face="bold"),legend.text = element_text(size=10, 
                                     face="bold"), title = element_text(size=20, face="bold") ) + theme(legend.text=element_text(size=16), legend.title=element_text(size=20), axis.title.x=element_blank())

plot_co1


```

## MiFsih
```{r}
#isolate groups

#Convert groups into vegan outputs
veganComm_surface_MiFish <- vegan_otu(surface_MiFish_sum.tax_sp)
veganComm_benthic_MiFish <- vegan_otu(subsurface_MiFish_sum.tax_sp)

###### iNEXT Species Accumulation Curves

#Create List for iNEXT Species Incidence Frequencies
mor_inc <-list ("Above 30m" =t(veganComm_surface_MiFish),
                "Below 30m"=t(veganComm_benthic_MiFish))

species_incidence <-lapply(mor_inc, as.incfreq)

#Convert to iNEXT format
t <- seq(1, 350, by=1)
out.inc_fish <- iNEXT(species_incidence, q=0, datatype="incidence_freq", size=t)

#out.inc2$DataInfo
```

```{r}
#Sample‐size‐based R/E curves
plot_fish <- ggiNEXT(out.inc_fish, type=1, color.var="Both") + 
  theme_bw(base_size = 18) + theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
                                                                                                    panel.background = element_blank()) + xlab("Stations") + ylab("Taxa") + scale_colour_manual(values = c("dodgerblue4","#Ed820e"), name = "Group", labels = c("Above 30m", "Below 30m")) + scale_shape_discrete(name = "Group", labels = c("Above 30m", "Below 30m")) + scale_fill_manual(values = c("dodgerblue4","#Ed820e"),name = "Group", labels = c("Above 30m", "Below 30m")) +ggtitle(label = "Accumulation", subtitle="MiFish") + ylim(0,600) +theme(axis.line = element_line(colour = "black"),panel.grid.major = element_blank(), panel.grid.minor = element_blank(),panel.background = element_blank(), axis.text=element_text(size=16),
        axis.title=element_text(size=20,face="bold"),legend.title = element_text( size=12, face="bold"),legend.text = element_text(size=10, 
                                     face="bold"), title = element_text(size=20, face="bold") ) + theme(legend.text=element_text(size=16), legend.title=element_text(size=20), axis.title.x=element_blank())
```

## Machida 
### iNEXT
```{r}

subsurface_Machida_2020_sum.tax <- subset_samples(subsurface_Machida_sum.tax, Cruise_ID_short != "SKQ2021")

subsurface_Machida_2020_sum.tax <- prune_samples(sample_sums(subsurface_Machida_2020_sum.tax) > 0, subsurface_Machida_2020_sum.tax)

subsurface_Machida_2020_sum.tax_sp = subset_samples(subsurface_Machida_2020_sum.tax, !is.na(Temp_C) & !is.na(lat) & !is.na(Depth_m))
subsurface_Machida_2020_sum.tax_sp = subset_taxa(subsurface_Machida_2020_sum.tax_sp, Kingdom!="Bacteria")
subsurface_Machida_2020_sum.tax_sp = subset_taxa(subsurface_Machida_2020_sum.tax_sp, Kingdom!="Environmental Unknown")


surface_Machida_sum.tax <- subset_samples(surface_Machida_sum.tax, Cruise_ID_short != "SKQ2021")

surface_Machida_sum.tax <- prune_samples(sample_sums(surface_Machida_sum.tax) > 0, surface_Machida_sum.tax)

surface_Machida_2020_sum.tax_sp = subset_samples(surface_Machida_sum.tax, !is.na(Temp_C) & !is.na(lat) & !is.na(Depth_m))
surface_Machida_2020_sum.tax_sp = subset_taxa(surface_Machida_2020_sum.tax_sp, Kingdom!="Bacteria")
surface_Machida_2020_sum.tax_sp = subset_taxa(surface_Machida_2020_sum.tax_sp, Kingdom!="Environmental Unknown")


#isolate groups

#Convert groups into vegan outputs
veganComm_surface_Machida <- vegan_otu(surface_Machida_2020_sum.tax_sp)
veganComm_benthic_Machida <- vegan_otu(subsurface_Machida_2020_sum.tax_sp)

###### iNEXT Species Accumulation Curves

#Create List for iNEXT Species Incidence Frequencies
mor_inc <-list ("Above 30m" =t(veganComm_surface_Machida),
                "Below 30m"=t(veganComm_benthic_Machida))

species_incidence <-lapply(mor_inc, as.incfreq)

#Convert to iNEXT format
t <- seq(1, 200, by=1)
out.inc_machida <- iNEXT(species_incidence, q=0, datatype="incidence_freq", size=t)

#out.inc2$DataInfo
```

```{r}
#Sample‐size‐based R/E curves
plot_machida <- ggiNEXT(out.inc_machida, type=1, color.var="Both") + 
  theme_bw(base_size = 18) + theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
                                                                                                    panel.background = element_blank()) + xlab("Stations") + ylab("Taxa") + scale_colour_manual(values = c("dodgerblue4","#Ed820e"), name = "Group", labels = c("Above 30m", "Below 30m")) + scale_shape_discrete(name = "Group", labels = c("Above 30m", "Below 30m")) + scale_fill_manual(values = c("dodgerblue4","#Ed820e"),name = "Group", labels = c("Above 30m", "Below 30m")) +ggtitle(label = "Machida 18Sv8-9") +ggtitle(label = "", subtitle="Machida 18Sv8-9") +theme(axis.line = element_line(colour = "black"),panel.grid.major = element_blank(), panel.grid.minor = element_blank(),panel.background = element_blank(), axis.text=element_text(size=16),
        axis.title=element_text(size=20,face="bold"),legend.title = element_text( size=12, face="bold"),legend.text = element_text(size=10, 
                                     face="bold"), title = element_text(size=20, face="bold") ) + theme(legend.text=element_text(size=16), legend.title=element_text(size=20))

plot_machida


```


```{r}
plot_fish +plot_co1+ plot_machida + plot_annotation(tag_levels = 'A') +
  plot_layout(guides = 'collect')


ggsave(filename = here("gold_out","Species_rarefaction.png"),
       width=12,
       height = 8,
       dpi = 400,
      units = c("in"))
```

```{r}

surface_north_2020 = subset_samples(surface_co1_2020_sum.tax_sp, lat > 65.89)
surface_north_2020 <- prune_taxa(taxa_sums(surface_north_2020) > 0, surface_north_2020)

surface_south_2020 = subset_samples(surface_co1_2020_sum.tax_sp, lat < 65.89)
surface_south_2020 <- prune_taxa(taxa_sums(surface_south_2020) > 0, surface_south_2020)

subsurface_south_2020 = subset_samples(subsurface_co1_2020_sum.tax_sp, lat < 65.89)
subsurface_south_2020 <- prune_taxa(taxa_sums(subsurface_south_2020) > 0, subsurface_south_2020)

subsurface_north_2020 = subset_samples(subsurface_co1_2020_sum.tax_sp, lat > 65.89)
subsurface_north_2020 <- prune_taxa(taxa_sums(subsurface_north_2020) > 0, subsurface_north_2020)

#isolate groups

#Convert groups into vegan outputs
veganComm_surface_CO1_north <- vegan_otu(surface_north_2020)
veganComm_surface_CO1_south <- vegan_otu(surface_south_2020)

veganComm_benthic_CO1_south <- vegan_otu(subsurface_south_2020)
veganComm_benthic_CO1_north <- vegan_otu(subsurface_north_2020)

###### iNEXT Species Accumulation Curves

#Create List for iNEXT Species Incidence Frequencies
mor_inc <-list ("Above 30m, North of Bering Strait" =t(veganComm_surface_CO1_north),
                "Above 30m, South of Bering Strait" =t(veganComm_surface_CO1_south),
                "Below 30m, North of Bering Strait"=t(veganComm_benthic_CO1_north),
                "Below 30m, South of Bering Strait"=t(veganComm_benthic_CO1_south))

species_incidence <-lapply(mor_inc, as.incfreq)

#Convert to iNEXT format
t <- seq(1, 100, by=1)
out.inc_merp <- iNEXT(species_incidence, q=0, datatype="incidence_freq", size=t)

#out.inc2$DataInfo

```

```{r}
#Sample‐size‐based R/E curves
plot_north_south <- ggiNEXT(out.inc_merp, type=1, color.var="Both") + 
  theme_bw(base_size = 18) + theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
                                                                                                    panel.background = element_blank()) + xlab("Stations") + ylab("Taxa") + scale_colour_manual(values = c("dodgerblue4","dodgerblue2","#Ed820e","#Ed8"), name = "Group", labels = c("Above 30m, North of Bering Strait", "Above 30m, South of Bering Strait","Below 30m, North of Bering Strait","Below 30m, South of Bering Strait")) + scale_shape_discrete(name = "Group", labels = c("Above 30m, North of Bering Strait", "Above 30m, South of Bering Strait","Below 30m, North of Bering Strait","Below 30m, South of Bering Strait")) + scale_fill_manual(values = c("dodgerblue4","dodgerblue2","#Ed820e","#Ed8"),name = "Group", labels = c("Above 30m, North of Bering Strait", "Above 30m, South of Bering Strait","Below 30m, North of Bering Strait","Below 30m, South of Bering Strait"))

plot_north_south
```

# Zeta Diversity

```{r}
library(zetadiv)
```

### surface_co1_2020_sum.tax_sp
```{r}

phyloseq_to_df(surface_co1_sum.tax_sp) %>% 
  dplyr::select(.,-Kingdom,-Phylum, -Class,-Order,-Family,-Genus, -Species,sum.taxonomy, -OTU) -> surface_co1_2020_sum.tax_sp_df
```

```{r}

#Read in Read Data

Data_asvs_sites <- surface_co1_2020_sum.tax_sp_df %>%  ungroup()

#Get OTU names

taxa_asvs_sites <- Data_asvs_sites$sum.taxonomy

#Make a presence/absence matrix
Data_asvs_sites %>% dplyr::select(-sum.taxonomy) %>% 
mutate_all(., ~replace(., . > 1, 1)) -> DataPA_asvs_sites

DataPA_asvs_sites <- as.data.frame(t(DataPA_asvs_sites))
colnames(DataPA_asvs_sites) <- taxa_asvs_sites

#Calculate zeta diversity decay parameters and plots.
surface_co1_2020_sum.tax_zetaDecay.ex_asvs_site <- Zeta.decline.ex(DataPA_asvs_sites,orders=1:20,plot=FALSE)

```

### subsurface_co1_2020_sum.tax_sp
```{r}

phyloseq_to_df(subsurface_co1_sum.tax_sp) %>% 
  dplyr::select(.,-Kingdom,-Phylum, -Class,-Order,-Family,-Genus, -Species,-OTU) -> subsurface_co1_2020_sum.tax_sp_df
```

```{r}

#Read in Read Data

Data_asvs_sites <- subsurface_co1_2020_sum.tax_sp_df %>%  ungroup()

#Get OTU names

taxa_asvs_sites <- Data_asvs_sites$sum.taxonomy

#Make a presence/absence matrix
Data_asvs_sites %>% dplyr::select(-sum.taxonomy) %>% 
mutate_all(., ~replace(., . > 1, 1)) -> DataPA_asvs_sites

DataPA_asvs_sites <- as.data.frame(t(DataPA_asvs_sites))
colnames(DataPA_asvs_sites) <- taxa_asvs_sites

#Calculate zeta diversity decay parameters and plots.
subsurface_co1_2020_sum.tax_sp_zetaDecay.ex_asvs_site <- Zeta.decline.ex(DataPA_asvs_sites,orders=1:20,plot=FALSE)

```


### Zeta Decay Figure

```{r}
subsurface_co1_2020_sum.tax_sp_zetaDecay.ex_asvs_site[1:4] %>% as.tibble() %>% mutate(Group = "Below 30m")-> subsurface_co1_tib

surface_co1_2020_sum.tax_zetaDecay.ex_asvs_site[1:4] %>% as.tibble() %>% mutate(Group = "Above 30m")-> surface_co1_tib

bind_rows(subsurface_co1_tib,surface_co1_tib) -> zdx_regions

zdv_decay_sites <- zdx_regions %>% 
ggplot(aes(x= zeta.order, y= zeta.val, group=Group,color=Group, fill=Group)) +geom_point() +geom_line() + geom_ribbon(aes(ymin = zeta.val - zeta.val.sd, ymax = zeta.val + zeta.val.sd),alpha =0.1, colour = NA) + ggtitle(label="")+theme(axis.line = element_line(colour = "black"),panel.grid.major = element_blank(), panel.grid.minor = element_blank(),panel.background = element_blank(), axis.text=element_text(size=16),
        axis.title=element_text(size=20,face="bold"),legend.position="none",legend.title = element_text( size=12, face="bold"),legend.text = element_text(size=10, 
                                     face="bold"), title = element_text(size=20, face="bold") ) + theme(axis.title.x=element_blank(),legend.text=element_text(size=16), legend.title=element_text(size=20)) +ylab("Zeta Diversity") + xlab("Zeta Order - Stations") + scale_x_continuous(breaks=seq(1, 10, 1), limits=c(1,10))+ scale_colour_manual(values = c("dodgerblue4","#Ed820e"), name = "Group", labels = c("Above 30m", "Below 30m")) + scale_shape_discrete(name = "Group", labels = c("Above 30m", "Below 30m")) + scale_fill_manual(values = c("dodgerblue4","#Ed820e"),name = "Group", labels = c("Above 30m", "Below 30m")) 
```
### Species Retention Figure

```{r}

subsurface_co1_2020_sum.tax_sp_zetaDecay.ex_asvs_site[5] %>% as.tibble() %>% mutate(Group = "Below 30m",zeta.order = 1:n()) -> ret_sub_co1_multi

surface_co1_2020_sum.tax_zetaDecay.ex_asvs_site[5] %>% as.tibble() %>% mutate(Group = "Above 30m",zeta.order = 1:n())-> ret_surf_co1_uni

bind_rows(ret_sub_co1_multi,ret_surf_co1_uni) -> zdx_regions_ret_sites

spec_ret_sites <- zdx_regions_ret_sites %>% 
ggplot(aes(x= zeta.order, y= ratio, group=Group,color=Group)) +geom_point() +geom_line() + ggtitle(label = "") +
 theme(axis.line = element_line(colour = "black"),panel.grid.major = element_blank(), panel.grid.minor = element_blank(),panel.background = element_blank(), axis.text=element_text(size=16),
        axis.title=element_text(size=20,face="bold"),legend.position="none",legend.title = element_text( size=12, face="bold"),legend.text = element_text(size=10, 
                                     face="bold"), title = element_text(size=20, face="bold"),axis.title.x=element_blank() ) + theme(legend.text=element_text(size=16), legend.title=element_text(size=20)) +ylab("Zeta Ratio") + xlab("Zeta Order - Stations") +ylim(0,1) +scale_colour_manual(values = c("dodgerblue4","#Ed820e"), name = "Group", labels = c("Above 30m", "Below 30m")) + scale_shape_discrete(name = "Group", labels = c("Above 30m", "Below 30m")) + scale_fill_manual(values = c("dodgerblue4","#Ed820e"),name = "Group", labels = c("Above 30m", "Below 30m"))+ xlim(0,20) 


```



### surface_MiFish_2020_sum.tax_sp
```{r}

phyloseq_to_df(surface_MiFish_sum.tax_sp) %>% 
  dplyr::select(.,-Kingdom,-Phylum, -Class,-Order,-Family,-Genus, -Species,sum.taxonomy, -OTU) -> surface_MiFish_2020_sum.tax_sp_df
```

```{r}

#Read in Read Data

Data_asvs_sites <- surface_MiFish_2020_sum.tax_sp_df %>%  ungroup()

#Get OTU names

taxa_asvs_sites <- Data_asvs_sites$sum.taxonomy

#Make a presence/absence matrix
Data_asvs_sites %>% dplyr::select(-sum.taxonomy) %>% 
mutate_all(., ~replace(., . > 1, 1)) -> DataPA_asvs_sites

DataPA_asvs_sites <- as.data.frame(t(DataPA_asvs_sites))
colnames(DataPA_asvs_sites) <- taxa_asvs_sites

#Calculate zeta diversity decay parameters and plots.
surface_MiFish_2020_sum.tax_sp_zetaDecay.ex_asvs_site <- Zeta.decline.ex(DataPA_asvs_sites,orders=1:20,plot=FALSE)

```

### subsurface_MiFish_2020_sum.tax_sp
```{r}

phyloseq_to_df(subsurface_MiFish_sum.tax_sp) %>% 
  dplyr::select(.,-Kingdom,-Phylum, -Class,-Order,-Family,-Genus, -Species,-OTU) -> subsurface_MiFish_2020_sum.tax_sp_df
```

```{r}

#Read in Read Data

Data_asvs_sites <- subsurface_MiFish_2020_sum.tax_sp_df %>%  ungroup()

#Get OTU names

taxa_asvs_sites <- Data_asvs_sites$sum.taxonomy

#Make a presence/absence matrix
Data_asvs_sites %>% dplyr::select(-sum.taxonomy) %>% 
mutate_all(., ~replace(., . > 1, 1)) -> DataPA_asvs_sites

DataPA_asvs_sites <- as.data.frame(t(DataPA_asvs_sites))
colnames(DataPA_asvs_sites) <- taxa_asvs_sites

#Calculate zeta diversity decay parameters and plots.
subsurface_MiFish_2020_sum.tax_sp_zetaDecay.ex_asvs_site <- Zeta.decline.ex(DataPA_asvs_sites,orders=1:20,plot=FALSE)

```


### Zeta Decay Figure

```{r}
subsurface_MiFish_2020_sum.tax_sp_zetaDecay.ex_asvs_site[1:4] %>% as.tibble() %>% mutate(Group = "Below 30m")-> subsurface_MiFish_tib

surface_MiFish_2020_sum.tax_sp_zetaDecay.ex_asvs_site[1:4] %>% as.tibble() %>% mutate(Group = "Above 30m")-> surface_MiFish_tib

bind_rows(subsurface_MiFish_tib,surface_MiFish_tib) -> zdx_regions_mifish

zdv_decay_sites_mifish <- zdx_regions_mifish %>% 
ggplot(aes(x= zeta.order, y= zeta.val, group=Group,color=Group, fill=Group)) +geom_point() +geom_line() + geom_ribbon(aes(ymin = zeta.val - zeta.val.sd, ymax = zeta.val + zeta.val.sd),alpha =0.1, colour = NA) + ggtitle(label = "Turnover")+theme(axis.line = element_line(colour = "black"),panel.grid.major = element_blank(), panel.grid.minor = element_blank(),panel.background = element_blank(), axis.text=element_text(size=16),
        axis.title=element_text(size=20,face="bold"),legend.position="none",legend.title = element_text( size=12, face="bold"),legend.text = element_text(size=10, 
                                     face="bold"),axis.title.x=element_blank(), title = element_text(size=20, face="bold") ) + theme(legend.text=element_text(size=16), legend.title=element_text(size=20)) +ylab("Zeta Diversity") + xlab("Zeta Order - Stations") + scale_x_continuous(breaks=seq(1, 10, 1), limits=c(1,10)) +scale_colour_manual(values = c("dodgerblue4","#Ed820e"), name = "Group", labels = c("Above 30m", "Below 30m")) + scale_shape_discrete(name = "Group", labels = c("Above 30m", "Below 30m")) + scale_fill_manual(values = c("dodgerblue4","#Ed820e"),name = "Group", labels = c("Above 30m", "Below 30m")) 
```
### Species Retention Figure

```{r}

subsurface_MiFish_2020_sum.tax_sp_zetaDecay.ex_asvs_site[5] %>% as.tibble() %>% mutate(Group = "Below 30m",zeta.order = 1:n()) -> ret_sub_MiFish_multi

surface_MiFish_2020_sum.tax_sp_zetaDecay.ex_asvs_site[5] %>% as.tibble() %>% mutate(Group = "Above 30m",zeta.order = 1:n())-> ret_surf_MiFish_uni

bind_rows(ret_sub_MiFish_multi,ret_surf_MiFish_uni) -> zdx_regions_ret_sites_Mifish

spec_ret_sites_mifish <- zdx_regions_ret_sites_Mifish %>% 
ggplot(aes(x= zeta.order, y= ratio, group=Group,color=Group)) +geom_point() +geom_line() + ggtitle(label = "Retention") +
 theme(axis.line = element_line(colour = "black"),panel.grid.major = element_blank(), panel.grid.minor = element_blank(),panel.background = element_blank(), axis.text=element_text(size=16),
        axis.title=element_text(size=20,face="bold"),legend.position="none",axis.title.x=element_blank(),legend.title = element_text( size=12, face="bold"),legend.text = element_text(size=10, 
                                     face="bold"), title = element_text(size=20, face="bold") ) + theme(legend.text=element_text(size=16), legend.title=element_text(size=20)) +ylab("Zeta Ratio") + xlab("Zeta Order - Stations") +ylim(0,1) +scale_colour_manual(values = c("dodgerblue4","#Ed820e"), name = "Group", labels = c("Above 30m", "Below 30m")) + scale_shape_discrete(name = "Group", labels = c("Above 30m", "Below 30m")) + scale_fill_manual(values = c("dodgerblue4","#Ed820e"),name = "Group", labels = c("Above 30m", "Below 30m"))+ xlim(0,20) 


```

### surface_Machida_2020_sum.tax_sp
```{r}

phyloseq_to_df(surface_Machida_2020_sum.tax_sp) %>% 
  dplyr::select(.,-Kingdom,-Phylum, -Class,-Order,-Family,-Genus, -Species,sum.taxonomy, -OTU) -> surface_Machida_2020_sum.tax_sp_df
```

```{r}

#Read in Read Data

Data_asvs_sites <- surface_Machida_2020_sum.tax_sp_df %>%  ungroup()

#Get OTU names

taxa_asvs_sites <- Data_asvs_sites$sum.taxonomy

#Make a presence/absence matrix
Data_asvs_sites %>% dplyr::select(-sum.taxonomy) %>% 
mutate_all(., ~replace(., . > 1, 1)) -> DataPA_asvs_sites

DataPA_asvs_sites <- as.data.frame(t(DataPA_asvs_sites))
colnames(DataPA_asvs_sites) <- taxa_asvs_sites

#Calculate zeta diversity decay parameters and plots.
surface_Machida_2020_sum.tax_sp_zetaDecay.ex_asvs_site <- Zeta.decline.ex(DataPA_asvs_sites,orders=1:100,plot=FALSE)

```

### subsurface_MiFish_2020_sum.tax_sp
```{r}

phyloseq_to_df(subsurface_Machida_2020_sum.tax_sp) %>% 
  dplyr::select(.,-Kingdom,-Phylum, -Class,-Order,-Family,-Genus, -Species,-OTU) -> subsurface_Machida_2020_sum.tax_sp_df
```

```{r}

#Read in Read Data

Data_asvs_sites <- subsurface_Machida_2020_sum.tax_sp_df %>%  ungroup()

#Get OTU names

taxa_asvs_sites <- Data_asvs_sites$sum.taxonomy

#Make a presence/absence matrix
Data_asvs_sites %>% dplyr::select(-sum.taxonomy) %>% 
mutate_all(., ~replace(., . > 1, 1)) -> DataPA_asvs_sites

DataPA_asvs_sites <- as.data.frame(t(DataPA_asvs_sites))
colnames(DataPA_asvs_sites) <- taxa_asvs_sites

#Calculate zeta diversity decay parameters and plots.
subsurface_Machida_2020_sum.tax_sp_zetaDecay.ex_asvs_site <- Zeta.decline.ex(DataPA_asvs_sites,orders=1:100,plot=FALSE)

```


### Zeta Decay Figure

```{r}
subsurface_Machida_2020_sum.tax_sp_zetaDecay.ex_asvs_site[1:4] %>% as.tibble() %>% mutate(Group = "Below 30m")-> subsurface_Machida_tib

surface_Machida_2020_sum.tax_sp_zetaDecay.ex_asvs_site[1:4] %>% as.tibble() %>% mutate(Group = "Above 30m")-> surface_Machida_tib

bind_rows(subsurface_Machida_tib,surface_Machida_tib) -> zdx_regions_machida

zdv_decay_sites_machida <- zdx_regions_machida %>% 
ggplot(aes(x= zeta.order, y= zeta.val, group=Group,color=Group, fill=Group)) +geom_point() +geom_line() + geom_ribbon(aes(ymin = zeta.val - zeta.val.sd, ymax = zeta.val + zeta.val.sd),alpha =0.1, colour = NA) + ggtitle(label = "")+theme(axis.line = element_line(colour = "black"),panel.grid.major = element_blank(), panel.grid.minor = element_blank(),panel.background = element_blank(), axis.text=element_text(size=16),
        axis.title=element_text(size=20,face="bold"),legend.position="none",legend.title = element_text( size=12, face="bold"),legend.text = element_text(size=10, 
                                     face="bold"), title = element_text(size=20, face="bold") ) + theme(legend.text=element_text(size=16), legend.title=element_text(size=20)) +ylab("Zeta Diversity") + xlab("Zeta Order - Stations") + scale_x_continuous(breaks=seq(1, 10, 1), limits=c(1,10)) +scale_colour_manual(values = c("dodgerblue4","#Ed820e"), name = "Group", labels = c("Above 30m", "Below 30m")) + scale_shape_discrete(name = "Group", labels = c("Above 30m", "Below 30m")) + scale_fill_manual(values = c("dodgerblue4","#Ed820e"),name = "Group", labels = c("Above 30m", "Below 30m")) 
```
### Species Retention Figure

```{r}

subsurface_Machida_2020_sum.tax_sp_zetaDecay.ex_asvs_site[5] %>% as.tibble() %>% mutate(Group = "Below 30m",zeta.order = 1:n()) -> ret_sub_Machida_multi

surface_Machida_2020_sum.tax_sp_zetaDecay.ex_asvs_site[5] %>% as.tibble() %>% mutate(Group = "Above 30m",zeta.order = 1:n())-> ret_surf_Machida_uni

bind_rows(ret_sub_Machida_multi,ret_surf_Machida_uni) -> zdx_regions_ret_sites_Machida

spec_ret_sites_Machida <- zdx_regions_ret_sites_Machida %>% 
ggplot(aes(x= zeta.order, y= ratio, group=Group,color=Group)) +geom_point() +geom_line() + ggtitle(label = "") +
 theme(axis.line = element_line(colour = "black"),panel.grid.major = element_blank(), panel.grid.minor = element_blank(),panel.background = element_blank(), axis.text=element_text(size=16),
        axis.title=element_text(size=20,face="bold"),legend.position="none",legend.title = element_text( size=12, face="bold"),legend.text = element_text(size=10, 
                                     face="bold"), title = element_text(size=20, face="bold") ) + theme(legend.text=element_text(size=16), legend.title=element_text(size=20)) +ylab("Zeta Ratio") + xlab("Zeta Order - Stations") +ylim(0,1) +scale_colour_manual(values = c("dodgerblue4","#Ed820e"), name = "Group", labels = c("Above 30m", "Below 30m")) + scale_shape_discrete(name = "Group", labels = c("Above 30m", "Below 30m")) + scale_fill_manual(values = c("dodgerblue4","#Ed820e"),name = "Group", labels = c("Above 30m", "Below 30m"))+ xlim(0,20) 


```

```{r}
(plot_fish + zdv_decay_sites_mifish + spec_ret_sites_mifish)/ (plot_co1 + zdv_decay_sites + spec_ret_sites)  +
  plot_layout(guides = 'collect')


ggsave(filename = here("gold_out","Zeta_diveristy.png"),
       width=14,
       height = 10,
       dpi = 400,
      units = c("in"))
```



## Pseudocalanus
```{r}

long_COI_EcoFOCI %>% 
  dplyr::select(ASV_combo, Kingdom, Phylum, Class, Order, Family, Genus, Species) %>% 
  distinct() %>% 
  filter(., Genus=="Pseudocalanus")
```

```{r}
long_COI_EcoFOCI %>% 
  filter(., Genus=="Pseudocalanus") %>% 
  dplyr::group_by(ASV_combo, Kingdom, Phylum, Class, Order, Family, Genus, Species) %>%
  dplyr::summarise(sum_reads=sum(nReads)) %>% 
  arrange(desc(sum_reads)) -> Pseudocalanus_tot

Pseudocalanus_tot %>% 
  filter(., sum_reads > 5000) -> Pseudocalanus_to_keep
  # Pseudocalanus mimus versus Pseudocalanus minutus	

```

```{r}

c(setdiff(varnames,to_remove),"Tot") -> varnames_to_group

long_COI_EcoFOCI %>% 
 group_by(Sample) %>%
    mutate(Tot = sum(nReads)) %>% 
    filter(., Genus=="Pseudocalanus") %>% 
    filter(., !is.na(Species)) %>% 
    group_by(!!!syms(varnames_to_group)) %>% 
    dplyr::summarise(nReads=sum(nReads)) %>% 
    mutate (Prop.abund = nReads / Tot) %>% 
      ungroup() %>% 
    group_by (Species) %>%
    mutate (Colmax = max(Prop.abund),
            eDNA.Index = Prop.abund / Colmax) %>% 
  mutate(., Type = case_when(Species == "Pseudocalanus mimus" ~ "Temperate",
                             Species == "Pseudocalanus newmani"~ "Temperate",
                             Species ==  "Pseudocalanus minutus" ~"Arctic",
                             Species ==  "Pseudocalanus acuspes" ~"Arctic",
                             Species ==  "Pseudocalanus moultoni" ~"Temperate",
                             TRUE ~"Unknown"))-> Pseudocalanus_Df

```

```{r}

Pseudocalanus_Df %>% 
  filter(., Species =="Pseudocalanus mimus") %>% 
  ggplot(., aes(y=eDNA.Index, x=Temp_C)) +geom_point() + facet_wrap(Species~.) +
  stat_poly_line() +
  stat_poly_eq() +
  geom_point()
```


```{r}


Pseudocalanus_Df %>% 
  ggplot(., aes(y=log10(nReads), x=Temp_C, color=Cruise_ID_short)) +geom_point() + facet_wrap(Species~Type)
```


```{r}

Pseudocalanus_Df %>% 
  mutate(., PA = if_else(nReads >0, 1,0)) %>% 
  ggplot(., aes(y=PA, x=Temp_C)) +geom_point() + facet_wrap(Species~Type) +
  geom_smooth(method = "glm", 
    method.args = list(family = "binomial"), 
    se = FALSE) 
```

```{r}
Pseudocalanus_Df %>% 
  mutate(., PA = if_else(nReads >0, 1,0)) %>% 
  mutate(., Depth_bin = if_else(Depth_m >29, "bottom","surface" )) %>% 
  ggplot(., aes(y=lat, x=Temp_C, colour = PA)) +geom_point() + facet_wrap(Species~Type~Depth_bin)
```

```{r}
Pseudocalanus_Df %>% 
  mutate(., PA = if_else(nReads >0, 1,0)) %>% 
  ggplot(., aes(y=log10(Chla_ugrams.per.l), x=Temp_C, colour = eDNA.Index)) +geom_point() + facet_wrap(Species~Type) +    scale_color_gradientn(colours = c("black", "#C6DBEF", 'dodgerblue4'),
                       values = c(0, .Machine$double.eps, 1))
```

```{r}
Pseudocalanus_Df %>% 
  mutate(., PA = if_else(nReads >0, 1,0)) %>% 
  mutate(., Time2 = as.POSIXct(Time)) %>% 
  ggplot(., aes(x=Time2, y=Temp_C, colour = eDNA.Index)) +geom_point() + facet_wrap(Species~Type) +    scale_color_gradientn(colours = c("grey80", "lightyellow", 'red'),
                       values = c(0, .Machine$double.eps, 1)) +theme_bw()
```

```{r}
Pseudocalanus_Df %>% 
  mutate(., PA = if_else(nReads >0, 1,0)) %>% 
  filter(., PA >0) %>% 
  ggplot(., aes(, x=Temp_C)) +geom_density() + facet_wrap(Species~Type)

```

```{r}
Pseudocalanus_Df %>% 
  mutate(., PA = if_else(nReads > 0, "Present","Absent")) %>% 
  mutate(., Temp_bin = if_else(Temp_C > 2, "Warm","Cold")) %>% 
  group_by(Species, Temp_bin, PA) %>% 
  dplyr::summarise(N=n()) %>% 
  mutate(freq = N / sum(N)) %>% 
  filter(., !is.na(Temp_bin)) %>% 
  mutate(., Type = case_when(Species == "Pseudocalanus mimus" ~ "Temperate",
                             Species == "Pseudocalanus newmani"~ "Temperate",
                             Species ==  "Pseudocalanus minutus" ~"Arctic",
                             Species ==  "Pseudocalanus acuspes" ~"Arctic",
                             Species ==  "Pseudocalanus moultoni" ~"Temperate",
                             TRUE ~"Unknown")) %>% 
ggplot(., aes(x=PA, y =freq)) + geom_col() + facet_wrap(Species~Type ~Temp_bin)
```


