---
title: "REVAMP & Sample Metadata Merge Into Long Format Files"
author: "Zack Gold"
date: "2025-01-03"
output: html_document
---

```{r}
library(here)
library(tidyverse)
library(Biostrings)
```

# Load in Sample Metadata

```{r}
run1_4_revamp_formatted_metadata <- read.csv(here("run1_4_revamp_formatted_metadata.csv"))

run1_4_revamp_formatted_metadata %>% 
  filter(., str_detect(Sample, "E2084_SKQ23_12S"))
```

# 18Sv4 - REVAMP

## Unite ASVs

```{r}
run2_18Sv4_tax <- read.table(here("data","REVAMP_out","Run2_from_poseidon_20250103","18Sv4","ASV2Taxonomy","18Sv4_RUN1_asvTaxonomyTable.txt"), sep="\t", header=T) #IS IT REALLY CALLED RUN1 when it is RUN2?

run2_18Sv4_fasta <- readDNAStringSet(here("data","REVAMP_out","Run2_from_poseidon_20250103","18Sv4","dada2","ASVs.fa"), format="fasta" )

as.data.frame(run2_18Sv4_fasta) %>%
  as_tibble(rownames = "ASV") %>% 
  left_join(run2_18Sv4_tax) %>% 
  rename_with(.cols = everything(), function(x){paste0("Run2:", x)}) %>% 
  dplyr::rename(Sequence=`Run2:x`) -> run2_18Sv4_fasta_df

run3_18Sv4_tax <- read.table(here("data","REVAMP_out","Run3_from_poseidon_20250103","18Sv4_REVAMP_out","ASV2Taxonomy","18Sv4_REVAMP_out_asvTaxonomyTable.txt"), sep="\t", header=T) 

run3_18Sv4_fasta <- readDNAStringSet(here("data","REVAMP_out","Run3_from_poseidon_20250103","18Sv4_REVAMP_out","dada2","ASVs.fa"), format="fasta" )

as.data.frame(run3_18Sv4_fasta) %>% 
  as_tibble(rownames = "ASV") %>% 
  left_join(run3_18Sv4_tax) %>% 
   rename_with(.cols = everything(), function(x){paste0("Run3:", x)}) %>% 
  dplyr::rename(Sequence=`Run3:x`)-> run3_18Sv4_fasta_df

full_join(run2_18Sv4_fasta_df, run3_18Sv4_fasta_df) %>% 
 mutate(., ASV_combo = row_number(),
        ASV_combo = str_c("ASVc_",ASV_combo)) %>% 
 mutate(., Kingdom = if_else(!is.na(`Run3:Kingdom`), `Run3:Kingdom`,`Run2:Kingdom`),
        Phylum= if_else(!is.na(`Run3:Phylum`), `Run3:Phylum`,`Run2:Phylum`), 
        Class= if_else(!is.na(`Run3:Class`), `Run3:Class`,`Run2:Class`),
        Order= if_else(!is.na(`Run3:Order`), `Run3:Order`,`Run2:Order`), 
        Family= if_else(!is.na(`Run3:Family`), `Run3:Family`,`Run2:Family`),
        Genus= if_else(!is.na(`Run3:Genus`), `Run3:Genus`,`Run2:Genus`),
        Species= if_else(!is.na(`Run3:Species`), `Run3:Species`,`Run2:Species`)) %>% 
  dplyr::select(ASV_combo, `Run3:ASV`, `Run2:ASV`, Sequence, Kingdom,Phylum,Class, Order, Family, Genus, Species) -> run2_3_18Sv4_united_ASV_tables

saveRDS(run2_3_18Sv4_united_ASV_tables, here("data","long_data","run2_3_18Sv4_united_ASV_tables.RDS"))

```

## Run 1 - There are no EcoFOCI samples in Run1
I am including this here because there is something a little off about the taxonomy for 18Sv4 and the REVAMP output with Run 1 and Run 2 taxonomy appearing to be flip flopped.
```{r}

# 
# run1_18Sv4_asv <- read.table(here("data","REVAMP_out","run1_from_poseidon_20250103","18Sv4","dada2","ASVs_counts.tsv"), sep="\t", header=T) %>% 
#   rename(ASV=x)
# 
# run1_18Sv4_tax <- read.table(here("data","REVAMP_out","run1_from_poseidon_20250103","18Sv4","ASV2Taxonomy","18Sv4_RUN2_sansPOS_asvTaxonomyTable.txt"), sep="\t", header=T) #IS IT REALLY CALLED RUN2 when it is run1?
# 
# run1_18Sv4_asv %>% 
#   left_join(run1_18Sv4_tax) %>% 
#    rename_at(vars(starts_with("MP_")), funs(str_replace(., "MP_", ""))) -> run1_18Sv4_merge
# 
# run1_18Sv4_merge %>% 
#   pivot_longer(cols = c(-ASV, -Kingdom, -Phylum, -Class, -Order, -Family,-Genus, -Species), names_to = "Sample", values_to = "nReads") %>% 
#   left_join(run1_4_revamp_formatted_metadata) %>% 
#   filter(., !is.na(Sample_Name)) %>% # this removes all non-EcoFOCI samples. Note this also excludes positive and negative controls. No decontamination has been performed. We should definitely do this at somepoint, but need a conversation as to what approaches
#     group_by(Sample) %>%
#     mutate (Tot = sum(nReads),
#             Prop.abund = nReads / Tot) %>% 
#     group_by (ASV) %>%
#     mutate (Colmax = max(Prop.abund),
#             eDNA.Index = Prop.abund / Colmax) %>%
#   dplyr::select(-Tot,-Colmax) -> run1_18Sv4_long
  
```


## Run 2
```{r}

run2_18Sv4_asv <- read.table(here("data","REVAMP_out","Run2_from_poseidon_20250103","18Sv4","dada2","ASVs_counts.tsv"), sep="\t", header=T) %>% 
  rename(ASV=x)

run2_18Sv4_asv %>% 
  dplyr::rename(`Run2:ASV`=ASV) %>% 
  left_join(run2_3_18Sv4_united_ASV_tables) %>% 
   rename_at(vars(starts_with("MP_")), funs(str_replace(., "MP_", ""))) -> run2_18Sv4_merge
colnames(run2_18Sv4_merge)

run2_18Sv4_merge %>% 
  pivot_longer(cols = c(-`ASV_combo`, -`Run2:ASV`, -`Run3:ASV`,-`Sequence`,-Kingdom, -Phylum, -Class, -Order, -Family, -Genus, -Species), names_to = "Sample", values_to = "nReads") %>% 
  left_join(run1_4_revamp_formatted_metadata) %>% 
  filter(., !is.na(Sample_Name)) %>% # this removes all non-EcoFOCI samples. Note this also excludes positive and negative controls. No decontamination has been performed. We should definitely do this at somepoint, but need a conversation as to what approaches
    group_by(Sample) %>%
    mutate (Tot = sum(nReads),
            Prop.abund = nReads / Tot) %>% 
    group_by (ASV_combo) %>%
    mutate (Colmax = max(Prop.abund),
            eDNA.Index = Prop.abund / Colmax) %>%
  dplyr::select(-Tot,-Colmax) -> run2_18Sv4_long
  
  
```


## Run 3
```{r}


run3_18Sv4_asv <- read.table(here("data","REVAMP_out","Run3_from_poseidon_20250103","18Sv4_REVAMP_out","dada2","ASVs_counts.tsv"), sep="\t", header=T) %>% 
  dplyr::rename(ASV=x)

run3_18Sv4_asv %>% 
  dplyr::rename(`Run3:ASV`=ASV) %>% 
  left_join(run2_3_18Sv4_united_ASV_tables) %>% 
   rename_at(vars(starts_with("MP_")), funs(str_replace(., "MP_", ""))) -> run3_18Sv4_merge

run3_18Sv4_merge %>% 
   pivot_longer(cols = c(-`ASV_combo`, -`Run2:ASV`, -`Run3:ASV`,-`Sequence`,-Kingdom, -Phylum, -Class, -Order, -Family, -Genus, -Species), names_to = "Sample", values_to = "nReads") %>% 
  left_join(run1_4_revamp_formatted_metadata) %>% 
  filter(., !is.na(Sample_Name)) %>% # this removes all non-EcoFOCI samples. Note this also excludes positive and negative controls. No decontamination has been performed. We should definitely do this at somepoint, but need a conversation as to what approaches
    group_by(Sample) %>%
    mutate (Tot = sum(nReads),
            Prop.abund = nReads / Tot) %>% 
    group_by (ASV_combo) %>%
    mutate (Colmax = max(Prop.abund),
            eDNA.Index = Prop.abund / Colmax) %>%
  dplyr::select(-Tot,-Colmax) -> run3_18Sv4_long
  
  
```

## Merge

```{r}

bind_rows(run2_18Sv4_long, run3_18Sv4_long) -> long_18Sv4_EcoFOCI

saveRDS(long_18Sv4_EcoFOCI, here("data","long_data","long_18Sv4_EcoFOCI.RDS"))

```

# 18Sv4 - PR2

## Unite ASVs

```{r}
run2_18Sv4_tax_pr2 <- read.table(here("data","pr2_setta_20240106","Run2_18Sv4_tax_pr2_sklearn.tsv"), sep="\t", header=T) %>% 
  dplyr::rename(ASV=Feature.ID) %>% separate(., Taxon, into=c("Domain","Supergroup","Division","Subdivision","Class","Order","Family","Genus","Species"), sep=";")

as.data.frame(run2_18Sv4_fasta) %>%
  as_tibble(rownames = "ASV") %>% 
  left_join(run2_18Sv4_tax_pr2) %>% 
  rename_with(.cols = everything(), function(x){paste0("Run2:", x)}) %>% 
  dplyr::rename(Sequence=`Run2:x`) -> run2_18Sv4_fasta_df_pr2

run3_18Sv4_tax_pr2 <- read.table(here("data","pr2_setta_20240106","Run3_18Sv4_tax_pr2_sklearn.tsv"), sep="\t", header=T) %>% 
  dplyr::rename(ASV=Feature.ID) %>% separate(., Taxon, into=c("Domain","Supergroup","Division","Subdivision","Class","Order","Family","Genus","Species"), sep=";")

as.data.frame(run3_18Sv4_fasta) %>% 
  as_tibble(rownames = "ASV") %>% 
  left_join(run3_18Sv4_tax_pr2) %>% 
   rename_with(.cols = everything(), function(x){paste0("Run3:", x)}) %>% 
  dplyr::rename(Sequence=`Run3:x`)-> run3_18Sv4_fasta_df_pr2

full_join(run2_18Sv4_fasta_df_pr2, run3_18Sv4_fasta_df_pr2) %>% 
 mutate(., ASV_combo = row_number(),
        ASV_combo = str_c("ASVc_",ASV_combo)) %>% 
 mutate(., Domain = if_else(!is.na(`Run3:Domain`), `Run3:Domain`,`Run2:Domain`),
        Supergroup= if_else(!is.na(`Run3:Supergroup`), `Run3:Supergroup`,`Run2:Supergroup`), 
         Division= if_else(!is.na(`Run3:Division`), `Run3:Division`,`Run2:Division`), 
         Subdivision= if_else(!is.na(`Run3:Subdivision`), `Run3:Subdivision`,`Run2:Subdivision`), 
        Class= if_else(!is.na(`Run3:Class`), `Run3:Class`,`Run2:Class`),
        Order= if_else(!is.na(`Run3:Order`), `Run3:Order`,`Run2:Order`), 
        Family= if_else(!is.na(`Run3:Family`), `Run3:Family`,`Run2:Family`),
        Genus= if_else(!is.na(`Run3:Genus`), `Run3:Genus`,`Run2:Genus`),
        Species= if_else(!is.na(`Run3:Species`), `Run3:Species`,`Run2:Species`)) %>% 
  dplyr::select(ASV_combo, `Run3:ASV`, `Run2:ASV`, Sequence, Domain,Supergroup,Division,Subdivision,Class, Order, Family, Genus, Species) -> run2_3_18Sv4_pr2_united_ASV_tables

saveRDS(run2_3_18Sv4_pr2_united_ASV_tables, here("data","long_data","run2_3_18Sv4_pr2_united_ASV_tables.RDS"))

```

## Run 1 - There are no EcoFOCI samples in Run1

## Run 2
```{r}

run2_18Sv4_asv %>% 
  dplyr::rename(`Run2:ASV`=ASV) %>% 
  left_join(run2_3_18Sv4_pr2_united_ASV_tables) %>% 
   rename_at(vars(starts_with("MP_")), funs(str_replace(., "MP_", ""))) -> run2_18Sv4_merge_pr2

run2_18Sv4_merge_pr2 %>% 
  pivot_longer(cols = c(-`ASV_combo`, -`Run2:ASV`, -`Run3:ASV`,-`Sequence`,-Domain,-Supergroup,-Division,-Subdivision, -Class, -Order, -Family, -Genus, -Species), names_to = "Sample", values_to = "nReads") %>% 
  left_join(run1_4_revamp_formatted_metadata) %>% 
  filter(., !is.na(Sample_Name)) %>% # this removes all non-EcoFOCI samples. Note this also excludes positive and negative controls. No decontamination has been performed. We should definitely do this at somepoint, but need a conversation as to what approaches
    group_by(Sample) %>%
    mutate (Tot = sum(nReads),
            Prop.abund = nReads / Tot) %>% 
    group_by (ASV_combo) %>%
    mutate (Colmax = max(Prop.abund),
            eDNA.Index = Prop.abund / Colmax) %>%
  dplyr::select(-Tot,-Colmax) -> run2_18Sv4_long_pr2
  
  
```


## Run 3
```{r}

run3_18Sv4_asv %>% 
  dplyr::rename(`Run3:ASV`=ASV) %>% 
  left_join(run2_3_18Sv4_pr2_united_ASV_tables) %>% 
   rename_at(vars(starts_with("MP_")), funs(str_replace(., "MP_", ""))) -> run3_18Sv4_merge_pr2

run3_18Sv4_merge_pr2 %>% 
  pivot_longer(cols = c(-`ASV_combo`, -`Run2:ASV`, -`Run3:ASV`,-`Sequence`,-Domain,-Supergroup,-Division,-Subdivision, -Class, -Order, -Family, -Genus, -Species), names_to = "Sample", values_to = "nReads") %>% 
  left_join(run1_4_revamp_formatted_metadata) %>% 
  filter(., !is.na(Sample_Name)) %>% # this removes all non-EcoFOCI samples. Note this also excludes positive and negative controls. No decontamination has been performed. We should definitely do this at somepoint, but need a conversation as to what approaches
    group_by(Sample) %>%
    mutate (Tot = sum(nReads),
            Prop.abund = nReads / Tot) %>% 
    group_by (ASV_combo) %>%
    mutate (Colmax = max(Prop.abund),
            eDNA.Index = Prop.abund / Colmax) %>%
  dplyr::select(-Tot,-Colmax) -> run3_18Sv4_long_pr2
  
  
```

## Merge

```{r}

bind_rows(run2_18Sv4_long_pr2, run3_18Sv4_long_pr2) -> long_18Sv4_EcoFOCI_pr2

saveRDS(long_18Sv4_EcoFOCI_pr2, here("data","long_data","long_18Sv4_EcoFOCI_pr2.RDS"))

```

# 18Sv9

## Unite ASVs

```{r}
run2_18Sv9_tax <- read.table(here("data","REVAMP_out","Run2_from_poseidon_20250103","18Sv9","ASV2Taxonomy","18Sv9_RUN1_asvTaxonomyTable.txt"), sep="\t", header=T) #IS IT REALLY CALLED RUN1 when it is RUN2?

run2_18Sv9_fasta <- readDNAStringSet(here("data","REVAMP_out","Run2_from_poseidon_20250103","18Sv9","dada2","ASVs.fa"), format="fasta")

as.data.frame(run2_18Sv9_fasta) %>%
  as_tibble(rownames = "ASV") %>% 
  left_join(run2_18Sv9_tax) %>% 
  rename_with(.cols = everything(), function(x){paste0("Run2:", x)}) %>% 
  dplyr::rename(Sequence=`Run2:x`) -> run2_18Sv9_fasta_df

run3_18Sv9_tax <- read.table(here("data","REVAMP_out","Run3_from_poseidon_20250103","18Sv9_REVAMP_out","ASV2Taxonomy","18Sv9_REVAMP_out_asvTaxonomyTable.txt"), sep="\t", header=T) 

run3_18Sv9_fasta <- readDNAStringSet(here("data","REVAMP_out","Run3_from_poseidon_20250103","18Sv9_REVAMP_out","dada2","ASVs.fa"), format="fasta" )

as.data.frame(run3_18Sv9_fasta) %>% 
  as_tibble(rownames = "ASV") %>% 
  left_join(run3_18Sv9_tax) %>% 
   rename_with(.cols = everything(), function(x){paste0("Run3:", x)}) %>% 
  dplyr::rename(Sequence=`Run3:x`)-> run3_18Sv9_fasta_df

full_join(run2_18Sv9_fasta_df, run3_18Sv9_fasta_df) %>% 
 mutate(., ASV_combo = row_number(),
        ASV_combo = str_c("ASVc_",ASV_combo)) %>% 
 mutate(., Kingdom = if_else(!is.na(`Run3:Kingdom`), `Run3:Kingdom`,`Run2:Kingdom`),
        Phylum= if_else(!is.na(`Run3:Phylum`), `Run3:Phylum`,`Run2:Phylum`), 
        Class= if_else(!is.na(`Run3:Class`), `Run3:Class`,`Run2:Class`),
        Order= if_else(!is.na(`Run3:Order`), `Run3:Order`,`Run2:Order`), 
        Family= if_else(!is.na(`Run3:Family`), `Run3:Family`,`Run2:Family`),
        Genus= if_else(!is.na(`Run3:Genus`), `Run3:Genus`,`Run2:Genus`),
        Species= if_else(!is.na(`Run3:Species`), `Run3:Species`,`Run2:Species`)) %>% 
  dplyr::select(ASV_combo, `Run3:ASV`, `Run2:ASV`, Sequence, Kingdom,Phylum,Class, Order, Family, Genus, Species) -> run2_3_18Sv9_united_ASV_tables

saveRDS(run2_3_18Sv9_united_ASV_tables, here("data","long_data","run2_3_18Sv9_united_ASV_tables.RDS"))

```

## Run 2
```{r}

run2_18Sv9_asv <- read.table(here("data","REVAMP_out","Run2_from_poseidon_20250103","18Sv9","dada2","ASVs_counts.tsv"), sep="\t", header=T) %>% 
  dplyr::rename(ASV=x)

run2_18Sv9_asv %>% 
 dplyr::rename(`Run2:ASV`=ASV) %>% 
  left_join(run2_3_18Sv9_united_ASV_tables) %>% 
  rename_at(vars(starts_with("MP_")), funs(str_replace(., "MP_", ""))) -> run2_18Sv9_merge

run2_18Sv9_merge %>% 
    pivot_longer(cols = c(-`ASV_combo`, -`Run2:ASV`, -`Run3:ASV`,-`Sequence`,-Kingdom, -Phylum, -Class, -Order, -Family, -Genus, -Species), names_to = "Sample", values_to = "nReads") %>% 
  left_join(run1_4_revamp_formatted_metadata) %>% 
  filter(., !is.na(Sample_Name)) %>% # this removes all non-EcoFOCI samples. Note this also excludes positive and negative controls. No decontamination has been performed. We should definitely do this at somepoint, but need a conversation as to what approaches
    group_by(Sample) %>%
    mutate (Tot = sum(nReads),
            Prop.abund = nReads / Tot) %>% 
    group_by (ASV_combo) %>%
    mutate (Colmax = max(Prop.abund),
            eDNA.Index = Prop.abund / Colmax) %>%
  dplyr::select(-Tot,-Colmax)-> run2_18Sv9_long
  
  
```


## Run 3
```{r}


run3_18Sv9_asv <- read.table(here("data","REVAMP_out","Run3_from_poseidon_20250103","18Sv9_REVAMP_out","dada2","ASVs_counts.tsv"), sep="\t", header=T) %>% 
  dplyr::rename(ASV=x)

run3_18Sv9_asv %>% 
  dplyr::rename(`Run3:ASV`=ASV) %>% 
  left_join(run2_3_18Sv4_united_ASV_tables) %>% 
   rename_at(vars(starts_with("MP_")), funs(str_replace(., "MP_", ""))) -> run3_18Sv9_merge

run3_18Sv9_merge %>% 
    pivot_longer(cols = c(-`ASV_combo`, -`Run2:ASV`, -`Run3:ASV`,-`Sequence`,-Kingdom, -Phylum, -Class, -Order, -Family, -Genus, -Species), names_to = "Sample", values_to = "nReads") %>% 
  left_join(run1_4_revamp_formatted_metadata) %>% 
  filter(., !is.na(Sample_Name)) %>% # this removes all non-EcoFOCI samples. Note this also excludes positive and negative controls. No decontamination has been performed. We should definitely do this at somepoint, but need a conversation as to what approaches
    group_by(Sample) %>%
    mutate (Tot = sum(nReads),
            Prop.abund = nReads / Tot) %>% 
    group_by (ASV_combo) %>%
    mutate (Colmax = max(Prop.abund),
            eDNA.Index = Prop.abund / Colmax) %>%
  dplyr::select(-Tot,-Colmax) -> run3_18Sv9_long
  
  
```

## Merge

```{r}

bind_rows(run2_18Sv9_long, run3_18Sv9_long) -> long_18Sv9_EcoFOCI

saveRDS(long_18Sv9_EcoFOCI, here("data","long_data","long_18Sv9_EcoFOCI.RDS"))

```


# COI

## Unite ASVs

```{r}
run2_COI_tax <- read.table(here("data","REVAMP_out","Run2_from_poseidon_20250103","COI","ASV2Taxonomy","COI_RUN1_asvTaxonomyTable.txt"), sep="\t", header=T) #IS IT REALLY CALLED RUN1 when it is RUN2?

run2_COI_fasta <- readDNAStringSet(here("data","REVAMP_out","Run2_from_poseidon_20250103","COI","dada2","ASVs.fa"), format="fasta" )

as.data.frame(run2_COI_fasta) %>%
  as_tibble(rownames = "ASV") %>% 
  left_join(run2_COI_tax) %>% 
  rename_with(.cols = everything(), function(x){paste0("Run2:", x)}) %>% 
  dplyr::rename(Sequence=`Run2:x`) -> run2_CO1_fasta_df

run3_COI_tax <- read.table(here("data","REVAMP_out","Run3_from_poseidon_20250103","COI_REVAMP_out","ASV2Taxonomy","COI_REVAMP_out_asvTaxonomyTable.txt"), sep="\t", header=T) 

run3_COI_fasta <- readDNAStringSet(here("data","REVAMP_out","Run3_from_poseidon_20250103","COI_REVAMP_out","dada2","ASVs.fa"), format="fasta" )

as.data.frame(run3_COI_fasta) %>% 
  as_tibble(rownames = "ASV") %>% 
  left_join(run3_COI_tax) %>% 
   rename_with(.cols = everything(), function(x){paste0("Run3:", x)}) %>% 
  dplyr::rename(Sequence=`Run3:x`)-> run3_COI_fasta_df

full_join(run2_CO1_fasta_df, run3_COI_fasta_df) %>% 
 mutate(., ASV_combo = row_number(),
        ASV_combo = str_c("ASVc_",ASV_combo)) %>% 
 mutate(., Kingdom = if_else(!is.na(`Run3:Kingdom`), `Run3:Kingdom`,`Run2:Kingdom`),
        Phylum= if_else(!is.na(`Run3:Phylum`), `Run3:Phylum`,`Run2:Phylum`), 
        Class= if_else(!is.na(`Run3:Class`), `Run3:Class`,`Run2:Class`),
        Order= if_else(!is.na(`Run3:Order`), `Run3:Order`,`Run2:Order`), 
        Family= if_else(!is.na(`Run3:Family`), `Run3:Family`,`Run2:Family`),
        Genus= if_else(!is.na(`Run3:Genus`), `Run3:Genus`,`Run2:Genus`),
        Species= if_else(!is.na(`Run3:Species`), `Run3:Species`,`Run2:Species`)) %>% 
  dplyr::select(ASV_combo, `Run3:ASV`, `Run2:ASV`, Sequence, Kingdom,Phylum,Class, Order, Family, Genus, Species) -> run2_3_COI_united_ASV_tables

saveRDS(run2_3_COI_united_ASV_tables, here("data","long_data","run2_3_COI_united_ASV_tables.RDS"))

```

## Run 2
```{r}


run2_COI_asv <- read.table(here("data","REVAMP_out","Run2_from_poseidon_20250103","COI","dada2","ASVs_counts.tsv"), sep="\t", header=T) %>% 
  dplyr::rename(ASV=x)

run2_COI_asv %>% 
 dplyr::rename(`Run2:ASV`=ASV) %>% 
  left_join(run2_3_COI_united_ASV_tables) %>% 
  rename_at(vars(starts_with("MP_")), funs(str_replace(., "MP_", ""))) -> run2_COI_merge

run2_COI_merge %>% 
    pivot_longer(cols = c(-`ASV_combo`, -`Run2:ASV`, -`Run3:ASV`,-`Sequence`,-Kingdom, -Phylum, -Class, -Order, -Family, -Genus, -Species), names_to = "Sample", values_to = "nReads") %>% 
  left_join(run1_4_revamp_formatted_metadata) %>% 
  filter(., !is.na(Sample_Name)) %>% # this removes all non-EcoFOCI samples. Note this also excludes positive and negative controls. No decontamination has been performed. We should definitely do this at somepoint, but need a conversation as to what approaches
    group_by(Sample) %>%
    mutate (Tot = sum(nReads),
            Prop.abund = nReads / Tot) %>% 
    group_by (ASV_combo) %>%
    mutate (Colmax = max(Prop.abund),
            eDNA.Index = Prop.abund / Colmax) %>%
  dplyr::select(-Tot,-Colmax) -> run2_COI_long
  
```


## Run 3
```{r}


run3_COI_asv <- read.table(here("data","REVAMP_out","Run3_from_poseidon_20250103","COI_REVAMP_out","dada2","ASVs_counts.tsv"), sep="\t", header=T) %>% 
  dplyr::rename(ASV=x)

run3_COI_asv %>% 
   dplyr::rename(`Run3:ASV`=ASV) %>% 
  left_join(run2_3_COI_united_ASV_tables) %>% 
   rename_at(vars(starts_with("MP_")), funs(str_replace(., "MP_", ""))) -> run3_COI_merge

run3_COI_merge %>% 
    pivot_longer(cols = c(-`ASV_combo`, -`Run2:ASV`, -`Run3:ASV`,-`Sequence`,-Kingdom, -Phylum, -Class, -Order, -Family, -Genus, -Species), names_to = "Sample", values_to = "nReads") %>% 
  left_join(run1_4_revamp_formatted_metadata) %>% 
  filter(., !is.na(Sample_Name)) %>% # this removes all non-EcoFOCI samples. Note this also excludes positive and negative controls. No decontamination has been performed. We should definitely do this at somepoint, but need a conversation as to what approaches
    group_by(Sample) %>%
    mutate (Tot = sum(nReads),
            Prop.abund = nReads / Tot) %>% 
    group_by (ASV_combo) %>%
    mutate (Colmax = max(Prop.abund),
            eDNA.Index = Prop.abund / Colmax) %>%
  dplyr::select(-Tot,-Colmax) -> run3_COI_long
  
```

## Merge

```{r}

bind_rows(run2_COI_long, run3_COI_long) -> long_COI_EcoFOCI

saveRDS(long_COI_EcoFOCI, here("data","long_data","long_COI_EcoFOCI.RDS"))

```


# MiFish

## Unite ASVs

```{r}
run2_MiFish_tax <- read.table(here("data","REVAMP_out","Run2_from_poseidon_20250103","MiFish","ASV2Taxonomy","MiFish_RUN1_asvTaxonomyTable.txt"), sep="\t", header=T) #IS IT REALLY CALLED RUN1 when it is RUN2?

run2_MiFish_fasta <- readDNAStringSet(here("data","REVAMP_out","Run2_from_poseidon_20250103","MiFish","dada2","ASVs.fa"), format="fasta" )

as.data.frame(run2_MiFish_fasta) %>%
  as_tibble(rownames = "ASV") %>% 
  left_join(run2_MiFish_tax) %>% 
  rename_with(.cols = everything(), function(x){paste0("Run2:", x)}) %>% 
  dplyr::rename(Sequence=`Run2:x`) -> run2_MiFish_fasta_df

run3_MiFish_tax <- read.table(here("data","REVAMP_out","Run3_from_poseidon_20250103","MiFish_REVAMP_out","ASV2Taxonomy","MiFish_REVAMP_out_asvTaxonomyTable.txt"), sep="\t", header=T) 

run3_MiFish_fasta <- readDNAStringSet(here("data","REVAMP_out","Run3_from_poseidon_20250103","MiFish_REVAMP_out","dada2","ASVs.fa"), format="fasta" )

as.data.frame(run3_MiFish_fasta) %>% 
  as_tibble(rownames = "ASV") %>% 
  left_join(run3_MiFish_tax) %>% 
   rename_with(.cols = everything(), function(x){paste0("Run3:", x)}) %>% 
  dplyr::rename(Sequence=`Run3:x`)-> run3_MiFish_fasta_df

full_join(run2_MiFish_fasta_df, run3_MiFish_fasta_df) %>% 
 mutate(., ASV_combo = row_number(),
        ASV_combo = str_c("ASVc_",ASV_combo)) %>% 
 mutate(., Kingdom = if_else(!is.na(`Run3:Kingdom`), `Run3:Kingdom`,`Run2:Kingdom`),
        Phylum= if_else(!is.na(`Run3:Phylum`), `Run3:Phylum`,`Run2:Phylum`), 
        Class= if_else(!is.na(`Run3:Class`), `Run3:Class`,`Run2:Class`),
        Order= if_else(!is.na(`Run3:Order`), `Run3:Order`,`Run2:Order`), 
        Family= if_else(!is.na(`Run3:Family`), `Run3:Family`,`Run2:Family`),
        Genus= if_else(!is.na(`Run3:Genus`), `Run3:Genus`,`Run2:Genus`),
        Species= if_else(!is.na(`Run3:Species`), `Run3:Species`,`Run2:Species`)) %>% 
  dplyr::select(ASV_combo, `Run3:ASV`, `Run2:ASV`, Sequence, Kingdom,Phylum,Class, Order, Family, Genus, Species) -> run2_3_MiFish_united_ASV_tables

saveRDS(run2_3_MiFish_united_ASV_tables, here("data","long_data","run2_3_MiFish_united_ASV_tables.RDS"))

```

## Run 2
```{r}


run2_MiFish_asv <- read.table(here("data","REVAMP_out","Run2_from_poseidon_20250103","MiFish","dada2","ASVs_counts.tsv"), sep="\t", header=T) %>% 
  dplyr::rename(ASV=x)

run2_MiFish_asv %>% 
dplyr::rename(`Run2:ASV`=ASV) %>% 
  left_join(run2_3_MiFish_united_ASV_tables) %>% 
  rename_at(vars(starts_with("MP_")), funs(str_replace(., "MP_", ""))) -> run2_MiFish_merge

run2_MiFish_merge %>% 
    pivot_longer(cols = c(-`ASV_combo`, -`Run2:ASV`, -`Run3:ASV`,-`Sequence`,-Kingdom, -Phylum, -Class, -Order, -Family, -Genus, -Species), names_to = "Sample", values_to = "nReads") %>% 
  left_join(run1_4_revamp_formatted_metadata) %>% 
  filter(., !is.na(Sample_Name)) %>% # this removes all non-EcoFOCI samples. Note this also excludes positive and negative controls. No decontamination has been performed. We should definitely do this at somepoint, but need a conversation as to what approaches
    group_by(Sample) %>%
    mutate (Tot = sum(nReads),
            Prop.abund = nReads / Tot) %>% 
    group_by (ASV_combo) %>%
    mutate (Colmax = max(Prop.abund),
            eDNA.Index = Prop.abund / Colmax) %>%
  dplyr::select(-Tot,-Colmax)-> run2_MiFish_long
  
```


## Run 3
```{r}


run3_MiFish_asv <- read.table(here("data","REVAMP_out","Run3_from_poseidon_20250103","MiFish_REVAMP_out","dada2","ASVs_counts.tsv"), sep="\t", header=T) %>% 
  dplyr::rename(ASV=x)

run3_MiFish_asv %>% 
    dplyr::rename(`Run3:ASV`=ASV) %>% 
  left_join(run2_3_MiFish_united_ASV_tables) %>% 
   rename_at(vars(starts_with("MP_")), funs(str_replace(., "MP_", ""))) -> run3_MiFish_merge

run3_MiFish_merge %>% 
    pivot_longer(cols = c(-`ASV_combo`, -`Run2:ASV`, -`Run3:ASV`,-`Sequence`,-Kingdom, -Phylum, -Class, -Order, -Family, -Genus, -Species), names_to = "Sample", values_to = "nReads") %>% 
  left_join(run1_4_revamp_formatted_metadata) %>% 
  filter(., !is.na(Sample_Name)) %>% # this removes all non-EcoFOCI samples. Note this also excludes positive and negative controls. No decontamination has been performed. We should definitely do this at somepoint, but need a conversation as to what approaches
    group_by(Sample) %>%
    mutate (Tot = sum(nReads),
            Prop.abund = nReads / Tot) %>% 
    group_by (ASV_combo) %>%
    mutate (Colmax = max(Prop.abund),
            eDNA.Index = Prop.abund / Colmax) %>%
  dplyr::select(-Tot,-Colmax) -> run3_MiFish_long
  
  
```

## Merge

```{r}

bind_rows(run2_MiFish_long, run3_MiFish_long) -> long_MiFish_EcoFOCI

saveRDS(long_MiFish_EcoFOCI, here("data","long_data","long_MiFish_EcoFOCI.RDS"))

```
