---
title: "Arctic Metadata Merging"
author: "Zack Gold"
date: "2024-07-01"
output: html_document
---

```{r}
library(tidyverse)
library(readxl)
library(here)
library(readr)
```

```{r}
# Load sample metadata - excel sheet with sample name and run names?
revamp_format_run1_4 <- read_excel(here("data","20241218_analysis","sample_metadata_smm_20241218.xlsx"))

revamp_format_run1_4
```

# Load Data

```{r}

FinalOME_Merge_DY20_12 <- read.csv(file = here("data","1_SampleCollection","2020_DY20-12_CruiseData","2020_Dyson_CruiseData","FinalOME_Merge_DY20-12.csv"), header=T) 

FinalOME_Merge_NO20_01 <-  read.csv(here("data","1_SampleCollection","2020_NO20-01_CruiseData","FinalOME_Merge_NO20-01.csv"), header=T)


FinalOME_Merge_SKQ21_12S <- read.csv(file = here("data","1_SampleCollection","2021_SikuliaqDBO_CruiseData","2021_SikuliaqDBO_CruiseData","FinalOME_Merge_SKQ21-12S.csv"), header=T) 


FinalOME_Merge_DY22_09 <- read.csv(here("data","1_SampleCollection","2022_DY22-09_CruiseData","FinalOME_Merge_DY22-09.csv"), header=T) 

FinalOME_Merge_DY22_06 <- read.csv(here("data","1_SampleCollection","2022_DY22-06_CruiseData","FinalOME_Merge_DY22-06.csv"), header=T)

FinalOME_Merge_RC0083 <- read.csv(here("data","1_SampleCollection","2022_GOA-OA_CruiseData","FinalOME_Merge_RC0083.csv"), header=T) 

FinalOME_Merge_DY23_06 <- read.csv(here("data","1_SampleCollection","2023_EcoFociSpringMooring_CruiseData","FinalOME_Merge_DY23-06.csv"), header=T) 

FinalOME_Merge_M2_PPS_0523 <- read.csv(here("data","1_SampleCollection","2023_M2_Spring-Summer_CruiseData","FinalOME_Merge_M2-PPS-0523.csv"), header=T)

FinalOME_Merge_SKQ23_12S <- read.csv(here("data","1_SampleCollection","2023_SikuliaqDBO_CruiseData","FinalOME_Merge_SKQ23-12S.csv"), header=T) 


FinalOME_Merge_AquaM_23 <- read.csv(here("data","1_SampleCollection","2023_aquamonitor_CruiseData","FinalOME_Merge_AquaM_23.csv"), header=T) 

FinalOME_Merge_RC0083
```

```{r}
# Gets list of unique column names from all the final merge dfs.
c(colnames(FinalOME_Merge_DY20_12), 
  colnames(FinalOME_Merge_NO20_01),
  colnames(FinalOME_Merge_SKQ21_12S),
  colnames(FinalOME_Merge_DY22_09), 
  colnames(FinalOME_Merge_DY22_06), 
  colnames(FinalOME_Merge_RC0083), 
  colnames(FinalOME_Merge_DY23_06), 
  colnames(FinalOME_Merge_M2_PPS_0523), 
  colnames(FinalOME_Merge_SKQ23_12S),
  colnames(FinalOME_Merge_AquaM_23)) %>% unique() -> all_unique

all_unique
```
# Function to Convert wind directions to main directions
```{r}
# %% 360 ensure any degree value wraps around to say within 0-359. So 370 degrees becomes 10 degrees. Handles any values where degree values might be outside the 0-359 range
# create break points offset by 22.5 to center
# labels() combines values into a vector (array).
# find_interval() determins which interval each degree falls into. 30 falls between 0 and 45 so returns 1
# return(labels[index ]) uses square brackets to access elements in the labels vector. So if index is 1, it will return NE.
convert_degrees_to_cardinal_directions <- function(degrees){
  labels <- c("N", "NNE", "NE", "ENE", "E", "ESE", "SE", "SSE", "S", "SSW", "SW", "WSW", "W", "WNW", "NW", "NNW")
  degree_ranges <- seq(0, 360, by=22.5)
  index <- findInterval(degrees, degree_ranges, rightmost.closed = TRUE)
  return(labels[index])
}
```

#Check Salt

```{r}
# converts all_unique list to a df (as.tibble) so that all the column names from all the merged datasets are a row in the column called "value". Then it creates a df called tibble_of_columns that creates a column for each final merged dataset (eg. SKQ21) with values of the column names each has. It checks if the "value" (aka column name) is in each dataset and if it is, it writes the column name to the row, and if its not it is NA. tibble_of_columns is a df to view which columns are present in which datasets.
all_unique %>%  as.tibble() %>%
  mutate(., SKQ21 = if_else(value %in% colnames(FinalOME_Merge_SKQ21_12S), value, NA),
         DY2020 = if_else(value %in% colnames(FinalOME_Merge_DY20_12), value, NA),
         DY2206_long = if_else(value %in% colnames(FinalOME_Merge_DY22_06), value, NA),
         DY2209_long = if_else(value %in% colnames(FinalOME_Merge_DY22_09), value, NA),
         Norseman2020_long = if_else(value %in% colnames(FinalOME_Merge_NO20_01), value, NA),
         RCGOA2022_long = if_else(value %in% colnames(FinalOME_Merge_RC0083), value, NA),
         M2_PPS_0523 = if_else(value %in% colnames(FinalOME_Merge_M2_PPS_0523), value, NA),
         DY23_06 = if_else(value %in% colnames(FinalOME_Merge_DY23_06), value, NA),
         SKQ202312S_long = if_else(value %in% colnames(FinalOME_Merge_SKQ23_12S), value, NA),
         AquaM_23 = if_else(value %in% colnames(FinalOME_Merge_AquaM_23), value, NA) ) -> tibble_of_columns

tibble_of_columns
```

We would like to keep temperature, salinity, oxygen, chlorphyll, SI, PO4, NO2, NO3. NH4, pH

#Check Temp

```{r}
# Filters tibble_of_columns df to get only rows that have a string match to those listed
tibble_of_columns %>% 
  filter(., str_detect(value,"temperature|temp|T090C|T190C|Temp|Temperature"))

# Ranking Temp ctd_temperature_ch1..degree_C., btl_temperature_ch1..degree_C., btl_T090C


```

# Salinity

```{r}
# filters tibble_of_columns df for rows with values that have a str match of sal or Sal
tibble_of_columns %>% 
  filter(., str_detect(value,"sal|Sal"))

# Ranking "ctd_salinity_ch1..PSU.", "btl_Sal00",  "btl_Salinity.psu.", "btl_salinity_ch1..PSU."

# select all data from values in the Norseman2020_long column and filter out any values that are NA. So get only rows with columns that are present in the Noresman dataset.
tibble_of_columns %>% 
  select(value, Norseman2020_long) %>% 
  filter(., !is.na(Norseman2020_long))
```

# Oxygen

```{r}

# filter for columns that have string matches below 
tibble_of_columns %>% 
  filter(., str_detect(value,"o2|ox|oxy|Oxy|Ox|O2"))

# Ranking Percent "ctd_oxy_percentsat_ch1..percent.", "btl_Sbeox0PS", "btl_oxy_percentsat_ch1..percent."
# Ranking Conc. "ctd_oxy_conc_ch1..umol.kg.", "btl_Sbox0Mm.Kg","btl_oxy_conc_ch1..micromole.kg."

```

#Check Chlorophyll

```{r}
# filter for columns with string matches in thr rows listed below
tibble_of_columns %>% 
  filter(., str_detect(value,"chl|par|chlor|Fluorescence|fluor|PAR|chlorophyll|FlECO.AFL"))

# Ranking Temp "ctd_chlor_fluorescence..ugrams.l."	, "btl_bottle.chlorophyll.µg.L.1", "btl_chlor_fluorescence..ugrams.l.", "btl_FlECO.AFL"	(this is in [mg/m^3] but is the same unit)
```

#Check PO4

```{r}
# String match for phosphorus related row string matches
tibble_of_columns %>% 
  filter(., str_detect(value,"PO4|phosphate|phos|Phosphate|Phos"))

# Ranking Temp "ctd_PO4..uM.l.", btl_phosphate.µM, btl2_Phosphate.umol.kg.1	(equivalent units)

# select only rows for the DY23_06 column with a value that is not NA
tibble_of_columns %>% 
  select(value, DY23_06) %>% 
  filter(., !is.na(DY23_06))
```

#Check NO2

```{r}
# filter for any row string matches for nitrogen related strings
tibble_of_columns %>% 
  filter(., str_detect(value,"NO2|Nitrite|nitrite|no2"))

# Ranking Temp "ctd_NO2..uM.l.", "btl_nitrate.nitriteµM" *need more info on this, "btl2_Nitrite.umol.kg.1"

# select only the DY23_06 dataset column and filter out an NA rows
tibble_of_columns %>% 
  select(value, DY23_06) %>% 
  filter(., !is.na(DY23_06))

# Access btl_nitrate.nitriteµM column values from the FinalOME_Merge_NO20_01 df
FinalOME_Merge_NO20_01$btl_nitrate.nitriteµM
```

#Check NO3

```{r}
# filter for nitrogen related strings in the rows
tibble_of_columns %>% 
  filter(., str_detect(value,"NO3|Nitrate|nitrate|no3"))

# Ranking Temp "	ctd_NO3..uM.l.", "btl_nitrate.nitriteµM" *need more info on this, "btl2_Nitrate.umol.kg.1	"

```

#Check NH4

```{r}
# filter for any ammonia related strings in the row values
tibble_of_columns %>% 
  filter(., str_detect(value,"NH4|ammonia|ammonium|nh4"))

# Ranking Temp "ctd_NH4..uM.l.", "btl_ammonia"

```

#Check pH
```{r}
# filter for any pH related strings in the row values but disallow characters after the "h" (to avoid phosphate returning.)
tibble_of_columns %>%
  filter(., str_detect(value,"[//./_][pP][hH](?![a-zA-Z])"))
```


# Check SI
```{r}
tibble_of_columns %>%
  filter(., str_detect(value, "[Ss]ilicate"))
```

# Check inorganic nitrogen
```{r}
tibble_of_columns %>%
  filter(., str_detect(value, "inroganic.N"))
```

# Check Turbidity
```{r}
tibble_of_columns %>%
  filter(., str_detect(value, "[Tt]urb"))
```


# check wind_direction and wind_speed
```{r}
tibble_of_columns %>%
  filter(., str_detect(value, "[Ww]ind"))
```


# Clean for REVAMP

## Have EcoFOCI CTD Data

SKQ21, DY2020, DY2206_long, DY2209_long, SKQ202312S_long

### REVAMP_DY20_12

```{r}

# select the specified columns from the Final_OME_Merge_DY20_12 dataset, and rename some of the columns. (e.g. rename the column ctd_pressure to Depth) and save as new df called REVAMP_DY20_12
FinalOME_Merge_DY20_12 %>% 
  mutate(., pH = NA,
            pH_method = NA,
            Silicate_umol = NA,
            Silicate_umol.per.kg = NA,
            Tot_Inorganic_N.umol = NA,
            wind_dirction = convert_degrees_to_cardinal_directions(ctd_wind_direction..degrees.)) %>%
  dplyr::select(Sample_Name, Biological_Replicate, Technical_Replicate, Negative_control, Cruise_ID_short, Cruise_ID_long, Cast_No.,Rosette_position, Station, Sample_volume_ml, Field_collection_method, Preservation, pH, pH_method, Silicate_umol, Silicate_umol.per.kg, Tot_Inorganic_N.umol, wind_dirction,
                Time=ctd_time..UTC., 
                Depth_m=ctd_pressure, 
                lat=ctd_latitude..degrees_north., 
                lon = ctd_longitude..degrees_east., 
                Temp_C= ctd_temperature_ch1..degree_C.,
                PSU_ppt = ctd_salinity_ch1..PSU., 
                OxySatPerc_percent = ctd_oxy_percentsat_ch1..percent.,
                OxyConc_umol.per.l = ctd_oxy_conc_ch1..umol.kg.,
                Chla_ugrams.per.l = ctd_chlor_fluorescence..ugrams.l.,
                PO4_umol.per.l = ctd_PO4..uM.l.,
                NO2_umol.per.l=ctd_NO2..uM.l.,
                NO3_umol.per.l=ctd_NO3..uM.l.,
                NH4_umol.per.l =ctd_NH4..uM.l.,
                # wind_direction.degrees = ctd_wind_direction..degrees.,
                wind_speed_m.per.s = ctd_wind_speed..m.s.1.,
                turbidity.NTU = btl_TurbWETntu0) -> REVAMP_DY20_12

REVAMP_DY20_12
```

### REVAMP_SKQ21_12S

```{r}

# edits the FinalOME_Merge_SKQ21_12S df columns like ctd_latitude..degrees_north (if the Cast_No. column values =1 then it makes the value in in the ctd_latitude..degreses_north the Lat_Notes, if not, it keeps the existing value). Does similar things to other columns. It does other data cleaning in the ctd time columns. Then it selects all the desired columns, changing some column names (e.g. ctd_pressure becomes Depth) and saves the new df as REVAMP_SKQ21_12S

  FinalOME_Merge_SKQ21_12S %>% 
  #For some reason we are missing the first bottle file for this cruise... so taking time and lat/lon data from notes
  mutate(., ctd_latitude..degrees_north. = if_else(Cast_No.==1, Lat_notes,ctd_latitude..degrees_north.)) %>% 
    mutate(., ctd_longitude..degrees_east. = if_else(Cast_No.==1, Lon_notes,ctd_longitude..degrees_east.)) %>% 
  mutate(., ctd_pressure = if_else(Cast_No.==1, Depth_m_notes,ctd_pressure)) %>% 
    mutate(., ctd_time..UTC. = if_else(Cast_No.==1, Date_UTC_notes,ctd_time..UTC.)) %>% 
      mutate(., ctd_time..UTC. = str_replace_all(ctd_time..UTC.,"\\/", "-")) %>% 
      mutate(., ctd_time..UTC. = str_replace_all(ctd_time..UTC.,"08 22", "08T22")) %>% 
  mutate(ctd_time..UTC. = if_else(Cast_No.==1, str_c(ctd_time..UTC.,"Z"),ctd_time..UTC.)) %>% 
  mutate(., pH = NA,
            pH_method = NA,
            Silicate_umol = NA,
            Silicate_umol.per.kg = NA,
            Tot_Inorganic_N.umol = NA,
            wind_direction = convert_degrees_to_cardinal_directions(ctd_wind_direction..degrees.)) %>%
  dplyr::select(Sample_Name, Biological_Replicate, Technical_Replicate, Negative_control, Cruise_ID_short, Cruise_ID_long, Cast_No.,Rosette_position, Station, Sample_volume_ml, Field_collection_method, Preservation, pH, pH_method, Silicate_umol, Silicate_umol.per.kg, Tot_Inorganic_N.umol, wind_direction,
                Time=ctd_time..UTC., 
                Depth_m=ctd_pressure, 
                lat=ctd_latitude..degrees_north., 
                lon = ctd_longitude..degrees_east., 
                Temp_C= ctd_temperature_ch1..degree_C.,
                PSU_ppt = ctd_salinity_ch1..PSU., 
                OxySatPerc_percent = ctd_oxy_percentsat_ch1..percent.,
                OxyConc_umol.per.l = ctd_oxy_conc_ch1..umol.kg.,
                Chla_ugrams.per.l = ctd_chlor_fluorescence..ugrams.l.,
                PO4_umol.per.l = ctd_PO4..uM.l.,
                NO2_umol.per.l=ctd_NO2..uM.l.,
                NO3_umol.per.l=ctd_NO3..uM.l.,
                NH4_umol.per.l =ctd_NH4..uM.l.,
                turbidity.NTU = btl_turbidity) -> REVAMP_SKQ21_12S
```

### REVAMP_DY22_09

```{r}
# Similar editing as other revamp dfs above.
  FinalOME_Merge_DY22_09 %>% 
    mutate(., pH = NA,
              pH_method = NA,
              Silicate_umol = NA,
              Silicate_umol.per.kg = NA,
              Tot_Inorganic_N.umol = NA) %>%
  dplyr::select(Sample_Name, Biological_Replicate, Technical_Replicate, Negative_control, Cruise_ID_short, Cruise_ID_long, Cast_No.,Rosette_position, Station, Sample_volume_ml, Field_collection_method, Preservation, pH, pH_method, Silicate_umol, Silicate_umol.per.kg, Tot_Inorganic_N.umol,
                Time=ctd_time..UTC., 
                Depth_m=ctd_pressure, 
                lat=ctd_latitude..degrees_north., 
                lon = ctd_longitude..degrees_east., 
                Temp_C= ctd_temperature_ch1..degree_C.,
                PSU_ppt = ctd_salinity_ch1..PSU., 
                OxySatPerc_percent = ctd_oxy_percentsat_ch1..percent.,
                OxyConc_umol.per.l = ctd_oxy_conc_ch1..umol.kg.,
                Chla_ugrams.per.l = ctd_chlor_fluorescence..ugrams.l.,
                PO4_umol.per.l = ctd_PO4..uM.l.,
                NO2_umol.per.l=ctd_NO2..uM.l.,
                NO3_umol.per.l=ctd_NO3..uM.l.,
                NH4_umol.per.l =ctd_NH4..uM.l.,
                wind_direction.degrees = ctd_wind_direction..degrees.,
                wind_speed_m.per.s = ctd_wind_speed..m.s.1.,
                turbidity.NTU = btl_turbidity..NTU.) -> REVAMP_DY22_09
```

### REVAMP_SKQ23_12S

```{r}
# Similar editing as before. Edits some columns that contain na as their values to be replaced with corresponding notes column values. If they aren't na, the values stay the same. More data cleaning to time columns. Selects the desired columns, renaming some columns and saving the new df as REVAMP_SKQ23_12S
  FinalOME_Merge_SKQ23_12S %>% 
    mutate(., ctd_latitude..degrees_north. = if_else(is.na(ctd_latitude..degrees_north.), Lat_notes,ctd_latitude..degrees_north.)) %>% 
    mutate(., ctd_longitude..degrees_east. = if_else(is.na(ctd_longitude..degrees_east.), Lon_notes,ctd_longitude..degrees_east.)) %>% 
  mutate(., ctd_pressure = if_else(is.na(ctd_pressure), Depth_m_notes,ctd_pressure)) %>% 
    mutate(., ctd_time..UTC. = if_else(is.na(ctd_time..UTC.), str_c(Date_UTC_notes,"Z"),ctd_time..UTC.)) %>% 
      mutate(., ctd_time..UTC. = str_replace_all(ctd_time..UTC.,"\\/", "-")) %>% 
      mutate(., ctd_time..UTC. = str_replace_all(ctd_time..UTC.," ", "T")) %>% 
      mutate(., pH = NA,
              pH_method = NA,
              Silicate_umol = NA,
              Silicate_umol.per.kg = NA,
              Tot_Inorganic_N.umol = NA,
              turbidity.NTU = NA) %>%
  dplyr::select(Sample_Name, Biological_Replicate, Technical_Replicate, Negative_control, Cruise_ID_short, Cruise_ID_long, Cast_No.,Rosette_position, Station, Sample_volume_ml, Field_collection_method, Preservation, pH, pH_method, Silicate_umol, Silicate_umol.per.kg, Tot_Inorganic_N.umol, turbidity.NTU,
                Time=ctd_time..UTC., 
                Depth_m=ctd_pressure, 
                lat=ctd_latitude..degrees_north., 
                lon = ctd_longitude..degrees_east., 
                Temp_C= ctd_temperature_ch1..degree_C.,
                PSU_ppt = ctd_salinity_ch1..PSU., 
                OxySatPerc_percent = ctd_oxy_percentsat_ch1..percent.,
                OxyConc_umol.per.l = ctd_oxy_conc_ch1..umol.kg.,
                Chla_ugrams.per.l = ctd_chlor_fluorescence..ugrams.l.,
                PO4_umol.per.l = ctd_PO4..uM.l.,
                NO2_umol.per.l=ctd_NO2..uM.l.,
                NO3_umol.per.l=ctd_NO3..uM.l.,
                NH4_umol.per.l =ctd_NH4..uM.l.,
                wind_direction.degrees = ctd_wind_direction..degrees.,
                wind_speed_m.per.s = ctd_wind_speed..m.s.1.) -> REVAMP_SKQ23_12S


```

### REVAMP_DY22_06 - Unique

```{r}
# Similar editing as above. New df is saved as REVAMP_DY22_06
  FinalOME_Merge_DY22_06 %>% 
   mutate(., ctd_latitude..degrees_north. = if_else(is.na(ctd_latitude..degrees_north.), btl_latitude..degrees_north.,ctd_latitude..degrees_north.)) %>% 
    mutate(., ctd_longitude..degrees_east. = if_else(is.na(ctd_longitude..degrees_east.), btl_longitude..degrees_east.,ctd_longitude..degrees_east.)) %>% 
    mutate(., ctd_pressure = if_else(is.na(ctd_pressure), btl_pressure..decibar.,ctd_pressure)) %>% 
  mutate(., ctd_pressure = if_else(is.na(ctd_pressure), Depth_m_notes,ctd_pressure)) %>% 
    mutate(., ctd_time..UTC. = if_else(is.na(ctd_time..UTC.), btl_time..UTC.,ctd_time..UTC.)) %>% 
    mutate(., pH = NA,
              pH_method = NA,
              Silicate_umol = NA,
              Silicate_umol.per.kg = NA,
              Tot_Inorganic_N.umol = NA) %>% 
  dplyr::select(Sample_Name, Biological_Replicate, Technical_Replicate, Negative_control, Cruise_ID_short, Cruise_ID_long, Cast_No.,Rosette_position, Station, Sample_volume_ml, Field_collection_method, Preservation, pH, pH_method, Silicate_umol, Silicate_umol.per.kg, Tot_Inorganic_N.umol,
                Time=ctd_time..UTC., 
                Depth_m=ctd_pressure, 
                lat=ctd_latitude..degrees_north., 
                lon = ctd_longitude..degrees_east., 
                Temp_C= btl_temperature_ch1..degree_C.,
                PSU_ppt = btl_salinity_ch1..PSU., 
                OxySatPerc_percent = btl_oxy_percentsat_ch1..percent.,
                OxyConc_umol.per.l = btl_oxy_conc_ch1..micromole.kg.,
                Chla_ugrams.per.l = btl_chlor_fluorescence..ugrams.l.,
                PO4_umol.per.l = ctd_PO4..uM.l.,
                NO2_umol.per.l=ctd_NO2..uM.l.,
                NO3_umol.per.l=ctd_NO3..uM.l.,
                NH4_umol.per.l =ctd_NH4..uM.l.,
                wind_direction.degrees = ctd_wind_direction..degrees.,
                wind_speed_m.per.s = ctd_wind_speed..m.s.1.,
                turbidity.NTU = btl_turbidity..NTU.) -> REVAMP_DY22_06
```

## Unique Cruise File Formats

### REVAMP_NO20_01

```{r}
# Similar editing as above except it also creates a new column called OxySatPerc_percent and sets all values to NA. Selects only desired columns and saves new df as REVAMP_NO20_01
# Norseman2020_long
    FinalOME_Merge_NO20_01 %>% 
        mutate(., ctd_time..UTC. = str_replace_all(Date_UTC_btl,"\\/", "-")) %>% 
      mutate(., ctd_time..UTC. = str_replace_all(ctd_time..UTC.," ", "T")) %>% 
  mutate(ctd_time..UTC. = str_c(ctd_time..UTC.,"Z"),ctd_time..UTC.) %>% 
  mutate(., OxySatPerc_percent= NA,
            pH = NA,
            pH_method = NA,
            Silicate_umol.per.kg = NA,
            wind_direction.degrees = NA,
            wind_speed_m.per.s = NA,
            turbidity.NTU = NA) %>% 
   dplyr::select(Sample_Name, Biological_Replicate, Technical_Replicate, Negative_control, Cruise_ID_short, Cruise_ID_long, Cast_No.,Rosette_position, Station, Sample_volume_ml, Field_collection_method, Preservation, pH, pH_method, Silicate_umol.per.kg, wind_direction.degrees, wind_speed_m.per.s, turbidity.NTU,
                Time=ctd_time..UTC., 
                Depth_m=btl_Pressure.db., 
                lat=btl_Latitude, 
                lon = btl_Longitude, 
                Temp_C= btl_Temperature.deg.C.,
                PSU_ppt = btl_Salinity.psu., 
                OxySatPerc_percent = OxySatPerc_percent,
                OxyConc_umol.per.l = btl_Oxygen.umol.kg.,
                Chla_ugrams.per.l = btl_Fluorescence.mg.m.3.,
                PO4_umol.per.l = btl_phosphate.µM,
                NO2_umol.per.l=btl_nitrate.nitriteµM,
                NO3_umol.per.l=btl_nitrate.nitriteµM,
                NH4_umol.per.l =btl_ammonia,
                Silicate_umol = btl_silicate.µM,
                Tot_Inorganic_N.umol = btl_total.inroganic.N) -> REVAMP_NO20_01
    
```

### REVAMP_RC0083

```{r}
#RCGOA2022_long
 #  carbonate chemistry not included since no other cruise has that data as of 12/26/2024 
  FinalOME_Merge_RC0083  %>% 
        mutate(., ctd_time..UTC. = str_replace_all(Date_UTC_btl,"\\/", "-")) %>% 
      mutate(., ctd_time..UTC. = str_replace_all(ctd_time..UTC.," ", "T")) %>% 
  mutate(ctd_time..UTC. = str_c(ctd_time..UTC.,"Z"),ctd_time..UTC.) %>% 
  mutate(., pH_method= "SBE-18 pH sensor",
            Silicate_umol = NA,
            wind_direction.degrees = NA,
            wind_speed_m.per.s = NA,
            Tot_Inorganic_N.umol = NA,
            turbidity.NTU = NA) %>% 
   dplyr::select(Sample_Name, Biological_Replicate, Technical_Replicate, Negative_control, Cruise_ID_short, Cruise_ID_long, Cast_No.,Rosette_position, Station, Sample_volume_ml, Field_collection_method, Preservation, pH_method, Silicate_umol, wind_direction.degrees, wind_speed_m.per.s, Tot_Inorganic_N.umol, turbidity.NTU,
                Time=ctd_time..UTC., 
                Depth_m=btl2_CTDPRES, 
                lat=btl2_Latitude, 
                lon = btl2_Longitude, 
                Temp_C= btl2_CTDTEMP_ITS90,
                PSU_ppt = btl2_CTDSAL_PSS78, 
                OxySatPerc_percent = btl_Sbeox0PS,
                OxyConc_umol.per.l = btl_Sbox0Mm.Kg,
                Chla_ugrams.per.l = btl_FlECO.AFL,
                PO4_umol.per.l = btl2_Phosphate.umol.kg.1,
                NO2_umol.per.l=btl2_Nitrite.umol.kg.1,
                NO3_umol.per.l=btl2_Nitrate.umol.kg.1,
                NH4_umol.per.l =btl2_Ammonium.umol.kg.1,
                pH = btl2_SBE.18.pH..BA.,
                Silicate_umol.per.kg = btl2_Silicate.umol.kg.1) -> REVAMP_RC0083 

REVAMP_RC0083
```

### REVAMP_DY23_06

```{r}

#This is also missing a bunch of cast data, Shaun said he would send the processed, raw data is missing casts for whatever reason
#DY23_06
      FinalOME_Merge_DY23_06 %>% 
   mutate(., btl_latitude..degrees_north. = if_else(is.na(btl_time..UTC.), Lat_notes,btl_latitude..degrees_north.)) %>% 
    mutate(., btl_longitude..degrees_east. = if_else(is.na(btl_time..UTC.), Lon_notes,btl_longitude..degrees_east.)) %>% 
  mutate(., btl_pressure..decibar. = if_else(is.na(btl_time..UTC.), Depth_m_notes,btl_pressure..decibar.)) %>% 
    mutate(., btl_time..UTC. = if_else(is.na(btl_time..UTC.), str_c(Date_UTC_notes,"Z"),btl_time..UTC.)) %>% 
      mutate(., btl_time..UTC. = str_replace_all(btl_time..UTC.,"\\/", "-")) %>% 
      mutate(., btl_time..UTC. = str_replace_all(btl_time..UTC.," ", "T")) %>% 
  mutate(.,PO4_umol.per.l = NA,
                NO2_umol.per.l=NA,
                NO3_umol.per.l=NA,
                NH4_umol.per.l =NA,
                pH = NA,
                pH_method = NA,
                Silicate_umol = NA,
                Silicate_umol.per.kg = NA,
                wind_direction.degrees = NA,
                wind_speed_m.per.s = NA,
                Tot_Inorganic_N.umol = NA)%>% 
   dplyr::select(Sample_Name, Biological_Replicate, Technical_Replicate, Negative_control, Cruise_ID_short, Cruise_ID_long, Cast_No.,Rosette_position, Station, Sample_volume_ml, Field_collection_method, Preservation, pH, pH_method, Silicate_umol, Silicate_umol.per.kg, wind_direction.degrees, wind_speed_m.per.s, Tot_Inorganic_N.umol,
                Time=btl_time..UTC., 
                Depth_m=btl_pressure..decibar., 
                lat=btl_latitude..degrees_north., 
                lon = btl_longitude..degrees_east., 
                Temp_C= btl_temperature_ch2..degree_C.,
                PSU_ppt = btl_salinity_ch1..PSU., 
                OxySatPerc_percent = btl_oxy_percentsat_ch1..percent.,
                OxyConc_umol.per.l = btl_oxy_conc_ch1..micromole.kg.,
                Chla_ugrams.per.l = btl_chlor_fluorescence..ugrams.l.,
                PO4_umol.per.l = PO4_umol.per.l,
                NO2_umol.per.l=NO2_umol.per.l,
                NO3_umol.per.l=NO3_umol.per.l,
                NH4_umol.per.l =NH4_umol.per.l,
                turbidity.NTU = btl_turbidity..NTU.) -> REVAMP_DY23_06

```

### REVAMP PPS_0523

```{r}
  FinalOME_Merge_M2_PPS_0523 %>% 
          mutate(., ctd_time..UTC. = str_replace_all(Date_UTC_notes,"\\/", "-")) %>% 
      mutate(., ctd_time..UTC. = str_replace_all(ctd_time..UTC.," ", "T")) %>% 
  mutate(ctd_time..UTC. = str_c(ctd_time..UTC.,"Z"),ctd_time..UTC.) %>% 
    mutate(., 
           Temp_C= NA,
                PSU_ppt = NA, 
                OxySatPerc_percent = NA,
                OxyConc_umol.per.l = NA,
                Chla_ugrams.per.l = NA,
                PO4_umol.per.l = NA,
                NO2_umol.per.l=NA,
                NO3_umol.per.l=NA,
                NH4_umol.per.l =NA,
                pH = NA,
                pH_method = NA,
                Silicate_umol = NA,
                Silicate_umol.per.kg = NA,
                wind_direction.degrees = NA,
                wind_speed_m.per.s = NA,
                Tot_Inorganic_N.umol = NA,
                turbidity.NTU = NA)%>% 
  dplyr::select(Sample_Name, Biological_Replicate, Technical_Replicate, Negative_control, Cruise_ID_short, Cruise_ID_long, Cast_No.,Rosette_position, Station, Sample_volume_ml, Field_collection_method, Preservation, pH, pH_method, Silicate_umol, Silicate_umol.per.kg, wind_direction.degrees, wind_speed_m.per.s, Tot_Inorganic_N.umol, turbidity.NTU, 
                Time=ctd_time..UTC., 
                Depth_m=Depth_m_notes, 
                lat=Lat_notes, 
                lon = Lon_notes, 
                Temp_C= Temp_C,
                PSU_ppt = PSU_ppt, 
                OxySatPerc_percent = OxySatPerc_percent,
                OxyConc_umol.per.l = OxyConc_umol.per.l,
                Chla_ugrams.per.l = Chla_ugrams.per.l,
                PO4_umol.per.l = PO4_umol.per.l,
                NO2_umol.per.l=NO2_umol.per.l,
                NO3_umol.per.l=NO3_umol.per.l,
                NH4_umol.per.l =NH4_umol.per.l) -> REVAMP_PPS_0523
  
```

### REVAMP PPS_0523

```{r}
  FinalOME_Merge_AquaM_23 %>% 
          mutate(., ctd_time..UTC. = str_replace_all(Date_UTC_notes,"\\/", "-")) %>% 
      mutate(., ctd_time..UTC. = str_replace_all(ctd_time..UTC.," ", "T")) %>% 
  mutate(ctd_time..UTC. = str_c(ctd_time..UTC.,"Z"),ctd_time..UTC.) %>% 
    mutate(., 
           Temp_C= NA,
                PSU_ppt = NA, 
                OxySatPerc_percent = NA,
                OxyConc_umol.per.l = NA,
                Chla_ugrams.per.l = NA,
                PO4_umol.per.l = NA,
                NO2_umol.per.l=NA,
                NO3_umol.per.l=NA,
                NH4_umol.per.l =NA,
                pH = NA,
                pH_method = NA,
                Silicate_umol = NA,
                Silicate_umol.per.kg = NA,
                wind_direction.degrees = NA,
                wind_speed_m.per.s = NA,
                Tot_Inorganic_N.umol = NA,
                turbidity.NTU = NA)%>% 
  dplyr::select(Sample_Name, Biological_Replicate, Technical_Replicate, Negative_control, Cruise_ID_short, Cruise_ID_long, Cast_No.,Rosette_position, Station, Sample_volume_ml, Field_collection_method, Preservation, pH, pH_method, Silicate_umol, Silicate_umol.per.kg, wind_direction.degrees, wind_speed_m.per.s, Tot_Inorganic_N.umol, turbidity.NTU,
                Time=ctd_time..UTC., 
                Depth_m=Depth_m_notes, 
                lat=Lat_notes, 
                lon = Lon_notes, 
                Temp_C= Temp_C,
                PSU_ppt = PSU_ppt, 
                OxySatPerc_percent = OxySatPerc_percent,
                OxyConc_umol.per.l = OxyConc_umol.per.l,
                Chla_ugrams.per.l = Chla_ugrams.per.l,
                PO4_umol.per.l = PO4_umol.per.l,
                NO2_umol.per.l=NO2_umol.per.l,
                NO3_umol.per.l=NO3_umol.per.l,
                NH4_umol.per.l =NH4_umol.per.l) -> REVAMP_AquaM_23
  
```

# Merge REVAMP files

```{r}
rbind(REVAMP_DY20_12, REVAMP_DY22_06, REVAMP_DY22_09, REVAMP_DY23_06, REVAMP_NO20_01, REVAMP_PPS_0523, REVAMP_RC0083, REVAMP_SKQ21_12S, REVAMP_SKQ23_12S, REVAMP_AquaM_23) -> revamp_sample_metadata


revamp_sample_metadata %>% 
  filter(., str_detect(Sample_Name,"E1148"))

revamp_sample_metadata
```

```{r}
revamp_format_run1_4 %>% 
  dplyr::select(-sites, -lat, -long) %>% 
  mutate(., Sample_Name = str_replace(Sample,"_","\\.")) %>% 
  mutate(.,Sample_Name = str_remove(Sample_Name,"_PCR1|_PCR2|_PCR3")) %>% 
    mutate(.,Sample_Name = str_replace(Sample_Name,"_NO20",".NO20")) %>% 
      mutate(.,Sample_Name = str_replace(Sample_Name,"_DY20",".DY2012")) %>% 
        mutate(.,Sample_Name = str_replace(Sample_Name,"E26.1B.DY2012","E26.1B.NC.DY2012")) %>% 
          mutate(.,Sample_Name = str_replace(Sample_Name,"E44.DY20","E44.DY2012")) %>% 
  left_join(revamp_sample_metadata) %>% 
  mutate(., sites = Station) -> run1_4_revamp_formatted_metadata

run1_4_revamp_formatted_metadata %>% 
  filter(., str_detect(Sample, "DY2206"))

readr::write_excel_csv(run1_4_revamp_formatted_metadata, file="run1_4_revamp_formatted_metadata.csv")


```
