---
title: "SKQ21_Alaska_HABs_maps"
author: "Sam Setta"
date: "`r Sys.Date()`"
output: html_document
---

# Goal of Analysis

Plot the distribution of HABs for the Alaska/Arctic region.

Note: Was going to use satellite measurements from ERDDAP, but I actually could not pull data from this region for November (SKQ_Alaska cruise, makes sense bc of sea ice), will just use the measurements that were made on the same cruise.

### Load libraries

```{r warning=FALSE, echo=FALSE}
# load libraries
library("readr")
library("here")
library("sf")
library("devtools")
library("rerddap")
library("lubridate")
library("dplyr")
library("reshape2")
library("ggplot2")
library("vegan")
library("tidyverse")
library("knitr")
library("rnaturalearth")
library("rnaturalearthdata")
library("readxl")
library("NatParksPalettes")
library("RColorBrewer")
library("maps")
library("mapdata")
library("measurements")
library("raster")
library("rasterVis")
library("marmap")
library("phyloseq")
library("microViz")
# BiocManager::install("scatterpie")
library("scatterpie")

# load map data:
# Deactivate s2 (spherical geometry for Alaska/Arctic)
sf::sf_use_s2(FALSE)
world <- ne_countries(scale = "medium", returnclass = "sf")
state_prov <- rnaturalearth::ne_states(c("united states of america", "canada","russia"))

here()
```

### Read in the data and format:

```{r,echo=FALSE, hide=TRUE, warning=FALSE, include=FALSE}
# load metabarcoding data (ITS1, 18Sv4, 18Sv9)
SKQ_18S_ITS_df<-readRDS(here("ITS_vs_18S","JonahVentures_RUN2","SKQ_ITS1_18S_all.rds"))

# read environmental csv data:
SKQ_env<-read_csv(here("SKQ_Alaska_Analysis","SKQ_Alaska_data","SKQ21_sample_data_20240627.csv"))

# read environmental csv data:
# Longitude
SKQ_lon.min<-min(SKQ_env$longitude, na.rm = TRUE)
SKQ_lon.max<-max(SKQ_env$longitude, na.rm = TRUE)
# Latitude
SKQ_lat.min<-min(SKQ_env$latitude, na.rm = TRUE)
SKQ_lat.max<-max(SKQ_env$latitude, na.rm = TRUE)
# start date
SKQ_date.start<-as.Date(min(SKQ_env$Date_local, na.rm = TRUE))
# end date
SKQ_date.end<-as.Date(max(SKQ_env$Date_local, na.rm = TRUE))

```


## Map of samples:

### Chlorophyll a along coastline for samples above 100 m

```{r fig1,echo=FALSE, warning=FALSE, fig.height= 12, fig.width= 8}
# ggplot map
SKQ_env_chla<-
  ggplot(data = state_prov) +
    geom_sf(border=1) +
    geom_point(data = SKQ_env , aes(x = longitude, y = latitude, fill=chlor_fluorescence), 
               size=4, shape=21, color="black") +
    # geom_text(data = WCOA_meta_seq %>% filter(Depth < 100), aes(x = Longitude, y = Latitude, label=Section_ID)) +
    coord_sf(xlim = c(-178,-150), ylim = c(55, 72), expand = FALSE) +
    scale_fill_continuous(high = "#284326", low = "#d4e5d2",
                          guide = guide_colorbar(frame.colour = "black", ticks.colour = "black")) +
    theme_bw() +
    xlab("Longitude") +
    ylab("Latitude") +
    labs(fill=expression('Chlorophyll '~italic(a)~''))
    # geom_text(data = summed_station, aes(x = Longitude, y = Latitude, label=Line_ID)) 

SKQ_env_chla

ggsave(
  file = here::here("SKQ_Alaska_Analysis", "SKQ_Alaska_Figures", "SKQ_Chla_map.pdf"),
  width = 8,
  height = 12, limitsize = FALSE
)
```


## Bathymetric map with transect IDs:

Load in depth data from the region:

```{r fig2, echo=FALSE}
# Try adding depth contours:
ylim <- c(-180,-150)
xlim <- c(50, 78)

depth <- 
  getNOAA.bathy(lat1 = xlim[1], lat2 = xlim[2], lon1 = ylim[1], lon2 = ylim[2], resolution = 1) %>% 
  fortify()
# take a look at depth data:
glimpse(depth)
# take a look at depth map:
depth %>% 
  filter(z <= 0) %>% 
  ggplot() +
  geom_raster(aes(x, y, fill = z))
# depth below sea level:
depth_f <- depth %>% filter(z <= 0)
```


```{r fig3, echo=FALSE, warning=FALSE, fig.height= 12, fig.width= 8}
# environmental data only:
# Phyto_env_data<-sample_data(WCOA_18Sv4_phyto_orig) %>% data.frame()
# For Transect ID's:
# Phyto_env_data_transectID<-Phyto_env_data %>% group_by(Transect_ID) %>%
#   mutate(Lat=min(Latitude), Lon=min(Longitude)) %>%
#  dplyr::select(-c("Latitude","Longitude")) %>%
#  filter(Transect_ID!="Anacapa" & Transect_ID!="CMAP")


# ggplot map
gmap<-ggplot(data = state_prov) +
    geom_sf(border=1, fill="#999999", lwd=0.5) +
    geom_raster(data=depth_f, aes(x, y, fill = z)) +
    geom_point(data = SKQ_env , aes(x = longitude, y = latitude), 
               size=4, shape=21, color="black", fill="white") +
    coord_sf(xlim = c(-178, -150), ylim = c(52, 75), expand=FALSE) +
    theme_bw() +
    xlab("Longitude") +
    ylab("Latitude") +
    theme(axis.text=element_text(size=14, face="bold"),
        axis.title=element_text(size=14, face="bold"),
        legend.title=element_text(size=14),
        legend.text=element_text(size=14),
        legend.key=element_rect(fill=NA),
        text = element_text(size=14))

gmap +
  geom_point(data = SKQ_env , aes(x = longitude, y = latitude, color=chlor_fluorescence), 
               size=3, shape=19) +
  scale_color_continuous(high = "#284326", low = "#d4e5d2",
                          guide = guide_colorbar(frame.colour = "black", ticks.colour = "black")) +
  guides(color=guide_legend(title="Chlorophyll a (ug/L)",reverse=T),
         fill=guide_legend(title="Depth (m)",reverse=T))

ggsave(
  file = here::here("SKQ_Alaska_Analysis", "SKQ_Alaska_Figures", "SKQ_Chla_Bathy_map.pdf"),
  width = 8,
  height = 10, limitsize = FALSE
)
```

Combine environmental data to get map of abundance and chlorophyll a

```{r,echo=FALSE, hide=TRUE, warning=FALSE, include=FALSE}
# Load dataframe with information on phytoplankton vs protist groups:
Protist_phyto_gp<-read_csv(here("ITS_vs_18S","Class_filtering.csv"))
# keep list of only phytoplankton:
Phyto_gp_only<-Protist_phyto_gp %>% filter(phytoplankton=="yes") %>% dplyr::select("Class")

# read environmental csv data:
SKQ_env<-read_csv(here("SKQ_Alaska_Analysis","SKQ_Alaska_data","SKQ21_sample_data_20240627.csv"))
SKQ_env2 <- SKQ_env %>% 
  filter(Sample_Name!="END.NC.SKQ2021" & Sample_Name!="START.NC.SKQ2021") %>%
  rename(Sample_ID=Sample_Name) %>%
  mutate(Sample_ID = str_replace_all(Sample_ID, "\\.", "_")) %>%
  dplyr::select("Sample_ID","latitude","longitude")

# load metabarcoding data (ITS1, 18Sv4, 18Sv9)
SKQ_18S_ITS_df<-readRDS(here("ITS_vs_18S","JonahVentures_RUN2","SKQ_ITS1_18S_all.rds"))

# filter out phytos from dataset:
SKQ_metabar<-SKQ_18S_ITS_df %>%
  dplyr::select(-c("MP_E610_NC","MP_POSITIVE","Feature_ID")) %>%
  # filter out samples that were sequenced for all three regions (ITS & 18S)
  pivot_longer(cols = MP_E1032_OC0821_PCR1:MP_E600_3B_PCR3, names_to = "Sample_ID", values_to = "counts") %>%
  filter(Class %in% Phyto_gp_only$Class) %>%
  mutate(Sample_ID = str_replace_all(Sample_ID, c("MP_"="","_PCR*"=""))) %>%
  filter(Sample_ID %in% SKQ_env2$Sample_ID) %>%
  # Correct Pseudo-nitzschia spelling variations
  mutate_if(is.character, str_replace_all, pattern="Pseudo_nitzschia", replacement = "Pseudo-nitzschia") %>%
  mutate(Species = str_replace_all(Species, " ", "_")) %>%
  mutate(Species = sub("^(([^_]*_){1}[^_]*).*", "\\1", Species),
         pres=if_else(counts>0, 1, 0))

# Presence/absence of Pseudo-nitzschia/Alexandrium/Karenia
SKQ_metabar2<-SKQ_metabar %>%
  filter(Genus %in% c("Alexandrium","Karenia","Pseudo-nitzschia")) %>%
  dplyr::select(c("Species","region","Sample_ID","counts","pres"))

# Join metabarcoding and environment for mapping:
SKQ_env_metabar<-left_join(SKQ_metabar2,SKQ_env2,by="Sample_ID") %>%
  drop_na(Species)

# read in presence as factor:
SKQ_env_metabar$pres <-as.factor(SKQ_env_metabar$pres)

```

## Map of species distribution

### Species presence/absence along transect

#### Karenia brevis

```{r fig4, echo=FALSE, warning=FALSE, fig.height= 12, fig.width= 8}
# save subset of data to plot:
SKQ_env_Karenia_df <- SKQ_env_metabar %>% filter(Species=="Karenia_brevis" & region=="18Sv9")

# ggplot map
SKQ_env_Karenia<-
  ggplot(data = state_prov) +
    geom_sf(border=1) +
    geom_point(data = SKQ_env_Karenia_df %>% filter(pres=="1"), 
               aes(x = longitude, y = latitude, fill=pres), 
               size=4, shape=21, color="black") +
    coord_sf(xlim = c(-178,-150), ylim = c(55, 72), expand = FALSE) +
    scale_fill_manual(breaks = c("0", "1"), 
                       values=c("white", "#bc8a31"),
                       labels=c("Absent","Present")) +
   annotate("text", x=-167.5, y=69, label= "n=6", size=5) + 
   facet_grid(. ~ Species) +
    theme_bw() +
    xlab("Longitude") +
    ylab("Latitude") +
    labs(fill=expression(paste("Presence of ", italic("Karenia brevis")))) +
    # geom_text(data = summed_station, aes(x = Longitude, y = Latitude, label=Line_ID)) 
    theme(axis.text=element_text(size=14, face="bold"),
        axis.title=element_text(size=14, face="bold"),
        legend.position = "none",
        text = element_text(size=14),
        strip.background=element_rect(fill="white"))

SKQ_env_Karenia

ggsave(
  file = here::here("SKQ_Alaska_Analysis", "SKQ_Alaska_Figures", "SKQ_Karenia_map.pdf"),
  width = 8,
  height = 10, limitsize = FALSE
)
```

#### Alexandrium catenella, another HAB species

```{r fig5, echo=FALSE, warning=FALSE, fig.height= 12, fig.width= 8}
# save subset of data to plot:
SKQ_env_alex_df <- SKQ_env_metabar %>% filter(Species=="Alexandrium_catenella" & region=="18Sv4")


# count the number of samples with pseudo-nitzschia:
SKQ_env_alex_df %>% filter(pres=="1") %>% nrow()

# ggplot map
SKQ_env_Alex<-
  ggplot(data = state_prov) +
    geom_sf(border=1) +
    geom_point(data = SKQ_env_alex_df %>% filter(pres=="1"), 
               aes(x = longitude, y = latitude, fill=pres), 
               size=4, shape=21, color="black") +
    coord_sf(xlim = c(-178,-150), ylim = c(55, 72), expand = FALSE) +
    scale_fill_manual(breaks = c("0", "1"), 
                       values=c("white", "#bc8a31"),
                       labels=c("Absent","Present")) +
   annotate("text", x=-167.5, y=69, label= "n=20", size=5) + 
   facet_grid(. ~ Species) +
    theme_bw() +
    xlab("Longitude") +
    ylab("Latitude") +
    # labs(fill=expression(paste("Presence of ", italic("Karenia brevis")))) +
    # geom_text(data = summed_station, aes(x = Longitude, y = Latitude, label=Line_ID)) 
    theme(axis.text=element_text(size=14, face="bold"),
        axis.title=element_text(size=14, face="bold"),
        legend.position = "none",
        text = element_text(size=14),
        strip.background=element_rect(fill="white"))

SKQ_env_Alex

ggsave(
  file = here::here("SKQ_Alaska_Analysis", "SKQ_Alaska_Figures", "SKQ_Alex_map.pdf"),
  width = 8,
  height = 10, limitsize = FALSE
)
```

#### Species presence/absence along transect of Pseudo-nitzschia species:
Pseudo-nitzschia_australis (18Sv9 only)
Pseudo-nitzschia_heimii (18Sv9 only)
Pseudo-nitzschia_delicatissima
Pseudo-nitzschia_pungens
Pseudo-nitzschia_seriata
Pseudo-nitzschia_granii
Pseudo-nitzschia_turgidula
Pseudo-nitzschia_fryxelliana
Pseudo-nitzschia_pseudodelicatissima (NOT IN ANY SAMPLES)


Pseudo-nitzschia_sp.


```{r fig6, echo=FALSE, warning=FALSE, fig.height= 12, fig.width= 8}
# save subset of data to plot:
SKQ_env_Paus_df <- SKQ_env_metabar %>% filter(region=="18Sv9" & Species=="Pseudo-nitzschia_australis")

# count the number of samples with pseudo-nitzschia:
SKQ_env_Paus_df %>% filter(pres=="1") %>% nrow()

# ggplot map
SKQ_env_Paus <- ggplot(data = state_prov) +
    geom_sf(border=1) +
    geom_point(data = SKQ_env_Paus_df %>% filter(pres=="1"), 
               aes(x = longitude, y = latitude, fill=pres), 
               size=4, shape=21, color="black") +
    coord_sf(xlim = c(-178,-150), ylim = c(55, 72), expand = FALSE) +
    scale_fill_manual(breaks = c("0", "1"), 
                       values=c("white", "#bc8a31"),
                       labels=c("Absent","Present")) +
   annotate("text", x=-167.5, y=69, label="n=57", size=5) + 
   facet_grid(. ~ Species) +
    theme_bw() +
    xlab("Longitude") +
    ylab("Latitude") +
    theme(axis.text=element_text(size=14, face="bold"),
        axis.title=element_text(size=14, face="bold"),
        legend.position = "none",
        text = element_text(size=14),
        strip.background=element_rect(fill="white"))

SKQ_env_Paus

ggsave(
  file = here::here("SKQ_Alaska_Analysis", "SKQ_Alaska_Figures", "SKQ_Paus_map.pdf"),
  width = 8,
  height = 10, limitsize = FALSE
)
```


```{r fig7, echo=FALSE, warning=FALSE, fig.height= 12, fig.width= 8}
# save subset of data to plot:
SKQ_env_Pheimii_df <- SKQ_env_metabar %>% filter(region=="18Sv9" & Species=="Pseudo-nitzschia_heimii")

# count the number of samples with pseudo-nitzschia:
SKQ_env_Pheimii_df %>% filter(pres=="1") %>% nrow()

# ggplot map
SKQ_env_Pheimii <- ggplot(data = state_prov) +
    geom_sf(border=1) +
    geom_point(data = SKQ_env_Pheimii_df %>% filter(pres=="1"), 
               aes(x = longitude, y = latitude, fill=pres), 
               size=4, shape=21, color="black") +
    coord_sf(xlim = c(-178,-150), ylim = c(55, 72), expand = FALSE) +
    scale_fill_manual(breaks = c("0", "1"), 
                       values=c("white", "#bc8a31"),
                       labels=c("Absent","Present")) +
   annotate("text", x=-167.5, y=69, label="n=2", size=5) + 
   facet_grid(. ~ Species) +
    theme_bw() +
    xlab("Longitude") +
    ylab("Latitude") +
    theme(axis.text=element_text(size=14, face="bold"),
        axis.title=element_text(size=14, face="bold"),
        legend.position = "none",
        text = element_text(size=14),
        strip.background=element_rect(fill="white"))

SKQ_env_Pheimii

ggsave(
  file = here::here("SKQ_Alaska_Analysis", "SKQ_Alaska_Figures", "SKQ_Pheimii_map.pdf"),
  width = 8,
  height = 10, limitsize = FALSE
)
```


```{r fig8, echo=FALSE, warning=FALSE, fig.height= 12, fig.width= 8}
# save subset of data to plot:
SKQ_env_Pdelact_df <- SKQ_env_metabar %>% filter(region=="ITS" & Species=="Pseudo-nitzschia_delicatissima")

# count the number of samples with pseudo-nitzschia:
Pdelact_samples<-SKQ_env_Pdelact_df %>% filter(pres=="1") %>% nrow()

# ggplot map
SKQ_env_Pdelact <- ggplot(data = state_prov) +
    geom_sf(border=1) +
    geom_point(data = SKQ_env_Pdelact_df %>% filter(pres=="1"), 
               aes(x = longitude, y = latitude, fill=pres), 
               size=4, shape=21, color="black") +
    coord_sf(xlim = c(-178,-150), ylim = c(55, 72), expand = FALSE) +
    scale_fill_manual(breaks = c("0", "1"), 
                       values=c("white", "#bc8a31"),
                       labels=c("Absent","Present")) +
   annotate("text", x=-167.5, y=69, label="n=874", size=5) + 
   facet_grid(. ~ Species) +
    theme_bw() +
    xlab("Longitude") +
    ylab("Latitude") +
    theme(axis.text=element_text(size=14, face="bold"),
        axis.title=element_text(size=14, face="bold"),
        legend.position = "none",
        text = element_text(size=14),
        strip.background=element_rect(fill="white"))

SKQ_env_Pdelact

ggsave(
  file = here::here("SKQ_Alaska_Analysis", "SKQ_Alaska_Figures", "SKQ_Pdelicat_map.pdf"),
  width = 8,
  height = 10, limitsize = FALSE
)
```


```{r fig9, echo=FALSE, warning=FALSE, fig.height= 12, fig.width= 8}
# save subset of data to plot:
SKQ_env_Ppung_df <- SKQ_env_metabar %>% filter(region=="ITS" & Species=="Pseudo-nitzschia_pungens")

# count the number of samples with pseudo-nitzschia:
SKQ_env_Ppung_df %>% filter(pres=="1") %>% nrow()

# ggplot map
SKQ_env_Ppung <- ggplot(data = state_prov) +
    geom_sf(border=1) +
    geom_point(data = SKQ_env_Ppung_df %>% filter(pres=="1"), 
               aes(x = longitude, y = latitude, fill=pres), 
               size=4, shape=21, color="black") +
    coord_sf(xlim = c(-178,-150), ylim = c(55, 72), expand = FALSE) +
    scale_fill_manual(breaks = c("0", "1"), 
                       values=c("white", "#bc8a31"),
                       labels=c("Absent","Present")) +
   annotate("text", x=-167.5, y=69, label="n=209", size=5) + 
   facet_grid(. ~ Species) +
    theme_bw() +
    xlab("Longitude") +
    ylab("Latitude") +
    theme(axis.text=element_text(size=14, face="bold"),
        axis.title=element_text(size=14, face="bold"),
        legend.position = "none",
        text = element_text(size=14),
        strip.background=element_rect(fill="white"))

SKQ_env_Ppung

ggsave(
  file = here::here("SKQ_Alaska_Analysis", "SKQ_Alaska_Figures", "SKQ_Ppung_map.pdf"),
  width = 8,
  height = 10, limitsize = FALSE
)
```



```{r fig10, echo=FALSE, warning=FALSE, fig.height= 12, fig.width= 8}
# save subset of data to plot:
SKQ_env_Pser_df <- SKQ_env_metabar %>% filter(region=="ITS" & Species=="Pseudo-nitzschia_seriata")

# count the number of samples with pseudo-nitzschia:
SKQ_env_Pser_df %>% filter(pres=="1") %>% nrow()

# ggplot map
SKQ_env_Pser <- ggplot(data = state_prov) +
    geom_sf(border=1) +
    geom_point(data = SKQ_env_Pser_df %>% filter(pres=="1"), 
               aes(x = longitude, y = latitude, fill=pres), 
               size=4, shape=21, color="black") +
    coord_sf(xlim = c(-178,-150), ylim = c(55, 72), expand = FALSE) +
    scale_fill_manual(breaks = c("0", "1"), 
                       values=c("white", "#bc8a31"),
                       labels=c("Absent","Present")) +
   annotate("text", x=-167.5, y=69, label="n=82", size=5) + 
   facet_grid(. ~ Species) +
    theme_bw() +
    xlab("Longitude") +
    ylab("Latitude") +
    theme(axis.text=element_text(size=14, face="bold"),
        axis.title=element_text(size=14, face="bold"),
        legend.position = "none",
        text = element_text(size=14),
        strip.background=element_rect(fill="white"))

SKQ_env_Pser

ggsave(
  file = here::here("SKQ_Alaska_Analysis", "SKQ_Alaska_Figures", "SKQ_Pser_map.pdf"),
  width = 8,
  height = 10, limitsize = FALSE
)
```


```{r fig11, echo=FALSE, warning=FALSE, fig.height= 12, fig.width= 8}
# save subset of data to plot:
SKQ_env_Pgran_df <- SKQ_env_metabar %>% filter(region=="ITS" & Species=="Pseudo-nitzschia_granii")

# count the number of samples with pseudo-nitzschia:
SKQ_env_Pgran_df %>% filter(pres=="1") %>% nrow()

# ggplot map
SKQ_env_Pgran <- ggplot(data = state_prov) +
    geom_sf(border=1) +
    geom_point(data = SKQ_env_Pgran_df %>% filter(pres=="1"), 
               aes(x = longitude, y = latitude, fill=pres), 
               size=4, shape=21, color="black") +
    coord_sf(xlim = c(-178,-150), ylim = c(55, 72), expand = FALSE) +
    scale_fill_manual(breaks = c("0", "1"), 
                       values=c("white", "#bc8a31"),
                       labels=c("Absent","Present")) +
   annotate("text", x=-167.5, y=69, label="n=127", size=5) + 
   facet_grid(. ~ Species) +
    theme_bw() +
    xlab("Longitude") +
    ylab("Latitude") +
    theme(axis.text=element_text(size=14, face="bold"),
        axis.title=element_text(size=14, face="bold"),
        legend.position = "none",
        text = element_text(size=14),
        strip.background=element_rect(fill="white"))

SKQ_env_Pgran

ggsave(
  file = here::here("SKQ_Alaska_Analysis", "SKQ_Alaska_Figures", "SKQ_Pgran_map.pdf"),
  width = 8,
  height = 10, limitsize = FALSE
)
```


```{r fig12, echo=FALSE, warning=FALSE, fig.height= 12, fig.width= 8}
# save subset of data to plot:
SKQ_env_Pturg_df <- SKQ_env_metabar %>% filter(region=="ITS" & Species=="Pseudo-nitzschia_turgidula")

# count the number of samples with pseudo-nitzschia:
SKQ_env_Pturg_df %>% filter(pres=="1") %>% nrow()

# ggplot map
SKQ_env_Pturg <- ggplot(data = state_prov) +
    geom_sf(border=1) +
    geom_point(data = SKQ_env_Pturg_df %>% filter(pres=="1"), 
               aes(x = longitude, y = latitude, fill=pres), 
               size=4, shape=21, color="black") +
    coord_sf(xlim = c(-178,-150), ylim = c(55, 72), expand = FALSE) +
    scale_fill_manual(breaks = c("0", "1"), 
                       values=c("white", "#bc8a31"),
                       labels=c("Absent","Present")) +
   annotate("text", x=-167.5, y=69, label="n=18", size=5) + 
   facet_grid(. ~ Species) +
    theme_bw() +
    xlab("Longitude") +
    ylab("Latitude") +
    theme(axis.text=element_text(size=14, face="bold"),
        axis.title=element_text(size=14, face="bold"),
        legend.position = "none",
        text = element_text(size=14),
        strip.background=element_rect(fill="white"))

SKQ_env_Pturg

ggsave(
  file = here::here("SKQ_Alaska_Analysis", "SKQ_Alaska_Figures", "SKQ_Pturg_map.pdf"),
  width = 8,
  height = 10, limitsize = FALSE
)
```

Pseudo-nitzschia_fryxelliana

```{r fig13, echo=FALSE, warning=FALSE, fig.height= 12, fig.width= 8}
# save subset of data to plot:
SKQ_env_Pfryx_df <- SKQ_env_metabar %>% filter(region=="ITS" & Species=="Pseudo-nitzschia_fryxelliana")

# count the number of samples with pseudo-nitzschia:
SKQ_env_Pfryx_df %>% filter(pres=="1") %>% nrow()

# ggplot map
SKQ_env_Pfryx <- ggplot(data = state_prov) +
    geom_sf(border=1) +
    geom_point(data = SKQ_env_Pfryx_df %>% filter(pres=="1"), 
               aes(x = longitude, y = latitude, fill=pres), 
               size=4, shape=21, color="black") +
    coord_sf(xlim = c(-178,-150), ylim = c(55, 72), expand = FALSE) +
    scale_fill_manual(breaks = c("0", "1"), 
                       values=c("white", "#bc8a31"),
                       labels=c("Absent","Present")) +
   annotate("text", x=-167.5, y=69, label="n=7", size=5) + 
   facet_grid(. ~ Species) +
    theme_bw() +
    xlab("Longitude") +
    ylab("Latitude") +
    theme(axis.text=element_text(size=14, face="bold"),
        axis.title=element_text(size=14, face="bold"),
        legend.position = "none",
        text = element_text(size=14),
        strip.background=element_rect(fill="white"))

SKQ_env_Pfryx

ggsave(
  file = here::here("SKQ_Alaska_Analysis", "SKQ_Alaska_Figures", "SKQ_Pfryx_map.pdf"),
  width = 8,
  height = 10, limitsize = FALSE
)
```

## Do we see any obvious blooms of Pn or other HABs?

Center log ratio transformation of phyto community data along transect

```{r,echo=FALSE, hide=TRUE, warning=FALSE}
# Load dataframe with information on phytoplankton vs protist groups:
Protist_phyto_gp<-read_csv(here("ITS_vs_18S","Class_filtering.csv"))
# keep list of only phytoplankton:
Phyto_gp_only<-Protist_phyto_gp %>% filter(phytoplankton=="yes") %>% dplyr::select("Class")

# read environmental csv data:
SKQ_env<-read_csv(here("SKQ_Alaska_Analysis","SKQ_Alaska_data","SKQ21_sample_data_20240627.csv"))
SKQ_env2 <- SKQ_env %>% 
  filter(Sample_Name!="END.NC.SKQ2021" & Sample_Name!="START.NC.SKQ2021") %>%
  rename(Sample_ID=Sample_Name) %>%
  mutate(Sample_ID = str_replace_all(Sample_ID, "\\.", "_")) %>%
  dplyr::select("Sample_ID","latitude","longitude")
# save as sample_data:
SKQ_env2_meta<-SKQ_env2 %>% column_to_rownames("Sample_ID") %>% sample_data()

# load metabarcoding data (ITS1, 18Sv4, 18Sv9)
SKQ_18S_ITS_df<-readRDS(here("ITS_vs_18S","JonahVentures_RUN2","SKQ_ITS1_18S_all.rds"))

# filter out phytos from dataset:
SKQ_metabar<-SKQ_18S_ITS_df %>%
  dplyr::select(-c("MP_E610_NC","MP_POSITIVE")) %>%
  # filter out samples that were sequenced for all three regions (ITS & 18S)
  pivot_longer(cols = MP_E1032_OC0821_PCR1:MP_E600_3B_PCR3, names_to = "Sample_ID", values_to = "counts") %>%
  filter(Class %in% Phyto_gp_only$Class) %>%
  mutate(Sample_ID = str_replace_all(Sample_ID, c("MP_"="","_PCR*"=""))) %>%
  filter(Sample_ID %in% SKQ_env2$Sample_ID)

# Save taxonomy for phyloseq object:
SKQ_ITS_taxa<- SKQ_metabar %>% as.data.frame() %>%
  filter(region=="ITS") %>%
  dplyr::select(c("Feature_ID":"Species","Kingdom","Phylum")) %>%
  unique() %>%
  remove_rownames() %>%
  column_to_rownames("Feature_ID") %>%
  dplyr::select("Kingdom","Phylum","Class","Order","Family","Genus","Species") %>%
  as.matrix()
# save as taxonomy table:
SKQ_ITS_taxa_ps<-tax_table(SKQ_ITS_taxa)

# Save counts for phyloseq object:
SKQ_ITS_counts<- SKQ_metabar %>% 
  filter(region=="ITS") %>%
  dplyr::select(c("Feature_ID","Sample_ID","counts")) %>%
  pivot_wider(names_from = "Sample_ID",values_from = "counts") %>%
  column_to_rownames("Feature_ID") %>%
  as.matrix()
# save as otu table:
SKQ_ITS_counts_ps<-otu_table(SKQ_ITS_counts, taxa_are_rows = TRUE)

# save ITS as phyloseq object:
SKQ_ITS_physeq<-phyloseq(SKQ_env2_meta,SKQ_ITS_taxa_ps,SKQ_ITS_counts_ps)

```

### Community Composition Bargraph

```{r fig14, warning=FALSE, message=FALSE}
# check out sample data to group by
sample_data(SKQ_ITS_physeq)
# add column to sample data for depth grouping:
SKQ_ITS_physeq2<-SKQ_ITS_physeq %>%
  ps_mutate(Latitude=round(as.numeric(latitude)),
            Longitude=round(as.numeric(longitude)))

# plot 
SKQ_ITS_physeq2 %>%
  tax_fix() %>%
   # tax_transform(trans = "rclr", rank = "unique") %>%
  # ps_arrange(desc("Latitude")) %>%
  comp_barplot(
    tax_level = "Species", n_taxa = 21, other_name = "Other",
    # label="Latitude",
    taxon_renamer = function(x) stringr::str_remove(x, " [ae]t rel."),
    palette = distinct_palette(n = 21, add = "grey90"),
    merge_other = FALSE, bar_outline_colour = "black"
  ) +
  coord_flip() +
   facet_grid(
    cols = vars(Longitude), rows = vars(Latitude),
    scales = "free_y", # this only frees y scale per row in grid faceting
    space = "free_y" # allows bars to be same size by freeing facet heights
  ) +
  # facet_wrap("Longitude", nrow = 1, scales = "free") +
  labs(x = NULL, y = NULL)

ggsave(
  file = here::here("SKQ_Alaska_Analysis", "SKQ_Alaska_Figures", "SKQ_CommunityComp_bargraph_ITS1.pdf"),
  width = 25,
  height = 15, limitsize = FALSE
)
```

### Stacked bargraphs on map of sites

```{r fig15, echo=FALSE, warning=FALSE, fig.height= 12, fig.width= 8}
# create variable to merge samples by in R
variable1 = as.character(get_variable(SKQ_ITS_physeq2, "Latitude"))
variable2 = as.character(get_variable(SKQ_ITS_physeq2, "Longitude"))
sample_data(SKQ_ITS_physeq2)$Lat_Lon <- mapply(paste0, variable1, variable2, collapse = "_")
merge_samples(SKQ_ITS_physeq2, "Lat_Lon")


# Create bargraph of community composition can we see HABs abundance change with sites
SKQ_ITS_physeq2_species<-SKQ_ITS_physeq2 %>%
  tax_fix() %>%
  tax_agg("Species") %>%
#  ps_mutate(Sample_ID= rownames(sample_data(SKQ_ITS_physeq2))) %>%
  merge_samples("Lat_Lon") %>%
  tax_transform("compositional") %>%
  psmelt()

# Filter for abundances greater than 1:
SKQ_ITS_physeq2_species_gt1<- SKQ_ITS_physeq2_species %>%
  group_by(Sample, Latitude, Longitude) %>%
filter(Abundance > 0.01)

# Sum remaining taxa with a relative abundance < 1% and make a new dataframe
Remainders_sed <- (SKQ_ITS_physeq2_species_gt1) %>%
  summarise(Abundance = (1-sum(Abundance))) %>% 
  as.data.frame()
Remainders_sed$Species<-"Species < 1%"
Remainders_sed$Phylum<-"_Species < 1%"
Remainders_sed$Class<-"_Species < 1%"
Remainders_sed$Order<-"_Species < 1%"
Remainders_sed$Genus<-"_Species < 1%"

# Join dataframes
SKQ_ITS_physeq2_species_bar <- full_join(SKQ_ITS_physeq2_species_gt1, Remainders_sed)

# Plot 
ggplot(data=SKQ_ITS_physeq2_species_bar , aes(x = Sample, y = Abundance, fill = Species)) + 
  geom_bar(stat = "identity", color='black') +
  scale_fill_manual(values=colorRampPalette(brewer.pal(12, "Paired"))(19)) +
  # Remove x axis title
  theme() +
  guides() +
  ylab("% Relative Abundance") +
  xlab("Lat_Lon") +
  theme_classic() +
  scale_y_continuous(expand = c(0,0)) +
  theme(plot.title = element_text(hjust=0.5, size=14),
        axis.text.x=element_text(angle=45, hjust=1))

ggsave(
  file = here::here("SKQ_Alaska_Analysis", "SKQ_Alaska_Figures", "SKQ_CommunityComp_bargraph3_ITS1.pdf"),
  width = 25,
  height = 15, limitsize = FALSE
)
```

### Add piechart to maps:

```{r fig16, echo=FALSE, warning=FALSE, fig.height= 12, fig.width= 8}
# Species color palette
Species_cols<-c("#d6c300","#3032b2","#91da4a","#00238e","#7ddb7c","#cf7cfe","#019257","#cc003e","#00a0e6","#f88c10","#220358","#ff8941","#7499ff","#b60304","#0b3500","#ff5a4e","#666d00","#b7384d","#935f44")

# Formatting rel abnd data:
SKQ_ITS_physeq2_species_bar2<-SKQ_ITS_physeq2_species_bar %>% 
  mutate(value=Abundance)
# Alaska map
Alaska_map<-ggplot(data = state_prov) +
    geom_sf(border=1) +
    coord_sf(xlim = c(-178,-152), ylim = c(55, 75))

# map with bargraphs:
Alaska_map + 
  # ggplot() +
  geom_scatterpie(aes(x=Longitude, y=Latitude, r=0.8), data=as.data.frame(SKQ_ITS_physeq2_species_bar2), 
                  cols="Species", color="black", long_format = TRUE) +
  scale_fill_manual(values=colorRampPalette(brewer.pal(12, "Paired"))(19)) +
  # scale_fill_manual(values=Species_cols) +
  theme_bw() +
  labs(fill="Species") +
  # scale_y_continuous(expand = c(0,0)) +
  theme(plot.title = element_text(hjust=0.5, size=14),
        axis.text.x=element_text(angle=45, hjust=1))

ggsave(
 file = here::here("SKQ_Alaska_Analysis", "SKQ_Alaska_Figures", "SKQ_ITS1_SpComp_map.pdf"),
 width = 12,
 height = 10, limitsize = FALSE
)
```

Separate grid of just pie charts:

```{r fig17, echo=FALSE, warning=FALSE, fig.height= 12, fig.width= 8}
# map with bargraphs:
ggplot() +
  geom_scatterpie(aes(x=Longitude, y=Latitude, r=0.8), data=as.data.frame(SKQ_ITS_physeq2_species_bar2), 
                  cols="Species", color="black", long_format = TRUE) +
  scale_fill_manual(values=colorRampPalette(brewer.pal(12, "Paired"))(19)) +
  coord_fixed() +
  # scale_fill_manual(values=Species_cols) +
  theme_bw() +
  labs(fill="Species") +
  # scale_y_continuous(expand = c(0,0)) +
  theme(plot.title = element_text(hjust=0.5, size=14),
        axis.text.x=element_text(angle=45, hjust=1))

ggsave(
 file = here::here("SKQ_Alaska_Analysis", "SKQ_Alaska_Figures", "SKQ_ITS1_SpComp_Piecharts.pdf"),
 width = 15,
 height = 10, limitsize = FALSE
)
```


Composition heatmap:

```{r fig18, warning=FALSE, message=FALSE}
SKQ_ITS_physeq2 %>%
    tax_fix() %>%
  tax_transform(trans = "rclr", rank = "Genus") %>%
  tax_filter(min_prevalence = 0.1, use_counts = TRUE) %>%
  comp_heatmap(
    colors = heat_palette(sym = TRUE), grid_col = NA,
    sample_side = "bottom", name = "Robust\nCLR",
    sample_anno = sampleAnnotation(
      "Latitude" = anno_sample("Latitude")
       # col = list("Birth mode" = c(
     #   "CS Elective" = "black", "CS Emergency" = "orange", "Vaginal" = "lightgrey"
     # ))
    )
  )
```

#### Software and package versions used for file:

```{r warning=FALSE}
print(sessionInfo(), locale = FALSE)
```




